### 2023-08-18 docs/appendix.Rmd
add appendix.Rmd (only contains some fragments at the moment)

## General principles in power analysis

```{r, message=FALSE}
library(tidyverse)
set.seed(1)
dat <- simulate(n_subjects = 100, n_items = 50)
dat$expert <- factor(dat$expert)
dat$advice_present <- factor(dat$advice_present)
dat$advice_correct <- factor(dat$advice_correct)
```

```{r, message=FALSE}
library(lme4)
f <- y_bin ~ 1 + expert + advice_present + advice_correct + 
  expert:advice_present + expert:advice_correct +
  (1|subject) + (1|item)
fit <- glmer(f, data = dat, family = "binomial")
```

```{r}
library(marginaleffects)
marginal_means(fit)
avg_comparisons(fit, variables = list(expert = c(0, 1)))
comparisons(fit,
  newdata = datagrid(expert = c("0", "1"), advice_present = c("0", "1"), advice_correct = c("0", "1"))
)
```

## Sample size justification based on precision: simulate expected confidence interval width

```{r}
simulate <- function(n_subjects = 100, n_items = 50,
  b_0 = 0.847, b_e = 1.350, b_a = -1.253, b_c = 2.603,
  b_ea = 0.790, b_ec = -1.393,
  sd_u0s = 0.5, sd_u0i = 0.5, ...){
  require(dplyr)
  require(faux)
  # simulate design
  dat <- add_random(subject = n_subjects, item = n_items) %>%
    add_between("subject", expert = c(1, 0), .prob = c(0.25, 0.75)) %>%
    mutate(advice_present = rbinom(n(), 1, prob = 2/3)) %>%
    mutate(advice_correct = if_else(advice_present == 1, 
                                    rbinom(n(), 1, prob = 0.8), 0)) %>%
    # add random effects
    add_ranef("subject", u0s = sd_u0s) %>%
    add_ranef("item", u0i = sd_u0i) %>%
    # compute dependent variable
    mutate(linpred = b_0 + u0i + u0s +
        b_e * expert + b_a * advice_present + b_c * advice_correct +
        b_ea * expert * advice_present + b_ec * expert * advice_correct) %>%
    mutate(y_prob = plogis(linpred)) %>%
    mutate(y_bin = rbinom(n = n(), size = 1, prob = y_prob))
  dat
}

sim_and_analyse <- function(
  formula_chr = "y_bin ~ 1 + expert + advice_present + advice_correct + 
    expert:advice_present + expert:advice_correct + (1|subject) + (1|item)", 
  null_hypotheses = c("expert:advice_correct = 0"), ...){
  require(lme4)
  require(multcomp)
  # simulate data
  dat <- simulate(...)
  # fit model
  model <- glmer(as.formula(formula_chr), data = dat, family = "binomial")
  # compute p-values
  glht <- glht(model, linfct = null_hypotheses)
  confint(glht, level = 0.95, calpha = univariate_calpha())$confint
}
```

```{r, message=FALSE, eval=TRUE}
library(future)
plan("multisession", workers = 6)
set.seed(2)
```

```{r, message=FALSE, eval=TURE}
library(tidyverse)
library(furrr)
sim_design <- crossing(
  rep = 1:200,
  n_subjects = c(150, 200),
  n_items = c(30, 50)
) %>%
  mutate(ci = future_pmap(., sim_and_analyse, 
    .options = furrr_options(seed = TRUE))) %>%
  unnest_wider(ci)
```

```{r, eval=FALSE}
precision <- sim_design %>%
  group_by(n_subjects, n_items) %>% 
  mutate(width = `expert:advice_correct`[,3] - `expert:advice_correct`[,2]) %>%
  summarise(precision = mean(width),
    ci.lwr = t.test(width)$conf.int[1],
    ci.upr = t.test(width)$conf.int[2], 
    .groups = "drop")
precision %>%
  mutate(across(c(n_subjects, n_items), factor)) %>%
  ggplot(aes(n_subjects, n_items, fill = precision)) +
  geom_tile() +
  geom_text(aes(label = sprintf("%.2f \n [%.2f; %.2f]", 
                                precision, ci.lwr, ci.upr)), 
    color = "white", size = 6) +
  scale_fill_viridis_c(limits = c(0, 2))
```