### 2024-03-27 docs/manuscript.Rmd
work on the "compute the estimate" section

---
title             : "A Tutorial on Tailored Simulation-Based Sample Size Planning for Experimental Designs with Generalized Linear Mixed Models"
shorttitle        : "Simulation-Based Sample Size Planning for GLMMs"
author:

  - name          : "Timo K. Koch"
    affiliation   : "1, 2"
    corresponding : yes    # Define only one corresponding author
    address       : "Torstrasse 25, 9000 St. Gallen, Switzerland"
    email         : "timo.koch@unisg.ch"
    equal_contrib: yes
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - "Conceptualization"
      - "Formal Analysis"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Florian Pargent"
    affiliation   : "1"
    equal_contrib: yes
    role:
      - "Formal Analysis"
      - "Methodology"
      - "Visualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Anne-Kathrin Kleine"
    affiliation   : "1"
    role:
      - "Formal Analysis"      
      - "Writing - Review & Editing"
  - name          : "Eva Lermer"
    affiliation   : "1,3"
    role:
      - "Conceptualization"
      - "Resources"
      - "Writing - Review & Editing"
  - name          : "Susanne Gaube"
    affiliation   : "1,4"
    role:
      - "Conceptualization"
      - "Supervision"
      - "Writing - Review & Editing"
affiliation:
  - id            : "1"
    institution   : "LMU Munich, Department of Psychology"
  - id            : "2"
    institution   : "University of St. Gallen, Institute of Behavioral Science & Technology"
  - id            : "3"
    institution   : "Technical University of Applied Sciences Augsburg, Department of Business Psychology"
  - id            : "4"
    institution   : "University College London, Global Business School for Health"
note: "\\clearpage" #added this to move entire author note to second page
authornote: |
  This research was funded by a grant from the Volkswagen Foundation (Grant No. 98525).
  All materials (reproducible manuscript incl. R code, simulation results) are available in the project's repository on the Open Science Framework (OSF): https://osf.io/dhwf4/
abstract: |
  When planning experimental research, determining an appropriate sample size and using suitable statistical models are crucial for robust and informative results. However, the recent replication crisis underlines the need for more rigorous statistical methodology and well-powered designs. Generalized linear mixed models (GLMMs) offer a flexible statistical framework to analyze experimental data with complex (e.g., dependent and hierarchical) data structures. Yet, available methods and software for a priori power analyses for GLMMs are often limited to specific designs, while data simulation approaches offer more flexibility. Based on a practical case study, the current tutorial equips researchers with a step-by-step guide and corresponding code for conducting tailored a priori power analyses to determine appropriate sample sizes with GLMMs. Finally, we give an outlook on the increasing importance of simulation-based power analysis in experimental research.

keywords          : "power analysis, data simulation, sample size, generalized linear mixed model"
wordcount         : "7504"
bibliography      : ["../Simulation Tutorial Paper.bib", "r-references.bib"]
floatsintext      : yes
linenumbers       : no
draft             : no
mask              : no
figurelist        : no
tablelist         : no
footnotelist      : no
classoption       : "man"
output: 
  papaja::apa6_pdf:
---

```{r setup, include = FALSE}
library("papaja")
library("tidyverse")
library("faux")
library("lme4")
library("multcomp")
library("marginaleffects")
library("future")
library("furrr")
library("binom")
r_refs("r-references.bib")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed, echo = TRUE, eval = FALSE, fig.keep = "high", dev = "jpeg")
```

# Introduction

<!-- General short intro paragraph, need for well-powered and analyzed experimental research -->
When planning experimental research, it is essential to determine an appropriate sample size to ensure that the results obtained are both robust and informative, and to use appropriate statistical models to analyze the data [@lakensSampleSizeJustification2022]. However, the recent replication crisis has illustrated many challenges surrounding the reproducibility and reliability of findings [@yarkoniGeneralizabilityCrisis2022]. As a result, there is a growing need for more rigorous statistical methodology and the adoption of well-powered experimental designs. While easy-to-use software solutions exist for simple statistical models and experimental designs, many researchers lack the skills and tools to conduct "a priori" (i.e., before data collection) sample size planning for more complex research designs using the flexible generalized linear mixed models (GLMM) framework in order to determine the required sample size in their experiments. In the present work, we provide a tutorial on performing tailored simulation based sample size planning for GLMMs. After introducing general concepts, we review existing software solutions in R and discuss under which circumstances tailored data simulations are necessary. Then we describe general steps and decisions that are necessary whenever performing tailored data simulation. Because the details of these steps vary considerably for each specific study, we finish with a case study from the field of human-AI (artificial intelligence) interaction research.

<!-- <!-- Old Summary / Wrap-Up, pick up main arguments and what the paper provides, keep this short! --> -->
<!-- Based on the need for well-powered experimental research using GLMMs and the lack of tools to conduct corresponding power analyses, in this tutorial paper, we present a case study that serves as a practical demonstration of how to perform a simulation-based a priori power analysis with GLMMs. Thereby, we aim to equip researchers with the tools needed to simulate data and determine appropriate sample sizes for their own research. -->

# Prior Knowledge

Simulation-based sample size planning for generalized linear models is a complex topic that touches on a variety of statistical concepts and programming skills.
Although we give some short introduction to clarify our terminology, to profit the most from our tutorial, our readers require prior knowledge in the following topics (or refer to our collection of background literature):

Readers should understand basic statistical concepts like hypothesis tests and their statistical power but also confidence intervals and their precision.
Some background knowledge in empirical estimands, statistical estimates, and causal inference is very useful but not absolutely necessary.

In addition, readers should have basic programming skills in R on how to simulate data and how to repeat these simulations in parallel to speed up computation.
While there are a myriad of options for data simulation in R beyond the basic functionality in the base and stats packages, in our tutorial we use functions from the tidyverse and the faux package.

Finally, readers should be familiar with regression modeling in general and GLMMs in particular.
In our tutorial, we simulate data by manually specifying the model equation of a GLMM that represents our assumed data generating process.
While it is not necessary to understand the technical details of how GLMMs are estimated, it is crucial to understand the structure of a basic GLMM (e.g., logistic regression with random intercepts) and how the model assumes that the values in the dependent variable are determined by the predictor variables and the random effects.

Literature:

- Lakens ebook evtl. Andy Field für die Basics?
- R for data science, Bruine article, faux documentation für die R skills
- kumle power tutorial und irgendwelche paper (am besten aus AMPPS) zu GLMMs
- what is your estimand? paper für das estimand topic

# General concepts

## Planning for statistical power or precision

Conducting research with too small sample sizes can have many negative consequences.
First, experiments may yield inconclusive or misleading results, hindering the accumulation of knowledge.
Second, studies that are doomed to never finding a postulated effect due to small sample size waste resources by consuming time, effort, and funding without delivering meaningful results.
For such reasons, many journals and funding agencies now require that a justification of sample size is included in study protocols and grant proposals, recognizing its significance in ensuring robust and meaningful findings.
Although many pragmatic justifications are possible [@lakensSampleSizeJustification2022], the ideal is often considered to determine a suitable sample size a priori (i.e., before the study is conducted) based on some meaningful computation to assure that the study will be able to fulfill its purpose.
In this way, adding a solid sample size calculation to the research process can act as a safeguard for ensuring high-quality research.

<!-- <!-- advantages of an a priori power analysis --> -->
<!-- Thereby, a priori power analysis offers a valuable contribution to the research process by allowing researchers to estimate the appropriate sample sizes required to achieve sufficient statistical power for results with high evidential value. -->
<!-- Moreover, conducting a careful a priori power analysis helps researchers decide which experimental design and statistical models are both feasible and appropriate for analyzing the data and answering their research questions. -->
<!-- Also, when conducting a proper power analysis, researchers have to consider every aspect of the experimental design and will notice statistical or design challenges before starting with data collection.  -->
<!-- Adding a solid sample size calculation to the research process can act as a safeguard for ensuring high-quality research. -->
<!-- Finally, many journals and funding agencies now require that a power analysis is included in study protocols and grant proposals, recognizing its significance in ensuring robust and meaningful findings. -->

The majority of empirical studies in psychology applies hypothesis testing.
As a consequence, the dominant approach for sample size planning is based on power analysis (i.e., planing for power).
However, sample size planning can also be based on the precision of parameter estimates (i.e., planning for precision).

### Planning for power

<!-- Some basics on statistical power -->
In empirical research relying on hypothesis testing, the most common strategy for determining an adequate sample size is based on statistical power [@lakensSampleSizeJustification2022].
Statistical power is defined as the probability that a hypothesis test has a significant p-value when analyzing repeated samples from a population with a true effect of some pre-specified size.
Less formally, power is described as the probability that a hypothesis test correctly rejects the null hypothesis when the alternative hypothesis is true.
If the sample size (i.e., the number of participants and/or stimuli) used for data collection is insufficient to detect the effects or relationships being investigated with high probability, the study would be considered "underpowered".

When planning for power, a target is set for the statistical power of a hypothesis test of interest (a very common heuristic is 0.8).
Assuming some effect size of interest and a desired significance level, a minimum sample size can be determined that (on average) would guarantee reaching this target.

### Planning for precision

Not all research questions are best answered using hypothesis testing.
Especially for exploratory studies when little previous research has been conducted, only estimating effects of interest is often more useful.
When no hypothesis tests are conducted when analysing the planned study, power analysis is not not available for sample size planning.
Nonetheless, sample size still has a crucial effect on how informative the planned study will be to learn about the effect of interest.
In this case, the expected width of a confidence interval is used as a target for sample size planning.
The values inside a confidence interval are commonly interpreted as plausible values for the quantity of interest the confidence interval is supposed to estimate.
More formally, a 95%-confidence interval provides the smallest interval with the property that upon repeated sampling, 95% of individual confidence intervals would include the true quantity of interest.
Thus, a narrow confidence interval (assuming the confidence level is kept at a high value) is more informative about the size of the true effect than a wide interval.
Apart from the desired confidence level, the width of a confidence interval depends strongly on the sample size.
Because bigger samples carry more information, they lead to smaller confidence intervals (assuming everything else stays the same).
When planning for precision, a target is set for the expected width of a confidence interval of interest.
Assuming some effect size of interest and a desired confidence level, a minimum sample size can be determined that (on average) would guarantee reaching this target.

## Generalized linear mixed models (GLMMs)

<!-- Advantages of GLMMs over "simpler" models -->
As study designs become more complex, psychological researchers require more sophisticated statistical models to capture the nuanced relationships and grouping structures introduced by their study designs [@yarkoniGeneralizabilityCrisis2022].
GLMMs (also called multilevel models) are gaining increasing popularity in analyzing data because they offer a flexible framework for handling data with outcome variables that are not normally distributed (e.g., categorical outcomes) while accounting for both fixed and random effects [@fahrmeirRegressionModelsMethods2021].

<!-- intro into GLMMs -->
GLMMs are an extension of LMMs (Linear Mixed Models), which are, in turn, extensions of linear regression models that account for correlated data including hierarchical structures  [@fahrmeirRegressionModelsMethods2021]. 
In this context, correlated data means that the value in the outcome variable for one observation may be related (i.e., more similar or less similar) to the value of another observation in a systematic way that is not already accounted for by the usual (fixed) predictor variables (e.g., age of participants).
This correlation can arise for various reasons: Responses to some stimuli from some participants might be more similar because the same person was measured twice (repeated measurements), both participants come from the same neighborhood (clustering) or both participants responded to the same stimulus (stimulus effects). 
Thus, modeling such correlations is especially important whenever the data has a clear structure, while the grouping variables can be hierarchically organized (e.g., students nested in schools, schools nested in districts) or not (e.g., students solve math exercises, but neither student sees all exercises).
LMMs are used when the outcome variable is continuous and follows a normal distribution (when conditioned on all fixed and random effects).
They allow for the modeling of fixed effects, which capture the relationships between our usual predictors and the outcome, as well as random effects, which account for the different types of correlation structure and grouping effects exemplified above.
Random effects are typically assumed to follow a normal distribution with a mean of zero and a variance that quantifies the heterogeneity across groups.

As mentioned, GLMMs extend the LMM framework to accommodate non-normally distributed continuous and categorical outcome variables.
GLMMs incorporate both fixed and random effects, similar to LMMs, but also involve a link function that connects the linear combination of predictor variables to the expected value of the outcome variable.
The link function allows for modeling the relationship between predictors and the outcome in a non-linear way that is appropriate for the specific distribution family of the outcome variable.
As an example, think of an experiment with different design factors (e.g., picture, headline) impacting the likelihood of users clicking on an online advertisement.
Here, participants’ behavior is measured repeatedly (e.g., over several sessions).
The click patterns of participants in one session are likely to be correlated with their previous sessions.
Finally, the outcome variable is binary (click/no click) for each session, which follows a binomial distribution.

# Simulation based sample size planning with GLMMs

For simpler statistical models, like t-tests, ANOVA, and linear regression, with common study designs (e.g., mean comparison between two groups), user-friendly software for a priori sample size planning is readily available [@faulStatisticalPowerAnalyses2009; @champelyPackagePwr2018; @kelleyObtainingPowerObtaining2003; @kelleySampleSizePlanning2006]. 
However, these software packages are often not flexible enough to perform sample size planning for complex designs.

Existing approaches for sample size planning with GLMMs have almost exlusively focused on planning for power. 
Power analysis methods for multilevel models can be categorized into formula-based methods and simulation-based methods [@murayamaSummarystatisticsbasedPowerAnalysis2022].
Formula-based methods rely on often complicated formulas that can be used to directly calculate power while simulation-based methods rely on repeatedly simulating data with a known true effect size and estimating power empirically (i.e., how often the hypothesis test is significant for the simulated data).
Currently available formula-based software packages for power analysis with multilevel models often do not include GLMMs or are limited to very simple designs [@westfallStatisticalPowerOptimal2014; @murayamaSummarystatisticsbasedPowerAnalysis2022], making it necessary to build data simulations tailored specifically to the study design. 
A number of tutorials have been published describing how to perform such simulation-based power analysis for multilevel models [@arendStatisticalPowerTwolevel2019; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2022; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021]. 
However, most of these tutorials focus on linear mixed models (LMMs) and the most common designs [but see @kumleEstimatingPowerGeneralized2021 for a tutorial that also covers more advanced settings]. This narrow focus provides limited guidance for researchers faced with more complex study designs, especially when little prior knowledge about plausible effect sizes is available [see the discussion in @kumleEstimatingPowerGeneralized2021]. 
The necessary presumptions for simulation-based power analysis with GLMMs include assumptions about the distributional form of the outcome variable, the random effects, and the correlation structure. The distributional assumption specifies the distributional family for the outcome variable (when conditioned on all fixed and random effects). Assumptions about the random effects include the assumption of normality (i.e., that the random effects follow a normal distribution) and the covariance structure among the random effects (i.e.,  if and how they are correlated). Interpreting these presumptions entails understanding the underlying presumptions of the model and ensuring they align with the characteristics of the data being analyzed.
Existing tutorials often rely on heuristics for specifying variance components (e.g., the standard deviation of random intercepts) or assume that results from meta-analyses or data from pilot studies are available to determine plausible values for all model parameters. However, in practice, knowledge about those parameters from prior studies is often limited, which makes specifying assumptions a practical challenge [see the discussion in @maxwellSampleSizePlanning2008 and @kumleEstimatingPowerGeneralized2021].

- Insert table on existing R packages for power analysis with GLMMs

# When do we profit from tailored data simulation?

Performing tailored simulation based sample size planning is more complicated and time consuming than using existing software tools for sample size planning.
Thus, it is worth discussing under which circumstances we most profit from or are forced to perform tailored simulation based sample size planning, because existing software solutions do not meet our requirements.

## Complex study designs

Real studies are often more complex than the simplified designs assumed by many user-friendly software packages for sample size planning.

### Missing data

Missing data is a frequent issue in applied data analysis.
There can be various reasons for why data is missing.
For example data can be missing completely at random.
Alternatively, subjects might drop out or produce missing data as a result of some covariate also measured in the dataset.
A a different note, many experimental designs contain some conditions in which some values of the predictor variables are missing by design.
Missing by design conditions can make data analysis more complicated because predictors have to be coded in highly specific ways that prevent the estimated GLMM to become unidentified.
Whether missing data has an effect on the result of sample size planning depends on the characteristics of the missing data process.
However, except for very simple scenarios, deciding whether missing data can be safely ignored in the sample size planning is often difficult to decide on a theoretical basis.
A better strategy is to explicitly include the (assumed) process of missing data into the data simulation process.

### Data generating processes beyond standard GLMMs

While GLMMs are a very flexible model class that can handle a large variety of outcome variables, psychological researchers are becoming increasingly aware that many psychological datasets might profit from even more sophisticated models.
Common examples are zero-inflated outcomes, censoring, and nonlinear predictor effects.

    - possibility to extend beyond standard GLMMs (e.g. zero-inflated outcomes, censoring, nonlinear effects, …)

## Complex statistical hypotheses

The most common hypotheses tested in psychological research are of the type $H_0: \beta = 0$, where $\beta$ is some coefficient in a regression model.
However, many research questions in psychology actually require testing more complex statistical hypotheses.

### Directed hypotheses and testing against values other than 0

In the new era of preregistration and registered reports, most research questions are tested with directed hypotheses, because good theories should at least postulate whether some psychological effect of interest is positive or negative.
Even better theories should be able to specify a smallest effect sizes of interest (SESOI) that must be exceeded if the psychological effect has some practical relevance in a specific application.
In combination, this might require a test like $H_0: \beta \leq 0.1$.

### Combined hypotheses

Elaborate research questions often require testing hypotheses that consist of a combination of model parameters, for example $H_0: \beta_0 + \beta_1 \leq 0$.
If the research question consists only of a single hypothesis of this sort, it is often possible to reduce the hypothesis to a single regression coefficient by clever coding and/or centering of predictor variables.
However, interesting research questions often consist of combined hypotheses that consist of more than one separate statistical hypotheses.
For example, a combined null hypothesis $H_0$ might consist of two single null hypotheses $H_{01}: \beta_1 \leq 0$ and $H_{02}: \beta_0 + \beta_1 \leq 0$.
For some research questions, the combined null hypothesis $H_0$ would be rejected if both $H_{01}$ *AND* $H_{02}$ are rejected.
For other research questions, the combined null hypothesis $H_0$ would be rejected if $H_{01}$ *OR* $H_{02}$ *OR* both are rejected.
If the global hypothesis $H_0$ is combined with *OR*, the p-values of the single hypotheses must be corrected for multiple testing to avoid $\alpha$-inflation for the global hypothesis.
However, if the global hypothesis $H_0$ is combined with *AND*, a correction for multiple testing is not necessary but rather a mistake that unnecessarily reduces the power of the global hypothesis test.

None of the software packages for sample size planning in table X can handle combined hypotheses as discussed here.
In contrast, our case study will demonstrate how we can test directed combined hypotheses with tailored simulation based sample size planning.

## Planning for precision

As discussed earlier, the two major frameworks for sample size planning are planning for power and planning for precision.
Although planning for precision has been increasingly discussed as a more suitable strategy for exploratory research, all available software packages for sample size planning with GLMMs are based on power analysis.

As long as one were interested only in confidence intervals of a single model parameter, it might be possible to adapt existing software packages to plan for precision.
However, we demonstrate in our case study that in realistic research projects with GLMMs, the contrasts of interests and their respective confidence intervals are usually situated on the scale of the (expected) outcome in contrast to the scale of the model parameters.
We will show that tailored simulation based sample size planning can easily handle the planning for precision scenario.
The only change in procedure is that instead of computing hypothesis tests for each simulated dataset and estimating statistical power across repetitions, confidence intervals are computed for each simulated dataset and the expected width is estimated.

## No prior studies or pilot data

All frameworks for sample size planning require the user to make assumptions about the expected effect size.
Assuming the true effect is of this size (or greater), one can compute the (minimum) power of a hypothesis test or the (maximum) expected width of a confidence interval.
Existing software packages for sample size planning for GLMMs usually require to provide the assumed effect in the unit of some standardized measure of effect size.
When the researcher has access to similar studies or pilot data, providing such standardized effect sizes is feasible.
However, this requirement can be an almost impossible challenge when no prior studies of pilot data are available.
This problem is further exacerbated by the fact that GLMMs are so flexible, that general heuristics of what should be considered a small effect do not exist or are difficult to defend.
In our experience, using domain knowledge to construct a tailored data simulation is the only solution to determine plausible effect sizes in the absence of prior evidence.
It would be possible to use these tailored simulations to extract plausible values for standardized effect sizes that could then be inserted in existing software packages for sample size planning.
However, we would argue that when tailored data simulations are necessary anyways, performing the whole desing analysis in a customized way is preferred over using the existing software packages.

# General steps in tailored simulation-based sample size planning

Although the details differ depending on the specific study characteristics, each tailored simulation based sample size planning requires a series of steps and decisions that we will outline in this section.

## Define the estimand

### THEORY

The first step in every research process should always be a clear definition of which theoretical quantity is necessary to answer the specific research question of the study under planning.
Unfortunately, many psychological researchers are still not well trained in this task.
The theoretical estimand is a unit specific quantity defined outside of any statistical model.
Depending on the research and whether the research question is causal, this quantity might be counterfactual and not directly observable.
Part of the theoretical estimand is also a clear definition of the target population, for which the quantity is of interest.

When the theoretical estimand has been specified, the next step is to define the empirical estimand, which reexpresses the theoretical estimand in a way that can be estimated based on the observed data that will be collected in the planned study.
For many common research questions in psychology, the empirical estimand can be expressed as a statistical quantities that can be estimated with a regression model.
Note however, that this is not always possible and that the literature on causal inference includes many estimation strategies beyond regression.

When the empirical estimand has been specified, the last step is to clearly define how a concrete estimate of the empirical estimand shall be computed based on the observed study data.
When a regression model is used for estimation, this includes decisions like which statistical framework and package is used to fit the model, and how to compute hypothesis tests or confidence intervals.

@lundbergWhatYourEstimand2021

### PRACTICE

In the present case study, we consider the effectiveness of feedback provided by an artificial intelligence (AI) in an AI-enabled diagnostic decision support system.
The context is a clinical setting, where expert radiologists and students under training must detect bleeding based on head scans from computer tomography.
In the investigated AI-enabled diagnostic decision support system, an AI model can provide initial diagnostic advice, which can be used as guidance by the humans require to make the final diagnostic decision.
The research goal is to validate the effectiveness of the AI-enabled advice.
We consider the AI-enable as effective, if the following pattern holds, which we will here describe verbally:

*We expect that for BOTH expert radiologists and students under training, correct AI-advice leads to a higher probability of accurately diagnosing a CT scan compared to no AI-advice presented, AND, we expect that for BOTH experts and non-experts, incorrect advice leads to a lower probability of accurately diagnosing a CT scan compared to no advice presented.*

I becomes clear that our theoretical estimand is be a causal quantity, based on individual quantities define outside of any statistical model.
To explain why, we focus on only on part of the research question concerning the comparison between correct AI-advice and no AI-advice for expert radiologists.
For a specific expert radiologist and a specific CT scan, our question is concerned with the difference between the probability that a correct diagnosis is made if correct AI-advice were presented and the probability that a correct diagnosis is made if no AI-advice were presented.
Mathematically, we could describe this unit-specific quantity with some pseudo-equation like this (where $s$ stands for a specific expert radiologist and $i$ stands for a specific CT scan):

$$
\begin{aligned}
& P(\text{correct diagnosis} | \text{correct advice, a specific expert, a specific scan}) \\
& \quad - P(\text{correct diagnosis} | \text{no advice, a specific expert, a specific scan})
\end{aligned}
$$
However, our verbal description is not precise enough yet, because the above quantity exists and may differ for each combination of expert and scan.
We must specify how we want to aggregate the quantity across different experts and scans.
If we would compute the average of our individual quantity across all experts and scans in the target population, the result would be called the average treatment effect (ATE) in the population of experts.
Although often debated whether this is a good idea, psychological research often focuses on a different quantity.
Instead of averaging over experts and scans, we focus in a hypothetical *typical* expert and a *typical* scan, where *typical* is usually defined as an average score on all attributes of the expert or scan.
With this definition, the theoretical estimand in our above example can be expressed as:

$$
\begin{aligned}
& P(\text{correct diagnosis} | \text{correct advice, an average expert, an average scan}) \\
& \quad - P(\text{correct diagnosis} | \text{no advice, an average expert, an average scan})
\end{aligned}
$$
In total, our theoretical estimand consists of four expressions like the one above, the remaining three are:

$$
\begin{aligned}
& P(\text{correct diagnosis} | \text{no advice, an average expert, an average scan}) \\
& \quad - P(\text{correct diagnosis} | \text{wrong advice, an average expert, an average scan})
\end{aligned}
$$
$$
\begin{aligned}
& P(\text{correct diagnosis} | \text{correct advice, an average student, an average scan}) \\
& \quad - P(\text{correct diagnosis} | \text{no advice, an average student, an average scan})
\end{aligned}
$$

$$
\begin{aligned}
& P(\text{correct diagnosis} | \text{no advice, an average student, an average scan}) \\
& \quad - P(\text{correct diagnosis} | \text{wrong advice, an average student, an average scan})
\end{aligned}
$$
To complete our definition of the theoretical estimand, we have to clearly define out target population that consists of both persons and scans:
With respect to persons, we are only interested in expert radiologists and medical students at German universities.
With respect to stimuli, we are only interested in the head CT scans made from subjects that do or do not suffer from intracerebral hemorrhage.
Similarly, we are only interested in AI-advice given by the specific AI-enabled diagnostic decision support system under study.

While a theoretical estimand is defined outside of any statistical model and be impossible to observe directly, an empirical estimand is a quantity that can be estimated based on observed data.
For our exemplary research question, it is possible to construct an experimental study, where all participants are confronted with the same set of brain scans, but the kind of AI-advice given for each brain scan is randomly assigned within participants.
This random intervention allows to estimate our theoretical estimand, although in reality, each person receives only one kind of AI-advice (correct advice, wrong advice, no advice) for each brain scan.

For our example, each probability expression in our theoretical estimand can be modeled with a GLMM.
Estimating the GLMM based on the data observed in our planned study produces an estimate for each probability and these estimates can be used to compute an estimate for each of the four probability contrasts.

## Simulate the data generating process

### Data simulation

#### THEORY

When the estimand has been defined, the next step in the research process is to write code that simulates the (assumed) data generating process of the planned study.
This requires specifying a generative process for all predictor variables used in the final data analysis.
While such assumptions can be quite challenging for observational studies and/or continuous predictor variables, this is less of a problem for experimental studies with only categorical predictor variables.
When all predictor variables have been simulated, one can use the structure of a suitable GLMM to simulate the dependent variable.
To simulate the GLMM, one requires plausible values for all model parameters.
We will discuss strategies for how these values can be obtained later.
Because one has full control over the data generating process in a tailored simulation, it is simple to also model highly specific aspects of the planned study, like data missing by design or assuming that subjects drop out (i.e. deleting their data from the simulated dataset) based on other variables in the simulated dataset.
The quality of the results of the sample size planning crucially depends on the plausibility of the simulated data generating process.
For this reason, deciding on a specific set of assumptions can easily feel overwelming.
However, we would argue that even a strongly simplified data generating process (e.g. including only a small number of interaction effect; including only random intercepts but no random slopes; making rather implausible assumptions for missing data) can yield informative result and should be strongly preferred to the current status quo in psychological research, where most studies do not justify their sample size or use highly questionable heuristics that are independent of their specific research question.

#### PRACTICE

In our case study, we simulate data for an experiment where the diagnostic performance of users of an AI-enabled diagnostic decision support system will be evaluated. 
<!-- The goal is to understand how AI advice influences medical decision-making. -->
Participants, radiologists (task experts; 1/4 of participants) and students/interns (non-task experts; 3/4), review a series of head computer tomography (CT) scans to assess the presence of a bleeding. 
To support their decision-making, an AI model provides initial diagnostic advice (2/3 of cases), which can be used as guidance by the participants.
In the control condition (1/3), no AI advice is presented, meaning that the participants have to read the CT scan without any support.
When AI advice is given, this advice can be either correct (4/5 of cases) or incorrect (1/5).
The type of advice (no advice, wrong advice, correct advice) is randomized within subjects across brain scans.
After reviewing a CT scan, participants deliver a medical diagnosis (bleeding or no bleeding), which may be either accurate or inaccurate.
This experimental design introduces some missing values by design since the advice is neither correct nor incorrect when no advice is present, which must be taken into account when simulating and analyzing the data.
<!-- With this experiment, we want to determine if (a) experts are better than non-experts in reading head CT scans and if (b) correct AI advice leads to better diagnostic accuracy than incorrect AI advice. -->
In this example, recruiting task experts (i.e., radiologists) is more challenging due to their limited availability, while non-experts (i.e., students/interns) are more readily accessible. 
The goal of simulation-based sample size planning is to determine how many task experts and non-experts must be recruited to achieve sufficient statistical power or precision in the planned experiment.

##### Our specific GLMM

In a GLMM, the expected value of the dependent variable Y conditioned on the vector of predictor variables $\mathbf{X}$ and random effects $\mathbf{U}$, transformed by a link function $g()$ is modeled as a linear combination $\eta$ of the predictor variables $\mathbf{X}$, the random effects $\mathbf{U}$ and the model parameters $\mathbf{\beta}$ [@fahrmeirRegressionModelsMethods2021]: 
$$
g(E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = \eta
$$
Equivalently, the conditional expected value is modeled as the linear combination $\eta$, transformed by the inverse link function $g^{-1}()$: 
$$
E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u}) = g^{-1}(\eta)
$$
If the dependent variable (i.e., diagnostic decision) $Y$ is a binary variable with values $0$ (i.e., inaccurate), or $1$ (i.e., accurate), the conditional expected value is equivalent to the probability: 
$$
P_{si} := P(Y = 1|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})
$$
In our case study, $P_{si}$ is the conditional probability that a subject $s$ gives the correct response to item (i.e., CT scan) $i$.

In such a setting, we model this probability as
$$
P_{si} = \text{inverse_logit}(\eta_{si})
$$
with the inverse-logit link $g^{-1}(\eta_{si}) = inverse\_logit(\eta_{si}) = \frac{exp(\eta_{si})}{1 + exp(\eta_{si})}$ or equivalently
$$
\text{logit}(P_{si}) = \eta_{si}
$$
with the logit link $g(P_{si}) = \text{logit}(P_{si}) = \text{ln} (\frac{P_{si}}{1 - P_{si}})$.

In our case study, the probability of making an accurate diagnostic decision is assumed to depend on the predictors:

- $advice\_present_{si}$: whether subject $s$ was presented with AI advice (1) or not (0) when asked to assess item $i$ 
- $advice\_correct_{si}$: whether this advice was correct (1) or not (0) 
- $expert_s$: whether subject $s$ was a task expert (1) or not (0)

and the random effects:

-   $u_{0s}$: the deviation of subject $s$ from the average ability to solve an item (i.e., CT scan) with average difficulty; assumed to be distributed as $u_{0s} \sim N(0, \sigma_S^2)$
-   $u_{0i}$: the deviation of item (i.e., CT scan) $i$ from the average difficulty to be solved by a person with average ability; assumed to be distributed as $u_{0i} \sim N(0, \sigma_I^2)$

In total, we assume the model
$$
\begin{aligned}
\text{logit}[P_{si}] =\ (&\beta_0 + u_{0s} + u_{0i}) + \\
&\beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
&\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}
\end{aligned}
$$
or equivalently
$$
\begin{aligned}
P_{si} = \text{inverse_logit}[&(\beta_0 + u_{0s} + u_{0i}) + \\
&\beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
&\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}]
\end{aligned}
$$
with model parameters $\beta_0$, $\beta_e$, $\beta_a$, $\beta_c$, $\beta_{ea}$, $\beta_{ec}$, $\sigma_S$, and $\sigma_I$.

In the GLMM literature, this would be called a binomial GLMM with two random intercepts (for subjects and items), two level-1 predictors ($advice\_present$, $advice\_correct$), one level-2 predictor ($expert$) and two cross-level interactions ($expert \cdot advice\_present$, $expert \cdot advice\_correct$).
To limit complexity, we do not consider random slopes, additional predictors or higher-level interactions.

##### Simulation function in R

The following R function simulates a full dataset structured according to the design of our case study.
The faux package [@R-faux] contains useful functions when simulating factorial designs, including random effects.

```{r}
simulate <- function(n_subjects = 100, n_items = 50,
  b_0 = 0.847, b_e = 1.350, b_a = -1.253, b_c = 2.603,
  b_ea = 0.790, b_ec = -1.393,
  sd_u0s = 0.5, sd_u0i = 0.5, ...){
  require(dplyr)
  require(faux)
  # simulate design
  dat <- add_random(subject = n_subjects, item = n_items) %>%
    add_between("subject", expert = c(1, 0), .prob = c(0.25, 0.75)) %>%
    mutate(advice_present = rbinom(n(), 1, prob = 2/3)) %>%
    mutate(advice_correct = if_else(advice_present == 1L, 
                                    rbinom(n(), 1L, prob = 0.8), 0L)) %>%
    # add random effects
    add_ranef("subject", u0s = sd_u0s) %>%
    add_ranef("item", u0i = sd_u0i) %>%
    # compute dependent variable
    mutate(linpred = b_0 + u0i + u0s +
        b_e * expert + b_a * advice_present + b_c * advice_correct +
        b_ea * expert * advice_present + b_ec * expert * advice_correct) %>%
    mutate(y_prob = plogis(linpred)) %>%
    mutate(y_bin = rbinom(n = n(), size = 1, prob = y_prob))
  dat
}
```

In the first six lines of the function definition, we set some default parameter values (which we will explain in the next section) and load the packages we use to manipulate and simulate data. 
In our case study, each subject (`n_subjects` in total) is assumed to respond to each item (i.e., CT scan; `n_items` in total). Thus, the `add_random` command creates a fully-crossed `data.frame` with `n_subjects` $\times$ `n_items` rows.
We add a between-subject effect with the `add_between` command, simulating that about $25\%$ of subjects are experts.
The next two lines simulate that in $\frac{2}{3}$ of trials, subjects will be presented with AI advice, and if advice is presented, the advice will be correct in about $80\%$ of cases (the variable `advice_correct` is always 0 when no advice is presented). Next, we simulate one random effect for each subject (`u0s`) and for each item (`u0i`).
As assumed by standard GLMMs, the `add_ranef` function draws the random effects from a normal distribution with a mean 0 and a standard deviation specified by the user.
With all design variables done, we are ready to simulate our model equation outlined in the last section.
The linear predictor variable `linpred` ($\eta$ in the GLMM model equations) combines the predictor variables, random effects, and model parameters as assumed by our model.
We then transform the linear predictor with the inverse-link function to compute `y_prob`, the probability that the subject correctly solved the item (in R, the inverse-logit link is computed with `plogis` and the logit link with `qlogis`).
In the final step, we simulate the binary dependent variable `y_bin` (i.e., whether the subject makes an accurate diagnostic decision for the CT scan) by -- for each trial -- drawing from a Bernoulli distribution with success probability `y_prob`.

### Strategies on specifying the population parameters

As discussed earlier, tailored simulation based sample size planning is especially relevant when studies have to be planned in the absence of previous studies with a similar design or any pilot data.
In this scenario, researchers require strategies on how to specify the population parameters used in their data simulation.
Population parameters are all model parameters estimated in a GLMM, in particular the regression coefficients of the fixed effects and the standard deviation of the random effects (and the correlation between random effects in more complicated models).
In contrast to non-hierarchical linear regression, common heuristics based on standardized effect sizes are less useful or not even available.
Our strategies to specify population parameters will require access to domain knowledge from domain experts.
Because most domain knowledge of domain experts can only be expressed in unstandardized measurement units of a specific application, we argue that unstandardized effect sizes are usually preferred over standardized effect sizes for tailored simulation based sample size planning.
This requires new strategies on how the data generating process implied by certain values of population parameters can be quantified or visualized in an intuitive way that enables a calibration of population parameters based on the available domain knowledge of domain experts.

Although we focus on frequentist statistics in our tutorial, selecting plausible values for data simulation shares many similarities with the specification of (weakly) informative prior distribution in applied Bayesian statistics.
In this literature, there have been developed some strategies to monitor the plausibility of (prior) model assumptions, that can be adapted for our purposes.

##### Insightful descriptive statistics

For most GLMMs, the interpretation of model parameters if quite complicated which makes it difficult to specify plausible values for a specific application.
However, the mathematical structure of GLMMs determines which patterns in data would be produced by the model, if a specific set of values for the population parameters would be specified.
The knowledge of how to simulate from a GLMM enables us to compute insightful descriptive statistics that can be compared to available domain knowledge much more easily than the opaque values of model parameters.
For example, domain experts might not be able to specify plausible values for the $\beta$ coefficients in a logistic regression model (which are measured on the log-odds scale).
However, they should be able to reason about the expected ratio of the binary dependent variable in different experimental condition, i.e. which descriptive statistics they expect to observe (on average), when they will analyse the actual dataset produced by the study under planning (if everything turns out as postulated by the investigated theory). 
The job of the analyst who is familiar with the mathematical structure of the GLMM is to produce the model implied value of the insightful descriptive statistic expected by the domain expert.
Although insightful descriptive statistics often depend on the model parameters in complicated ways, it is usually possible to manually adjust the population parameters (if only with the help of trial and error) until the model implied quantities produce the desired result.

##### Insightful model based quantities

Because GLMMs are complicated models, descriptive statistics alone are usually not enough to specify plausible values for all model parameters.
This is especially true for the standard deviation of random effects, that have complicated (and often unexpected) effects on the model-implied results.
An important advantage of data simulation (where one has full control over parameter values and sample sizes) is that one can produce insightful model based quantities that can never be directly observed in an actual empirical dataset.
For example, in a logistic model with random intercepts for participants, one can produce a visualization of the implied distribution of the probability that a participant on average solves a cognitive task.
Although domain knowledge will probably not suffice to specify this distribution completely, it should be possible to rule out unplausible boundary conditions.
For example, the domain expert might deem it unplausible that the 5% most able participants have a probability of more than 0.99 to solve the difficult cognitive task.

#### Iterative process with domain experts

In our experience, the gathering of domain knowledge by domain experts and the consecutive specification of population parameter values used in data simulation is not a one-time event but rather an iterative process.
In a first step, domain experts can be interviewed to "elicit" their domain knowledge about how the future data of the planned study is expected to look like.
As most domain experts are no experts in statistical modeling and GLMMs, they often struggle without further guidance to communicate their knowledge in a way that is useful when specifying the parameters for data simulation.
For this reason, we suggest that after an initial unstructured interview of domain experts, the analyst which is familiar with the structure of the GLMM under study selects an initial set of insightful descriptive statistics and model based quantities.
Then they reenter into an iterative discussion where some set of population values are selected and the plausibility of resulting implied quantities are discussed with the domain experts.
Then, the population parameters are updated based on this discussion until the domain experts are satisfied with the final result.
During this process, the monitored model based quantities and descriptive statistics can be updated or extended to capture as much available domain knowledge as possible.

##### PRACTICE

All parameter values in our present case study have been determined from results from related prior work and repeated discussions with our affiliated domain experts in Radiology to validate our assumptions.
Initially, we reviewed the literature to establish a reasonable baseline performance rate for examining head CT scans for ICH. Existing studies indicate that radiologists typically demonstrate high accuracies, often exceeding or hovering around 90%, while interns have been shown to perform below 80%, and medical students fall even shorter.
For simplicity, we assumed plausible probability values of .90 for experts and .70 for non-experts, respectively.
Our experts confirmed that these values are realistic baselines for reviewing diverse head CT images without AI assistance. Subsequently, we consulted several published papers investigating the effect of correct and incorrect advice on decision-making performance in other settings.
From their findings, we inferred that both experts and non-experts would benefit from correct and suffer losses from incorrect advice.
However, the magnitude of these effects would be substantially greater for non-experts, given their demonstrated reliance on advice compared to experts.
We further validated the plausibility of our estimated gains and losses with the collaborating radiologists.
For our simulation, we used the probabilities of average participants to solve an average case, as shown in Table 1.

## Estimate the statistical model

### THEORY

At this point, the researcher is capable of producing a simulated dataset similar to the actual dataset that will later be collected in the planned study.
The next step is to specify how the statistical model shall be estimated in the actual study collected later.
This usually includes the selection of 1) a statistical framework (e.g., frequentist statistics), 2) a software package that is capable of estimating the model class of interest (e.g., the lme4 R package), 3) an estimation algorithm (e.g., the default optimizer "bobyqa"), and 4) the specific model structure including all fixed effects, random effects, and the model family of the dependent variable.   
Note that this does not always mean that one will specify the same GLMM that was used when specifying the data generating process.
On the one hand, using a simpler model for data simulation than for model estimation can be a useful strategy in scenarios where making plausible assumptions for complicated random effect structures and interactions is not feasible.
On the other hand, using a more complex model (that perhaps even goes beyond a standard GLMM) for data simulation than for model estimation can be a useful strategy in scenarios where one has specific domain knowledge about aspects of the data generating process that are still difficult to estimate with the current state-of-the-art in multilevel modeling.
In such more complicated scenarios, the final model specification might also depend on the next decision on specifying the theoretical estimand.

### PRACTICE

In our case study, we use the lme4 R package [@batesFittingLinearMixedEffects2015], which is a state-of-the-art tool for fitting frequentist GLMMs.[^1]
The lme4 package includes a function called `simulate` that allows researchers to simulate the dependent variable based on the same model formula used for model fitting, enabling simulation-based power analyses and other related analyses.

[^1]: For Bayesian GLMMs, the brms R package is currently the most prominent option [@burknerBrmsPackageBayesian2017].

However, the model parameterization used by the lme4 package is quite technical, making it difficult for applied researchers to determine whether their specified population model (i.e., the theoretical model that describes the underlying data generation process for a specific population of interest) implies plausible associations in their simulated data. Therefore, in this tutorial, we simulate data for GLMMs from first principles (i.e., creating synthetic data step by step instead of using black box functions) to assist applied researchers in better understanding all model assumptions and then use lme4 to analyze the simulated datasets.[^2]

[^2]: A less flexible alternative would be to use the simr package [@greenSIMRPackagePower2016], which can be used to both simulate data and perform power analysis for models supported by the lme4 package.

In this section, we show how to fit a GLMM with lme4, interpret the model, and test hypotheses derived from a research question.
We simulate data according to our model, in which 100 subjects respond to 50 items (we use `set.seed` to make the simulation reproducible). However, for the sake of the exercise, we can imagine that this would be real data resulting from our future experiment and think about how we would analyze this data.

```{r, message=FALSE}
library(tidyverse)
set.seed(1)
dat <- simulate(n_subjects = 100, n_items = 50)
```

The lme4 package uses a special syntax for model specification. Our specific GLMM is represented by the formula:

```{r, message=FALSE}
library(lme4)
f <- y_bin ~ 1 + expert + advice_present + advice_correct + 
  expert:advice_present + expert:advice_correct +
  (1|subject) + (1|item)
```

The first two lines look similar to any linear model in R (general intercept indicated by `1`; main effects indicated by variable names in the dataset; interactions indicated by `variable1:variable2`). The third line specifies a random intercept for each subject `(1|subject)` and for each item `(1|item)`. The complete set of rules for the syntax is outlined in @batesFittingLinearMixedEffects2015 and in the documentation of the lme4 package.

In lme4, a GLMM is fitted with the `glmer` function. By setting `family =  "binomial"`, we request a binomial GLMM appropriate for our binary dependent variable `y_bin` (the binomial GLMM uses the canonical logit link by default), which is defined as an accurate (1) vs. inaccurate (0) diagnosis.

```{r}
fit <- glmer(f, data = dat, family = "binomial")
```

## Compute the estimate

### THEORY

When the theoretical and empirical estimands have been identified in the modeling framework used to estimate the GLMM, one has to decide on whether hypothesis testing or interval estimation should be used to answer the research question.
This decision is then followed by selecting the specific statistical method that shall be applied to compute the hypothesis test(s) or confidence interval(s) (e.g., compute hypothesis tests and confidence intervals with the marginaleffects R package using the delta method).

### PRACTICE

For our case study, we investigate the following research question:
*We expect that for both experts and non-experts, correct advice leads to a higher probability of accurately diagnosing a CT scan compared to no advice presented, AND, we expect that for both experts and non-experts, incorrect advice leads to a lower probability of accurately diagnosing a CT scan compared to no advice presented.*

In the earlier section on our theoretical estimand we translated this verbal description into four probability statements that are specified outside of any specific statistical model.
The next translation step is to specify how these probabilites can be estimated within the framework of our concrete GLMM.

The first of the four statements was:

$$
\begin{aligned}
& P(\text{correct diagnosis} | \text{correct advice, an average expert, an average scan}) \\
& \quad - P(\text{correct diagnosis} | \text{no advice, an average expert, an average scan})
\end{aligned}
$$
In the context of our GLMM, the corresponding model based quantity is:

$$
\begin{aligned}
& P(Y=1|advice\_present = 1, advice\_correct = 1, expert = 1, u_{0s} = 0, u_{0i} = 0) \\
& \quad - P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 1, u_{0s} = 0, u_{0i} = 0)
\end{aligned}
$$
Plugging in the model equation of the GLMM produces an equation on how to compute this contrast when all model parameters are known.

<!-- $$ -->
<!-- \begin{aligned} -->
<!-- & \text{inverse_logit}[\beta_0 + \beta_a \cdot 1 + \beta_c \cdot 1 + \beta_e \cdot 1 + \beta_{ea} \cdot 1 \cdot 1 + \beta_{ec} \cdot 1 \cdot 1]\\ -->
<!-- & \quad - \text{inverse_logit}[\beta_0 + \beta_a \cdot 0 + \beta_c \cdot 0 + \beta_e \cdot 1 + \beta_{ea} \cdot 1 \cdot 0 + \beta_{ec} \cdot 1 \cdot 0] = \\ -->
<!-- & = \text{inverse_logit}[\beta_0 + \beta_a  + \beta_c + \beta_e + \beta_{ea} + \beta_{ec} ]\\ -->
<!-- & \quad - \text{inverse_logit}[\beta_0 + \beta_e ] \\ -->
<!-- \end{aligned} -->
<!-- $$ -->

When model parameters are not known, the corresponding parameter estimates are plugged into the equation to compute an estimate of the contrast instead.

In general, we can inspect the estimates for all model parameters with the `summary` command:

```{r}
summary(fit)
```

In the lme4 output, the `Estimate` column in the `Fixed effects` table contains the estimates for the $\beta$ parameters, while the `Std.Dev.` column in the `Random effects` table contains the estimates for $\sigma_S$ and $\sigma_I$.

Unfortunately, there is no single model parameter that corresponds with our contrast of interest.
the model parameters in a binomial GLMM are hard to interpret because 1) the $\beta$ parameters are connected to the modeled probability via the non-linear inverse-logit link, and 2) we also have to consider the random effects.
When we are interested in a subject with average ability ($u_{0s} = 0$) responding to an item (i.e., CT scan) with average difficulty ($u_{0i} = 0$), the model implied probability that such a person solves such an item accurately is given by: 
$$
\begin{aligned}
P(Y=1|\mathbf{X=x}, \mathbf{U} = \mathbf{0}) = \\
= \text{inverse_logit}[&\beta_0 + \beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
&\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}]
\end{aligned}
$$
In fact, we would only need the full equation if the subject is an expert and correct advice is presented.
In all other experimental conditions, some terms drop from the equation because they are multiplied by $0$.
The other extreme case would be the probability that a non-expert with average ability solves an item with average difficulty when no advice is presented:
$$
\begin{aligned}
P(Y=1| advice\_present = 0, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0) = \\
= \text{inverse_logit}[\beta_0]
\end{aligned}
$$
One could use such reasoning, to determine the exact formula needed to compute our contrast of interest.
However, this would be very tedious and we can use R to compute this contrast for us without doing the math.

We can use the predict function of the lme4 package to compute the predicted probability for a correct diagnosis based on our model, plugin the two sets of predictor values, and compute the difference between the two estimated probabilities.

```{r}
pred_df <- data.frame(advice_present = c(1, 0), advice_correct = c(1, 0), expert = c(1, 1))
pred_df
preds <- predict(fit, newdata = pred_df, type = "response", re.form = NA)
preds
preds[1] - preds[2]
```

The argument `type = "response"` specifies that predictions are made on the probability scale, while `re.form = NA` sets all random effects to 0.

We could use this method to compute point estimates for all four contrast that are part of our estimand.
However, depending on whether we are interested in hypothesis testing or parameter estimation, we also need a method to compute hypothesis tests or confidence interval for our point estimates.

The marginaleffects package [@R-marginaleffects] is a very flexible package to compute hypothesis tests and confidence intervals for contrasts with a variety of statistical models, including GLMM estimated with lme4.

First, we compute estimated probabilities for all experimental conditions in our experiment with the `predictions` function:

```{r}
pred_df <- expand_grid(advice_present = 0:1, advice_correct = 0:1, expert = 0:1)
probs <- predictions(fit, newdata = pred_df, type = "response", re.form = NA)
probs
```

Note that the result also contains the two missing by design conditions that are never observed in the actual study ($advice\_present = 0, advice\_correct = 1, expert = 1$ and $advice\_present = 0, advice\_correct = 1, expert = 0$), and we should never interpret those.

Next, we use these predictions to compute the four specific contrasts that are part of our estimand.
For this we must specify which rows in `probs` have to be subtracted from each other.
We will use the `hypotheses` function to compute our four contrasts of interest together with hypothesis tests and confidence intervals.
We use the default inference options of the marginaleffects package that compute hypothesis tests and confidence intervals based on the approximate delta method.

```{r}
probs %>% 
  hypotheses(hypothesis = c(
    "b8 = b2",
    "b2 = b6",
    "b7 = b1",
    "b1 = b5"),
    equivalence = c(0, 0))
```

The expression `"b8 = b2"` is special syntax to subtract the estimate in row number 8 from the estimate in row number 2.
The argument `equivalence = c(0, 0)` can be used to compute one-sided p-values testing whether the constrast in the population is smaller than 0 (`p (NonSub)` column) or greater than zero (`p (NonInf)` column).
We computed the contrasts in a way that we expect positive values for all four of them.

#### Hypothesis testing

If we are interested in hypothesis testing in our case study, we test a combined null hypothesis that consists of four separate null hypotheses:

$$
\begin{aligned}
H_{01}:\ & P(Y=1|advice\_present = 1, advice\_correct = 1, expert = 1, u_{0s} = 0, u_{0i} = 0) \leq \\
& P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 1, u_{0s} = 0, u_{0i} = 0) \\
H_{02}:\ &P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 1, u_{0s} = 0, u_{0i} = 0) \leq \\
& P(Y=1|advice\_present = 1, advice\_correct = 0, expert = 1, u_{0s} = 0, u_{0i} = 0) \\
H_{03}:\ &P(Y=1|advice\_present = 1, advice\_correct = 1, expert = 0, u_{0s} = 0, u_{0i} = 0) \leq \\
& P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0) \\
H_{04}:\ & P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0) \leq \\
& P(Y=1|advice\_present = 1, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0)
\end{aligned}
$$
The combined null hypothesis should only be rejected if **all** individual null hypotheses are rejected [i.e., intersection-union setting; @dmitrienkoTraditionalMultiplicityAdjustment2013].
In such cases, the error probabilities do not accumulate, and we would waste power when correcting for multiple tests.

With a standard significance level of $\alpha = 0.05$, we would not reject all four null hypotheses (the p-value in the `p (NonInf)` column for the third hypothesis is not significant) and therefore also not reject the combined null hypothesis for this simulated dataset.
Note that this decision would be wrong because we have simulated the data such that the combined alternative hypothesis is actually true in the population.

#### Interval estimation

If we are interested in parameter estimation in our case study, we focus on the two-sided confidence intervals of the four contrasts of interest.
With a standard confidence level of $1 - \alpha = 0.95$, plausible values are clearly in the positive range for the first, third and fourth contrast, while both negative and positive values seem plausible for the second contrast.
Note that due to the constrained range of the probability scale, the width of the confidence interval differs between the four contrasts.
The smallest width is observed for the first contrast where both underlying probabilities are close to 1.
The largest width is observed for the fourth contrast, where both underlying probabilities are closer to 0.5.

<!-- Due to this complicated relationship, we argue not to focus too much on interpreting single model parameters when working with GLMMs. Instead, it can be more intuitive to consider model predictions and the model-implied distribution of the dependent variable for each experimental condition across all subjects and items. -->

<!-- With the marginaleffects package [@R-marginaleffects], we can easily compute predictions for all observations in the dataset based on the fitted GLMM (including all fixed **and** random effects), and plot the average probability with confidence intervals for each experimental condition in Figure \@ref(fig:marginaleffects): -->

<!-- (ref:marginaleffects) Marginal distributions including means and 95% confidence intervals for all experimental conditions computed with the marginaleffects package. -->

<!-- ```{r marginaleffects, message=FALSE, fig.cap="(ref:marginaleffects)"} -->
<!-- library(marginaleffects) -->
<!-- plot_predictions(fit, by = c("advice_present", "advice_correct", "expert"), -->
<!--   type = "response") + ylim(c(0.3, 1)) -->
<!-- ``` -->

## Perform repeated simulations

### THEORY

Conducting the previous steps enable the analyst to 1) simulate a dataset, 2) estimate a GLMM, and 3) compute hypothesis tests or confidence intervals for estimands of interest, mirroring the analysis that will later be performed for the actual dataset of the planned study.
The last missing piece is to write code to perform the above steps repeatedly and allow for a setting using different sample sizes.
On a conceptual level, we first require a function, that takes as input the sample size and the full set of population parameter.
When planning for power, the function should return the p-value(s) of the hypothesis test(s) of interest when conducted on the simulated dataset.
When planning for precision, the function should instead return the width of the confidence interval(s) of interest.
Secondly, we must run this function repeatedly with the same sample and population parameters.
Because even fitting GLMMs with frequentist methods can quickly become time-consuming, it is recommended to use parallel computations, that is running simulations on multiple cores of the computer at the same time to reduce total computation time.
Thirdly, the results of the repeated simulation must be collected and aggregated.
When planning for power, we compute the relative frequency of a significant p-value across repeated simulations.
When planning for precision, we compute the average width of the confidence interval.
Lastly, we have to repeat the complete simulation for different sample sizes, to determine how big the sample must be in order to achieve the targeted statistical power or the expected width of the confidence interval.

### PRACTICE

With all these considerations addressed, we are finally ready to perform a power analysis. Wrapping the `simulate` function already constructed earlier, the helper function `sim_and_analyse` performs all previous steps (simulate a dataset, fit a GLMM, compute p-values) in a single command.

```{r}
sim_and_analyse <- function(
  formula_chr = "y_bin ~ 1 + expert + advice_present + advice_correct + 
    expert:advice_present + expert:advice_correct + (1|subject) + (1|item)", 
  null_hypotheses = c("advice_present + advice_correct + 
    expert:advice_present + expert:advice_correct <= 0",
    "advice_present + advice_correct <= 0",
    "-1 * (advice_present + expert:advice_present) <= 0",
    "-1 * (advice_present) <= 0"), ...){
  require(lme4)
  require(multcomp)
  # simulate data
  dat <- simulate(...)
  # fit model
  model <- glmer(as.formula(formula_chr), data = dat, family = "binomial")
  # compute p-values
  glht <- glht(model, linfct = null_hypotheses)
  pvalues <- summary(glht, test = univariate())$test$pvalues
  setNames(pvalues, paste0("p_H0", 1:length(null_hypotheses)))
}
```

Power analysis can quickly become computationally intensive when we repeatedly simulate data and fit models for different parameter combinations or sample sizes. Thus, we use the future [@R-RJ-2021-048] and furrr [@R-furrr] packages to perform computations in parallel. First, we enable parallelization and specify how many parallel cores ("workers") of our computer to use (users can find out the maximum number of cores on their computer with the command `parallel::detectCores()`), and set a seed to make the simulation reproducible.

```{r, message=FALSE, eval=FALSE}
library(future)
plan("multisession", workers = 6)
set.seed(2)
```

The next code chunk specifies a simulation grid with different settings for both the number of subjects (`n_subjects`) and the number of items (`n_items`), each combination being repeated `rep` times. We chose 300 repetitions for the data simulation at hand as it strikes a balance between achieving a robust statistical estimate and remaining computationally feasible.
With the current settings, this simulation takes about one hour on a MacBook Pro from 2020 with M1 chip and 16 GB working memory. If you want to quickly experiment with the code yourself, a setting with `workers = 4` and `rep = 5` should finish in less than 5 minutes, even on smaller machines.

```{r, message=FALSE, echo=FALSE, eval=FALSE}
library(future)
library(furrr)
plan("multisession", workers = 6)
set.seed(2)
sim_design <- crossing(
  rep = 1:300,
  n_subjects = c(100, 150, 200, 250),
  n_items = c(10, 30, 50, 70)
) %>%
  mutate(pvalues = future_pmap(., sim_and_analyse, 
    .options = furrr_options(seed = TRUE))) %>%
  unnest_wider(pvalues)
saveRDS(sim_design, file = "results.rds")
```

```{r, message=FALSE, eval=FALSE}
library(furrr)
sim_design <- crossing(
  rep = 1:300,
  n_subjects = c(100, 150, 200, 250),
  n_items = c(10, 30, 50, 70)
) %>%
  mutate(pvalues = future_pmap(., sim_and_analyse, 
    .options = furrr_options(seed = TRUE))) %>%
  unnest_wider(pvalues)
```

The result of the computation is a data frame that contains the p-values of all tested hypotheses for each simulated dataset.
In some iterations (predominantly in conditions with small sample sizes), model estimation did not converge with the lme4 package. When the model fails to converge, it means that the statistical model being fitted to the data failed to reach a stable or valid solution during the estimation process. We do not remove these results because non-convergence can also happen when analyzing the real data we plan to collect, thus, we want to factor in this possibility to keep our simulation more realistic.

For our exemplary combined hypothesis, power is defined as the (long-run) percentage of simulations in which all four p-values of our component hypotheses are significant at the $\alpha = 0.05$ level. Based on our simulation outcomes, we compute a power estimate for each combination of `n_subjects` $\times$ `n_items` (including 95% confidence intervals) and visualize the results with the following code.[^3]

[^3]: This code was inspired by the "Mixed Design Simulation" vignette of the faux package at <https://debruine.github.io/faux/articles/sim_mixed.html>.

```{r, echo=FALSE}
sim_design <- readRDS(file = "results.rds")
```

```{r finalpwr, fig.cap="Simulation-based power estimates including 95% confidence interval of the case study for different numbers of subjects and items, based on a significance level of 0.05."}
library(binom)
alpha <- 0.05
power <- sim_design %>%
  group_by(n_subjects, n_items) %>% 
  summarise(power = mean(p_H01 < alpha & p_H02 < alpha & 
                           p_H03 < alpha & p_H04 < alpha), 
    n_sig = sum(p_H01 < alpha & p_H02 < alpha & 
                  p_H03 < alpha & p_H04 < alpha),
    n = n(),
    ci.lwr = binom.confint(n_sig, n, method = "wilson")$lower,
    ci.upr = binom.confint(n_sig, n, method = "wilson")$upper, 
    .groups = "drop")
power %>%
  mutate(across(c(n_subjects, n_items), factor)) %>%
  ggplot(aes(n_subjects, n_items, fill = power)) +
  geom_tile() +
  geom_text(aes(label = sprintf("%.2f \n [%.2f; %.2f]", 
                                power, ci.lwr, ci.upr)), 
    color = "white", size = 4) +
  scale_fill_viridis_c(limits = c(0, 1)) +
  xlab("number of subjects") + ylab("number of items")
```

<!-- How many participants/stimuli are needed for concrete example -->

As should be the case, power estimates in Figure \@ref(fig:finalpwr) increase with both the number of subjects and the number of items. The confidence intervals indicate how precisely power was estimated by our simulation. Higher precision (which would be reflected in narrower confidence intervals) could be obtained by increasing the number of repetitions (`rep`) in the simulation. In practice, data simulations are often run multiple times with adjusted combinations of sample sizes. When running for the first time, it might be revealed that power is way too low (or much higher than required) for some combinations of `n_subjects` and `n_items`. When narrowing down the best combination that achieves sufficient power while at the same time striking a good balance of how many subjects and items are practically feasible, later rounds of data simulation will typically include a smaller grid of sample sizes combined with a higher number of repetitions. This will assure high precision for the final power estimates, which are then used for the sample size justification of the future study.

Much has been written on the optimal amount of power to target in empirical research. The most prominent heuristic is to target a power of 0.8 (when combined with a type I error rate of $\alpha = 0.05$), but depending on the research goals of the study, there are often good reasons to move away from this standard depending on the research goals and resource constraints  [@lakensJustifyYourAlpha2018; @lakensSampleSizeJustification2022]. When target power has been specified, the number of subjects and the number of items in our study design can be traded against each other based on practical considerations. For the sake of the example, let the targeted power be indeed about 0.8, using an $\alpha$ of 0.05 to detect an effect of the expected size implied by our data simulation. This could be achieved by collecting data from 200 subjects (about 25% of which will be experts), each completing the same 50 items (with advice present in about 67% of cases, which is correct in about 80% of cases with present advice). If collecting data from 200 subjects is not feasible, an alternative would be to recruit 150 subjects but increase the length of the experiment to over 70 items. However, 70 items might take too long to complete for the radiologists participating in the study, who have a busy schedule. The simulation suggests that it might also be possible to plan a shorter experiment with only 30 items if it is feasible to recruit an even higher number of subjects (\> 250, to be determined by additional rounds of power analysis). Design parameters that also affect power, and which could be investigated in the simulation to find a more optimal trade-off, are the ratio of experts, the frequency of whether advice is presented and whether it is correct.

<!-- ## Sensitivity analysis -->

# THIS WAS THE OLD CASE STUDY CHAPTER

<!-- overview of this section -->

In this section, we outline the steps for performing data simulation and a priori power analysis for GLMMs using a case study based on a specific experimental study design from the area of human-AI (artificial intelligence) interaction research. All code in this manuscript and simulation results are available in the project's repository on the Open Science Framework (https://osf.io/dhwf4/).

## Experimental study design

<!-- describe the study idea and design -->
In the present case study, we simulate data for an experiment where the diagnostic performance of users of an AI-enabled diagnostic decision support system will be evaluated. 
The goal is to understand how AI advice influences medical decision-making. Participants, radiologists (task experts) and students/interns (non-task experts), review head computer tomography (CT) scans to assess the presence of a bleeding. 
To support their decision-making, an AI model provides initial diagnostic advice, which can be used as guidance by the participants. This AI advice can be either correct (80% of cases) or incorrect (20%). In the control condition, no AI advice is presented, meaning that the participants have to read the CT scan without any support. After reviewing the CT scan, participants deliver a medical diagnosis (bleeding or no bleeding), which may be either accurate or inaccurate.
This experimental design introduces some missing values by design since the advice is neither correct nor incorrect when no advice is present, which must be taken into account when simulating and analyzing the data. With this experiment, we want to determine if (a) experts are better than non-experts in reading head CT scans and if (b) correct AI advice leads to better diagnostic accuracy than incorrect AI advice.
In this example, recruiting task experts (i.e., radiologists) is more challenging due to their limited availability, while non-experts (i.e., students/interns) are more readily accessible. 
The goal of the present simulation-based power analysis is to determine how many task experts and non-experts must be recruited to achieve sufficient statistical power in the planned experiment.

## Our specific GLMM

<!-- deep dive into our specific example -->
In a GLMM, the expected value of the dependent variable Y conditioned on the vector of predictor variables $\mathbf{X}$ and random effects $\mathbf{U}$, transformed by a link function $g()$ is modeled as a linear combination $\eta$ of the predictor variables $\mathbf{X}$, the random effects $\mathbf{U}$ and the model parameters $\mathbf{\beta}$ [@fahrmeirRegressionModelsMethods2021]: 
$$
g(E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = \eta
$$
Equivalently, the conditional expected value is modeled as the linear combination $\eta$, transformed by the inverse link function $g^{-1}()$: 
$$
E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u}) = g^{-1}(\eta)
$$
If the dependent variable (i.e., diagnostic decision) $Y$ is a binary variable with values $0$ (i.e., inaccurate), or $1$ (i.e., accurate), the conditional expected value is equivalent to the probability: 
$$
P_{si} := P(Y = 1|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})
$$
In our case study, $P_{si}$ is the conditional probability that a subject $s$ gives the correct response to item (i.e., CT scan) $i$.

In such a setting, we model this probability as
$$
P_{si} = inverse\_logit(\eta_{si})
$$
with the inverse-logit link $g^{-1}(\eta_{si}) = inverse\_logit(\eta_{si}) = \frac{exp(\eta_{si})}{1 + exp(\eta_{si})}$ or equivalently
$$
logit(P_{si}) = \eta_{si}
$$
with the logit link $g(P_{si}) = logit(P_{si}) = ln (\frac{P_{si}}{1 - P_{si}})$.

In our case study, the probability of making an accurate diagnostic decision is assumed to depend on the predictors:

- $advice\_present_{si}$: whether subject $s$ was presented with AI advice (1) or not (0) when asked to assess item $i$ 
- $advice\_correct_{si}$: whether this advice was correct (1) or not (0) 
- $expert_s$: whether subject $s$ was a task expert (1) or not (0)

and the random effects:

-   $u_{0s}$: the deviation of subject $s$ from the average ability to solve an item (i.e., CT scan) with average difficulty; assumed to be distributed as $u_{0s} \sim N(0, \sigma_S^2)$
-   $u_{0i}$: the deviation of item (i.e., CT scan) $i$ from the average difficulty to be solved by a person with average ability; assumed to be distributed as $u_{0i} \sim N(0, \sigma_I^2)$

In total, we assume the model
$$
\begin{aligned}
logit[P_{si}] =\ (&\beta_0 + u_{0s} + u_{0i}) + \\
&\beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
&\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}
\end{aligned}
$$
or equivalently
$$
\begin{aligned}
P_{si} = inverse\_logit[&(\beta_0 + u_{0s} + u_{0i}) + \\
&\beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
&\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}]
\end{aligned}
$$
with model parameters $\beta_0$, $\beta_e$, $\beta_a$, $\beta_c$, $\beta_{ea}$, $\beta_{ec}$, $\sigma_S$, and $\sigma_I$.

In the GLMM literature, this would be called a binomial GLMM with two random intercepts (for subjects and items), two level-1 predictors ($advice\_present$, $advice\_correct$), one level-2 predictor ($expert$) and two cross-level interactions ($expert \cdot advice\_present$, $expert \cdot advice\_correct$).
To limit complexity, we do not consider random slopes, additional predictors or higher-level interactions.

## Data simulation

The following R function simulates a full dataset structured according to the design of our case study. The faux package [@R-faux] contains useful functions when simulating factorial designs, including random effects.

```{r}
simulate <- function(n_subjects = 100, n_items = 50,
  b_0 = 0.847, b_e = 1.350, b_a = -1.253, b_c = 2.603,
  b_ea = 0.790, b_ec = -1.393,
  sd_u0s = 0.5, sd_u0i = 0.5, ...){
  require(dplyr)
  require(faux)
  # simulate design
  dat <- add_random(subject = n_subjects, item = n_items) %>%
    add_between("subject", expert = c(1, 0), .prob = c(0.25, 0.75)) %>%
    mutate(advice_present = rbinom(n(), 1, prob = 2/3)) %>%
    mutate(advice_correct = if_else(advice_present == 1L, 
                                    rbinom(n(), 1L, prob = 0.8), 0L)) %>%
    # add random effects
    add_ranef("subject", u0s = sd_u0s) %>%
    add_ranef("item", u0i = sd_u0i) %>%
    # compute dependent variable
    mutate(linpred = b_0 + u0i + u0s +
        b_e * expert + b_a * advice_present + b_c * advice_correct +
        b_ea * expert * advice_present + b_ec * expert * advice_correct) %>%
    mutate(y_prob = plogis(linpred)) %>%
    mutate(y_bin = rbinom(n = n(), size = 1, prob = y_prob))
  dat
}
```

In the first six lines of the function definition, we set some default parameter values (which we will explain in a later section) and load the packages we use to manipulate and simulate data. 
In our case study, each subject (`n_subjects` in total) is assumed to respond to each item (i.e., CT scan; `n_items` in total). Thus, the `add_random` command creates a fully-crossed `data.frame` with `n_subjects` $\times$ `n_items` rows. We add a between-subject effect with the `add_between` command, simulating that about $25\%$ of subjects are experts. The next two lines simulate that in $\frac{2}{3}$ of trials, subjects will be presented with AI advice, and if advice is presented, the advice will be correct in about $80\%$ of cases (the variable `advice_correct` is always 0 when no advice is presented). Next, we simulate one random effect for each subject (`u0s`) and for each item (`u0i`). As assumed by standard GLMMs, the `add_ranef` function draws the random effects from a normal distribution with a mean 0 and a standard deviation specified by the user. With all design variables done, we are ready to simulate our model equation outlined in the last section. The linear predictor variable `linpred` ($\eta$ in the GLMM model equations) combines the predictor variables, random effects, and model parameters as assumed by our model. We then transform the linear predictor with the inverse-link function to compute `y_prob`, the probability that the subject correctly solved the item (in R, the inverse-logit link is computed with `plogis` and the logit link with `qlogis`). In the final step, we simulate the binary dependent variable `y_bin` (i.e., whether the subject makes an accurate diagnostic decision for the CT scan) by -- for each trial -- drawing from a Bernoulli distribution with success probability `y_prob`.

## Model fitting

<!-- ## The lme4 package in R -->

<!-- describe the lme4 package and how it works-->
In our case study, we use the lme4 R package [@batesFittingLinearMixedEffects2015], which is a state-of-the-art tool for fitting frequentist GLMMs.[^1]
The lme4 package includes a function called `simulate` that allows researchers to simulate the dependent variable based on the same model formula used for model fitting, enabling simulation-based power analyses and other related analyses.

[^1]: For Bayesian GLMMs, the brms R package is currently the most prominent option [@burknerBrmsPackageBayesian2017].

However, the model parameterization used by the lme4 package is quite technical, making it difficult for applied researchers to determine whether their specified population model (i.e., the theoretical model that describes the underlying data generation process for a specific population of interest) implies plausible associations in their simulated data. Therefore, in this tutorial, we simulate data for GLMMs from first principles (i.e., creating synthetic data step by step instead of using black box functions) to assist applied researchers in better understanding all model assumptions and then use lme4 to analyze the simulated datasets.[^2]

[^2]: A less flexible alternative would be to use the simr package [@greenSIMRPackagePower2016], which can be used to both simulate data and perform power analysis for models supported by the lme4 package.

In this section, we show how to fit a GLMM with lme4, interpret the model, and test hypotheses derived from a research question.
We simulate data according to our model, in which 100 subjects respond to 50 items (we use `set.seed` to make the simulation reproducible). However, for the sake of the exercise, we can imagine that this would be real data resulting from our future experiment and think about how we would analyze this data.

```{r, message=FALSE}
library(tidyverse)
set.seed(1)
dat <- simulate(n_subjects = 100, n_items = 50)
```

The lme4 package uses a special syntax for model specification. Our specific GLMM is represented by the formula:

```{r, message=FALSE}
library(lme4)
f <- y_bin ~ 1 + expert + advice_present + advice_correct + 
  expert:advice_present + expert:advice_correct +
  (1|subject) + (1|item)
```

The first two lines look similar to any linear model in R (general intercept indicated by `1`; main effects indicated by variable names in the dataset; interactions indicated by `variable1:variable2`). The third line specifies a random intercept for each subject `(1|subject)` and for each item `(1|item)`. The complete set of rules for the syntax is outlined in @batesFittingLinearMixedEffects2015 and in the documentation of the lme4 package.

In lme4, a GLMM is fitted with the `glmer` function. By setting `family =  "binomial"`, we request a binomial GLMM appropriate for our binary dependent variable `y_bin` (the binomial GLMM uses the canonical logit link by default), which is defined as an accurate (1) vs. inaccurate (0) diagnosis.

```{r}
fit <- glmer(f, data = dat, family = "binomial")
```

## Model interpretation

We can inspect the estimates for all model parameters with the `summary` command:

```{r}
summary(fit)
```

In the output, the `Estimate` column in the `Fixed effects` table contains the estimates for the $\beta$ parameters, while the `Std.Dev.` column in the `Random effects` table contains the estimates for $\sigma_S$ and $\sigma_I$.

Unfortunately, the model parameters in a binomial GLMM are hard to interpret because 1) the $\beta$ parameters are connected to the modeled probability via the non-linear inverse-logit link, and 2) we also have to consider the random effects. The most simple interpretation works by imagining a subject with average ability ($u_{0s} = 0$) responding to an item (i.e., CT scan) with average difficulty ($u_{0i} = 0$). Then the model implied probability that such a person solves such an item accurately is given by: 
$$
\begin{aligned}
P(Y=1|\mathbf{X=x}, \mathbf{U} = \mathbf{0}) = \\
= inverse\_logit[&\beta_0 + \beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
&\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}]
\end{aligned}
$$

In fact, we would only need the full equation if the subject is an expert and correct advice is presented. In all other experimental conditions, some terms drop from the equation because they are multiplied by $0$. The other extreme case would be the probability that a non-expert with average ability solves an item with average difficulty when no advice is presented:
$$
\begin{aligned}
P(Y=1|expert = 0, advice\_present = 0, advice\_correct = 0, u_{0s} = 0, u_{0i} = 0) = \\
= inverse\_logit[\beta_0]
\end{aligned}
$$

Due to this complicated relationship, we argue not to focus too much on interpreting single model parameters when working with GLMMs. Instead, it can be more intuitive to consider model predictions and the model-implied distribution of the dependent variable for each experimental condition across all subjects and items.

With the marginaleffects package [@R-marginaleffects], we can easily compute predictions for all observations in the dataset based on the fitted GLMM (including all fixed **and** random effects), and plot the average probability with confidence intervals for each experimental condition in Figure \@ref(fig:marginaleffects):

(ref:marginaleffects) Marginal distributions including means and 95% confidence intervals for all experimental conditions computed with the marginaleffects package.

```{r marginaleffects, message=FALSE, fig.cap="(ref:marginaleffects)"}
library(marginaleffects)
plot_predictions(fit, by = c("advice_present", "advice_correct", "expert"),
  type = "response") + ylim(c(0.3, 1))
```

## Hypothesis testing

However, we need to think about the model parameters again when we want to test hypotheses that we have theoretically derived from some research question. Because the inverse-logit link is still a continuously increasing function, positive parameter values always correspond to increases in probability and vice versa.

The `Fixed effects` table in the lme4 summary output also includes p-values for hypothesis tests with null hypotheses of the style $H_0: \beta = 0$. However, for many research questions of interest, we are not interested in these two-sided tests that refer to only a single parameter.

For our case study, imagine the following combined hypothesis: *We expect that for both experts and non-experts, correct advice leads to a higher probability of accurately diagnosing a CT scan compared to no advice presented, AND, we expect that for both experts and non-experts, incorrect advice leads to a lower probability of accurately diagnosing a CT scan compared to no advice presented.*

This combined hypothesis leads to the following four separate null hypotheses to be tested: $$
\begin{aligned}
H_{01}:\ &\beta_{a} + \beta_{c} + \beta_{ea} + \beta_{ec} \leq 0 \\
H_{02}:\ &\beta_{a} + \beta_{c} \leq 0 \\
H_{03}:\ &\beta_{a} + \beta_{ea} \geq 0 \\
H_{04}:\ &\beta_{a} \geq 0
\end{aligned}
$$

We arrive at these inequalities based on the following logic, exemplified here only for $H_{01}$: The first null hypothesis states that *an expert responding to an item while presented with correct advice has a lower or equal probability of solving the item compared to the same expert facing the same item without any advice.* This implies the following inequality for each subject $s$ and item $i$ $$
inverse\_logit[(\beta_0 + u_{0s} + u_{0i}) + \beta_e + \beta_{a} + \beta_{c} + \beta_{ea} + \beta_{ec}] \leq inverse\_logit[(\beta_0 + u_{0s} + u_{0i}) + \beta_e]
$$ which simplifies to $\beta_{a} + \beta_{c} + \beta_{ea} + \beta_{ec} \leq 0$.

We can specify and test hypotheses like these with the multcomp package [@hothornSimultaneousInferenceGeneral2008] as follows:

```{r, message=FALSE}
library(multcomp)
null_hypotheses <- c(
  "advice_present + advice_correct + expert:advice_present + 
   expert:advice_correct <= 0",
  "advice_present + advice_correct <= 0",
  "-1 * (advice_present + expert:advice_present) <= 0",
  "-1 * (advice_present) <= 0")
glht <- glht(fit, linfct = null_hypotheses)
summary(glht, test = univariate())$test$pvalues
```

Because all hypotheses tested simultaneously with the `glht` function must have the same direction, we flip the sign of inequalities three and four by multiplying them with $-1$. The multcomp package automatically adjusts p-values when multiple hypotheses are tested simultaneously [@hothornSimultaneousInferenceGeneral2008]. However, the combined null hypothesis in our exemplary research question should only be rejected if **all** individual null hypotheses are rejected [i.e., intersection-union setting; @dmitrienkoTraditionalMultiplicityAdjustment2013]. In such cases, the error probabilities do not accumulate, and we would waste power when correcting for multiple tests. Thus, we request unadjusted p-values by setting `test = univariate()` in the `summary` command. With a standard significance level of $\alpha = 0.05$, we would not reject all four null hypotheses (the p-value for hypothesis $H_{03}$ is not significant) and therefore also not reject the combined null hypothesis for this simulated dataset. Note that this decision would be wrong because we have simulated the data such that the combined alternative hypothesis is actually true in the population.

## Specification of plausible parameter values

When introducing our simulation function and simulating data for the above example, we have used theoretically plausible values as defaults for all model parameters ($\beta_0$, $\beta_e$, $\beta_a$, $\beta_c$, $\beta_{ea}$, $\beta_{ec}$, $\sigma_S$, and $\sigma_I$), but have not talked about where these numbers came from.

Ideally, one would rely on meta-analytic results or conclusive data from pilot studies. However, these are sometimes not readily available. All parameter values in our present case study have been determined based on results from related prior work. Additionally, we had repeated discussions with our affiliated domain experts in radiology to check our assumptions.

We now outline a few strategies on how to determine plausible parameter values. We have already seen in our discussion of model interpretation how we can derive the model implied probability for each experimental condition, that a subject with average ability solves an item with average difficulty. We can revert this perspective by choosing plausible probability values and deriving the parameter values implied by these probabilities (for an average subject and an average item).

```{r probtable, echo=FALSE}
probtable <- data.frame(
  col1 = 
    c("no advice, no expert", "no advice, expert", "false advice, no expert", 
      "false advice, expert", "correct advice, no expert", "correct advice, expert"),
  col2 =
    c("0.70", "0.90", "0.40", "0.85", "0.90", "0.95"),
  col3 = 
    c("$logit(0.70) = \\beta_0$",
      "$logit(0.90) = \\beta_0 + \\beta_e$", 
      "$logit(0.40) = \\beta_0 + \\beta_a$",
      "$logit(0.85) = \\beta_0 + \\beta_e + \\beta_{a} + \\beta_{ea}$",
      "$logit(0.90) = \\beta_0 + \\beta_a + \\beta_c$",
      "$logit(0.95) = \\beta_0 + \\beta_e + \\beta_a + \\beta_c + \\beta_{ea} + \\beta_{ec}$"))
apa_table(
  probtable
  , col.names = c("Experimental condition",
    "$P(Y=1|\\mathbf{X=x}, \\mathbf{U} = \\mathbf{0})$",
    "Implied equation")
  , caption = "Assumed probabilities that an average subject solves an average item in each experimental condition."
  , note = "Implied equations are derived based on the model equations and setting all random intercept terms to 0."
  , escape = FALSE
)
```

Table \@ref(tab:probtable) shows our set of assumptions concerning the probability that an average subject solves an average item for each experimental condition, as well as the corresponding equations implied by the model. The table can be used to compute the implied values for the $\beta$ parameters, starting with the first equation and reinserting the computed $\beta$ values in all following equations:

```{r}
b_0 <- qlogis(0.7)
b_e <- qlogis(0.9) - b_0
b_a <- qlogis(0.4) - b_0
b_ea <- qlogis(0.85) - b_0 - b_e - b_a
b_c <- qlogis(0.9) - b_0 - b_a
b_ec <- qlogis(0.95) - b_0 - b_e - b_a - b_c - b_ea
c(b_0 = b_0, b_e = b_e, b_a = b_a, b_c = b_c, b_ea = b_ea, b_ec = b_ec)
```

It is always possible to double-check these computations by transforming the parameter values back to probabilities, e.g. 
$$
\begin{aligned}
P(Y=1|expert = 1, advice\_present = 1, advice\_correct = 1, u_{0s} = 0, u_{0i} = 0) = \\
= inverse\_logit[\beta_0 + \beta_e + \beta_a + \beta_c + \beta_{ea} + \beta_{ec}]
\end{aligned}
$$

```{r}
plogis(b_0 + b_e + b_a + b_c + b_ea + b_ec)
```

Although the derivations above are straightforward, it is important not to misinterpret their implications: In binomial GLMMs, the average probability to solve an item (averaged across persons of varying ability and items of varying difficulty) is **not** equal to the probability that a person with average ability solves an item with average difficulty. The first perspective implies a so-called marginal interpretation, while the second one implies a conditional interpretation. For example, we determined the $\beta$ parameters in a way that corresponds to a desired conditional probability of $0.95$, that an expert with average ability solves an item with average difficulty when presented with correct advice. However, even if the model were true, we would not observe this probability value if we estimated the marginal probability in a group of experts responding to items presented with correct advice from a big sample of subjects drawn from their natural distribution of ability and items drawn from their natural distribution of difficulty.

The inequality of conditional and marginal effects in GLMMs [@fahrmeirRegressionModelsMethods2021] makes their interpretation more difficult. One must be careful when specifying parameter values based on previous studies or pilot data that use the marginal interpretation (e.g., a pilot study providing an estimate of how often neurologists make an accurate diagnosis based on brain scans). However, this does not mean that we cannot use the marginal interpretation (average probability across persons and items) to inform plausible parameter values: When parameter values have been selected, we can compute the implied marginal distributions and compare this information to our domain knowledge. Then, we can iteratively adjust the parameter values until we are satisfied with the implied distributions.

Earlier, we have already encountered one way to visualize the implied marginal distributions: We can fit our model to a simulated dataset and use the convenience functions from the marginaleffects package to compute averaged predictions that correspond to our quantities of interest. However, the model predictions will only be close to the true distribution if the simulated dataset is very large, but then the model fitting consumes a lot of time and memory. A more sophisticated strategy is to simulate a large dataset and directly compute the averages, contrasts and distributions we are interested in.

(ref:margdist1-caption) Marginal distributions including means, 66% and 95% confidence intervals for all experimental conditions.

```{r margdist1, fig.cap="(ref:margdist1-caption)"}
library(tidyverse)
library(ggdist)
dat <- simulate(n_subjects = 3000, n_items = 3000, sd_u0s = 0.5, sd_u0i = 0.5)
dat %>% 
  mutate(condition = fct_cross(
    factor(expert), factor(advice_present), factor(advice_correct))) %>%
  mutate(condition = fct_recode(condition,
    "no expert, no advice" = "0:0:0", "expert, no advice" = "1:0:0", 
    "no expert, wrong advice" = "0:1:0", "expert, wrong advice" = "1:1:0",
    "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) %>% 
  ggplot(aes(x = y_prob, y = condition)) +
  stat_histinterval(point_interval = "mean_qi", slab_color = "gray45") +
  scale_x_continuous(breaks = seq(0, 1, 0.1), limits = c(0, 1))
```

```{r}
library(tidyverse)
library(ggdist)
dat <- simulate(n_subjects = 3000, n_items = 3000, sd_u0s = 0.5, sd_u0i = 0.5)
dat %>% 
  mutate(condition = fct_cross(
    factor(expert), factor(advice_present), factor(advice_correct))) %>%
  mutate(condition = fct_recode(condition,
    "no expert, no advice" = "0:0:0", "expert, no advice" = "1:0:0", 
    "no expert, wrong advice" = "0:1:0", "expert, wrong advice" = "1:1:0",
    "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) %>%
  group_by(condition) %>%
  summarize(freq_y_1 = sum(y_bin) / n())
```

Figure \@ref(fig:margdist1) shows the model implied marginal distributions, including the mean, 66% and 95% intervals. We can see that, indeed, the average probabilities (black dots) slightly differ from the probabilities of average subjects and items considered in the previous section. This difference increases with the variability of the random effects.

Up to this point, we have not talked about plausible values for the standard deviations of the subject and item random intercepts ($\sigma_S$ and $\sigma_I$). Plots like the one above are a useful tool to decide whether the specified standard deviations are reasonable by comparing the ranges and overlap between conditions to domain knowledge.

In the next plot, we have set the item standard deviation to almost zero ($\sigma_I = 0.01$). This gives us a better way to see the variability between persons.

(ref:margdist2-caption) Marginal distributions including means, 66% and 95% confidence intervals for all experimental conditions while setting the standard deviation of item random intercepts to 0.01.

```{r margdist2, echo=FALSE, fig.cap="(ref:margdist2-caption)"}
dat <- simulate(n_subjects = 2000, n_items = 2000, sd_u0i = 0.01)
dat %>% 
  mutate(condition = fct_cross(
    factor(expert), factor(advice_present), factor(advice_correct))) %>%
  mutate(condition = fct_recode(condition,
    "no expert, no advice" = "0:0:0", "expert, no advice" = "1:0:0", 
    "no expert, wrong advice" = "0:1:0", "expert, wrong advice" = "1:1:0",
    "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) %>% 
  ggplot(aes(x = y_prob, y = condition)) +
  stat_histinterval(point_interval = "mean_qi", slab_color = "gray45") +
  scale_x_continuous(breaks = seq(0, 1, 0.1), limits = c(0, 1))
```

As an example, Figure \@ref(fig:margdist2) reveals a number of implicit assumptions about the comparison between experts and non-experts: With wrong advice, virtually all experts have a higher probability of making a correct diagnosis compared to non-experts when considering only items with average difficulty. In contrast, there is considerable overlap in probability between experts and non-experts with no advice and even higher overlap with correct advice. Patterns like these should be considered carefully and discussed with the domain experts. Parameter values ($\beta$ parameters, and $\sigma_S$) should be adjusted if the implications do not seem reasonable.

We could also have a closer look at variability between items by setting the subject standard deviation to almost zero ($\sigma_S = 0.01$).

```{r, include=FALSE}
dat <- simulate(n_subjects = 2000, n_items = 2000, sd_u0s = 0.01)
dat %>% 
  mutate(condition = fct_cross(
    factor(expert), factor(advice_present), factor(advice_correct))) %>%
  mutate(condition = fct_recode(condition,
    "no expert, no advice" = "0:0:0", "expert, no advice" = "1:0:0", 
    "no expert, wrong advice" = "0:1:0", "expert, wrong advice" = "1:1:0",
    "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) %>% 
  ggplot(aes(x = y_prob, y = condition)) +
  stat_histinterval(point_interval = "mean_qi", slab_color = "gray45") +
  scale_x_continuous(breaks = seq(0, 1, 0.1), limits = c(0, 1))
```

The final plot demonstrates that these plots are also useful for spotting standard deviations that are specified too high. For Figure \@ref(fig:margdist3), we have set $\sigma_S = 3$ and $\sigma_I = 3$. This implies that in each experimental condition, the probabilities that a subject solves an item are usually close to either 0 or 1, which is not a plausible assumption. However, these high standard deviations do not account for the inherent variability and complexity of human performance. For example, we would expect that a participant with low ability compared to other task experts to solve a difficult item with a probability substantially larger than zero even when presented with wrong advice.

(ref:margdist3-caption) Marginal distributions including means, 66% and 95% confidence intervals for all experimental conditions while setting the standard deviation of subject and item random intercepts to 3.

```{r margdist3, echo=FALSE, fig.cap="(ref:margdist3-caption)"}
dat <- simulate(n_subjects = 2000, n_items = 2000, sd_u0s = 3, sd_u0i = 3)
dat %>% 
  mutate(condition = fct_cross(
    factor(expert), factor(advice_present), factor(advice_correct))) %>%
  mutate(condition = fct_recode(condition,
    "no expert, no advice" = "0:0:0", "expert, no advice" = "1:0:0", 
    "no expert, wrong advice" = "0:1:0", "expert, wrong advice" = "1:1:0",
    "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) %>% 
  ggplot(aes(x = y_prob, y = condition)) +
  stat_histinterval(point_interval = "mean_qi", slab_color = "gray45") +
  scale_x_continuous(breaks = seq(0, 1, 0.1), limits = c(0, 1))
```

## Results

With all these considerations addressed, we are finally ready to perform a power analysis. Wrapping the `simulate` function already constructed earlier, the helper function `sim_and_analyse` performs all previous steps (simulate a dataset, fit a GLMM, compute p-values) in a single command.

```{r}
sim_and_analyse <- function(
  formula_chr = "y_bin ~ 1 + expert + advice_present + advice_correct + 
    expert:advice_present + expert:advice_correct + (1|subject) + (1|item)", 
  null_hypotheses = c("advice_present + advice_correct + 
    expert:advice_present + expert:advice_correct <= 0",
    "advice_present + advice_correct <= 0",
    "-1 * (advice_present + expert:advice_present) <= 0",
    "-1 * (advice_present) <= 0"), ...){
  require(lme4)
  require(multcomp)
  # simulate data
  dat <- simulate(...)
  # fit model
  model <- glmer(as.formula(formula_chr), data = dat, family = "binomial")
  # compute p-values
  glht <- glht(model, linfct = null_hypotheses)
  pvalues <- summary(glht, test = univariate())$test$pvalues
  setNames(pvalues, paste0("p_H0", 1:length(null_hypotheses)))
}
```

Power analysis can quickly become computationally intensive when we repeatedly simulate data and fit models for different parameter combinations or sample sizes. Thus, we use the future [@R-RJ-2021-048] and furrr [@R-furrr] packages to perform computations in parallel. First, we enable parallelization and specify how many parallel cores ("workers") of our computer to use (users can find out the maximum number of cores on their computer with the command `parallel::detectCores()`), and set a seed to make the simulation reproducible.

```{r, message=FALSE, eval=FALSE}
library(future)
plan("multisession", workers = 6)
set.seed(2)
```

The next code chunk specifies a simulation grid with different settings for both the number of subjects (`n_subjects`) and the number of items (`n_items`), each combination being repeated `rep` times. We chose 300 repetitions for the data simulation at hand as it strikes a balance between achieving a robust statistical estimate and remaining computationally feasible.
With the current settings, this simulation takes about one hour on a MacBook Pro from 2020 with M1 chip and 16 GB working memory. If you want to quickly experiment with the code yourself, a setting with `workers = 4` and `rep = 5` should finish in less than 5 minutes, even on smaller machines.

```{r, message=FALSE, echo=FALSE, eval=FALSE}
library(future)
library(furrr)
plan("multisession", workers = 6)
set.seed(2)
sim_design <- crossing(
  rep = 1:300,
  n_subjects = c(100, 150, 200, 250),
  n_items = c(10, 30, 50, 70)
) %>%
  mutate(pvalues = future_pmap(., sim_and_analyse, 
    .options = furrr_options(seed = TRUE))) %>%
  unnest_wider(pvalues)
saveRDS(sim_design, file = "results.rds")
```

```{r, message=FALSE, eval=FALSE}
library(furrr)
sim_design <- crossing(
  rep = 1:300,
  n_subjects = c(100, 150, 200, 250),
  n_items = c(10, 30, 50, 70)
) %>%
  mutate(pvalues = future_pmap(., sim_and_analyse, 
    .options = furrr_options(seed = TRUE))) %>%
  unnest_wider(pvalues)
```

The result of the computation is a data frame that contains the p-values of all tested hypotheses for each simulated dataset.
In some iterations (predominantly in conditions with small sample sizes), model estimation did not converge with the lme4 package. When the model fails to converge, it means that the statistical model being fitted to the data failed to reach a stable or valid solution during the estimation process. We do not remove these results because non-convergence can also happen when analyzing the real data we plan to collect, thus, we want to factor in this possibility to keep our simulation more realistic.

For our exemplary combined hypothesis, power is defined as the (long-run) percentage of simulations in which all four p-values of our component hypotheses are significant at the $\alpha = 0.05$ level. Based on our simulation outcomes, we compute a power estimate for each combination of `n_subjects` $\times$ `n_items` (including 95% confidence intervals) and visualize the results with the following code.[^3]

[^3]: This code was inspired by the "Mixed Design Simulation" vignette of the faux package at <https://debruine.github.io/faux/articles/sim_mixed.html>.

```{r, echo=FALSE}
sim_design <- readRDS(file = "results.rds")
```

```{r finalpwr, fig.cap="Simulation-based power estimates including 95% confidence interval of the case study for different numbers of subjects and items, based on a significance level of 0.05."}
library(binom)
alpha <- 0.05
power <- sim_design %>%
  group_by(n_subjects, n_items) %>% 
  summarise(power = mean(p_H01 < alpha & p_H02 < alpha & 
                           p_H03 < alpha & p_H04 < alpha), 
    n_sig = sum(p_H01 < alpha & p_H02 < alpha & 
                  p_H03 < alpha & p_H04 < alpha),
    n = n(),
    ci.lwr = binom.confint(n_sig, n, method = "wilson")$lower,
    ci.upr = binom.confint(n_sig, n, method = "wilson")$upper, 
    .groups = "drop")
power %>%
  mutate(across(c(n_subjects, n_items), factor)) %>%
  ggplot(aes(n_subjects, n_items, fill = power)) +
  geom_tile() +
  geom_text(aes(label = sprintf("%.2f \n [%.2f; %.2f]", 
                                power, ci.lwr, ci.upr)), 
    color = "white", size = 4) +
  scale_fill_viridis_c(limits = c(0, 1)) +
  xlab("number of subjects") + ylab("number of items")
```

<!-- How many participants/stimuli are needed for concrete example -->

As should be the case, power estimates in Figure \@ref(fig:finalpwr) increase with both the number of subjects and the number of items. The confidence intervals indicate how precisely power was estimated by our simulation. Higher precision (which would be reflected in narrower confidence intervals) could be obtained by increasing the number of repetitions (`rep`) in the simulation. In practice, data simulations are often run multiple times with adjusted combinations of sample sizes. When running for the first time, it might be revealed that power is way too low (or much higher than required) for some combinations of `n_subjects` and `n_items`. When narrowing down the best combination that achieves sufficient power while at the same time striking a good balance of how many subjects and items are practically feasible, later rounds of data simulation will typically include a smaller grid of sample sizes combined with a higher number of repetitions. This will assure high precision for the final power estimates, which are then used for the sample size justification of the future study.

Much has been written on the optimal amount of power to target in empirical research. The most prominent heuristic is to target a power of 0.8 (when combined with a type I error rate of $\alpha = 0.05$), but depending on the research goals of the study, there are often good reasons to move away from this standard depending on the research goals and resource constraints  [@lakensJustifyYourAlpha2018; @lakensSampleSizeJustification2022]. When target power has been specified, the number of subjects and the number of items in our study design can be traded against each other based on practical considerations. For the sake of the example, let the targeted power be indeed about 0.8, using an $\alpha$ of 0.05 to detect an effect of the expected size implied by our data simulation. This could be achieved by collecting data from 200 subjects (about 25% of which will be experts), each completing the same 50 items (with advice present in about 67% of cases, which is correct in about 80% of cases with present advice). If collecting data from 200 subjects is not feasible, an alternative would be to recruit 150 subjects but increase the length of the experiment to over 70 items. However, 70 items might take too long to complete for the radiologists participating in the study, who have a busy schedule. The simulation suggests that it might also be possible to plan a shorter experiment with only 30 items if it is feasible to recruit an even higher number of subjects (\> 250, to be determined by additional rounds of power analysis). Design parameters that also affect power, and which could be investigated in the simulation to find a more optimal trade-off, are the ratio of experts, the frequency of whether advice is presented and whether it is correct.

## Expected effect size vs. smallest effect size of interest: sensitivity analysis

In our case study, we have performed simulation-based power analysis from a single set of parameter values that reflect our assumptions of an expected effect size. Instead of extracting this expected effect size from meta-analyses or pilot data, which has been the main focus of previous tutorials, we have demonstrated some strategies to determine plausible parameter values in GLMMs based on domain knowledge. Domain knowledge can be considered a vague theoretical model about the data-generating process that is less formal and can only be accessed by a back-and-forth exchange in which domain experts assess the plausibility of simulated data. When sample sizes are chosen based on the results of our simulation-based power analysis, a future study will be informative to reject the null hypothesis if an effect of our *expected size* is present. However, if the true effect is indeed smaller, the power will be lower, and the study might not be sufficiently informative. A common, more conservative strategy for sample size justification is to perform power analysis for the smallest effect size of interest (SESOI). An effect smaller than the SESOI would be considered too small to be interesting or practically meaningful, even if the effect is not actually zero [@kingPointMinimalImportant2011]. For strategies on the even more difficult task of specifying a plausible SESOI, as well as a thorough discussion of various topics concerning power analysis, see [@lakensImprovingYourStatistical2022]. When domain knowledge or formal theories about the research topic of interest are too vague to specify a meaningful SESOI, it is still recommended to demonstrate power for different effect sizes in what is called *sensitivity power analysis*. By simulating power for different effect sizes (in addition to the different number of subjects and items), one can make sure that power would still be sufficient to detect smaller effect sizes than our expected effect or at least get an impression of how strongly power depends on the size of the true effect. In simple study designs, it is possible to perform sensitivity power analysis based on a single standardized effect size (e.g., analyze power in a two-sample t-test for a standardized mean difference varying between 0.1 and 0.8). However, for our case study that investigates combined hypotheses in a GLMM modeling framework, the effect size is implicitly represented by the complex distribution of probabilities within and between experimental conditions. In this setting, sensitivity power analysis would require manually specifying additional sets of plausible parameter values that reflect scenarios with smaller or larger differences between groups with respect to our specific research question. Power could then be simulated for several of these scenarios (across different numbers of subjects and items, as considered earlier).

# Outlook

<!-- Outlook -->
Beyond the specifics of our concrete case study, we want to outline six developments regarding the future role of simulation-based sample size planning in experimental research:

1.  The growing need for simulation-based sample size planning in experimental research: In order to conduct informative research using varying complex experimental designs with GLMMs formula-based heuristics and user-friendly software tools for a priori power analysis are often not suitable. Therefore, simulation-based power analysis is becoming increasingly needed since it provides experimental researchers with a tailored approach to estimating required sample sizes before data collection.

2.  Managing data simulations more easily with discrete predictor variables: Simulation-based sample size planning becomes more manageable when all predictor variables are discrete (like in the presented case study) and fixed by the study design. 
This allows researchers to focus on simulating outcome variables while avoiding the need for complex simulations of predictor values, which would introduce additional assumptions. By simplifying the simulation process, researchers can obtain reliable estimates for power or precision without compromising realistic assumptions about the data-generating process implied by the study design.

3.  Teaching data simulation skills: The ability to conduct simulation-based sample size planning is a valuable skill that should be taught to experimental researchers. By incorporating such training into research methods courses and workshops, researchers can gain a deeper understanding of statistical power and improve the quality of their experimental designs. 
Equipping researchers with the knowledge and tools to perform simulation-based sample size planning enables them to make informed decisions and enhance the rigor of their studies. 
The need to reason about how to simulate plausible data that is in line with the research hypothesis, while not violating domain expertise on how plausible data should look, might also contribute to planning more insightful studies that can answer more precise research questions [@yarkoniGeneralizabilityCrisis2022].

4.  Addressing the mismatch in effort perception: There is often a significant disconnect between the amount of effort required to perform simulation-based a priori sample size planning and the perceived effort estimated by researchers and collaborators in experimental research. Many researchers request simulation-based power analyses from statisticians or methodological experts without fully comprehending the complexity and time-consuming nature of these tailored simulations. It is crucial to raise awareness about the effort involved to ensure realistic expectations and effective collaboration between researchers and methodological experts.

5.  Recognizing the value of simulation-based design analysis: Simulation-based power analyses are not mere technicalities; they are valuable research contributions that deserve recognition in experimental research. They offer insights into the robustness and sensitivity of experimental designs, helping researchers make informed decisions about sample sizes, effect sizes, and statistical power. Their importance can be reflected by allocating them a separate publication or incorporating them as a significant component of stage 1 preregistered reports [@chambersPresentFutureRegistered2022].

6.  Integration with Open Science and preregistration practices: Simulation-based sample size planning aligns well with the principles of Open Science and preregistration in experimental research. When researchers have access to simulated data based on their pre-specified model, analyzing the collected dataset becomes straightforward and unambiguous. By preregistering their simulation-based sample size plan, researchers enhance the transparency and accountability of their experimental procedures, contributing to the credibility and reproducibility of research.

# Conclusion

In the wake of the replication crisis and myriad of underpowered experimental work, generalized linear mixed models (GLMMs) offer a flexible statistical framework to analyze experimental data with complex (e.g., dependent and hierarchical) data structures. Yet, analytic methods and software cannot be applied to conduct a priori sample size planning for GLMMs necessitating data simulation-based approaches. Through this applied tutorial, we aim to provide researchers with the necessary skills and tools to perform simulation-based sample size planning with GLMMs themselves. By incorporating GLMMs and a priori sample size planning into their work, researchers can enhance the replicability and credibility of their experiments [@yarkoniGeneralizabilityCrisis2022].

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::