### 2024-06-11 full_simulation.qmd
update bibfile, add full_simulation.qmd notebook, change notebook-links option

---
title: "Full simulation"
---

::: {.callout-note}
This Notebook is used to run the full simulation for the case study discussed in the manuscript. The simulation takes several hours to run on a Mac Book Pro (13-inch, M1, 2020) with 16 GB memory.
:::

## Simulation functions

```{r}
#| echo: true
simulate <- function(n_subjects = 100, n_items = 50,
  b_0 = 0.847, b_e = 1.350, b_a = -1.253, b_c = 2.603,
  b_ea = 0.790, b_ec = -1.393,
  sd_u0s = 0.5, sd_u0i = 0.5, ...){
  require(dplyr)
  require(faux)
  # simulate design
  dat <- add_random(subject = n_subjects, item = n_items) |>
    add_between("subject", expert = c(1, 0), .prob = c(0.25, 0.75)) |>
    mutate(advice_present = rbinom(n(), 1, prob = 2/3)) |>
    mutate(advice_correct = if_else(advice_present == 1L, 
                                    rbinom(n(), 1L, prob = 0.8), 0L)) |>
    # add random effects
    add_ranef("subject", u0s = sd_u0s) |>
    add_ranef("item", u0i = sd_u0i) |>
    # compute dependent variable
    mutate(linpred = b_0 + u0i + u0s +
        b_e * expert + b_a * advice_present + b_c * advice_correct +
        b_ea * expert * advice_present + b_ec * expert * advice_correct) |>
    mutate(y_prob = plogis(linpred)) |>
    mutate(y_bin = rbinom(n = n(), size = 1, prob = y_prob))
  dat
}
```

```{r}
#| echo: true
sim_and_analyse <- function(
  formula_chr = "y_bin ~ 1 + expert + advice_present + advice_correct + 
    expert:advice_present + expert:advice_correct + (1|subject) + (1|item)",
  contrasts = c("b8 = b2", "b2 = b6", "b7 = b1", "b1 = b5"), ...){
  require(lme4)
  require(marginaleffects)
  require(tidyr)
  # simulate data
  dat <- simulate(...)
  # fit model
  model <- glmer(as.formula(formula_chr), data = dat, family = "binomial")
  # compute contrasts
  contr_df <- expand_grid(advice_present = 0:1, advice_correct = 0:1,
    expert = 0:1)
  predictions(model, newdata = contr_df, type = "response", re.form = NA) |>
    hypotheses(hypothesis = contrasts, equivalence = c(0, 0)) |>
    data.frame()
}
```

## Run simulation

```{r}
#| echo: true
library(tidyverse)
library(future)
library(furrr)

plan("multisession", workers = 6)
set.seed(2)

sim_result <- crossing(
  rep = 1:300,
  n_subjects = c(100, 150, 200, 250),
  n_items = c(10, 30, 50, 70)
) |>
  mutate(res = future_pmap(., sim_and_analyse, 
    .options = furrr_options(seed = TRUE))) |>
  unnest(col = res)
```

## Save results file to be used in the manuscript

```{r}
saveRDS(sim_result, file = "results.rds")
```