### 2024-06-12 _tex/manuscript.tex
Built site for gh-pages

\documentclass[
  man,
  floatsintext,
  longtable,
  nolmodern,
  notxfonts,
  notimes,
  colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{apa7}

\usepackage{amsmath}
\usepackage{amssymb}

\RequirePackage{longtable}
\RequirePackage{threeparttablex}

\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
	{0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
	{-.5em}%
	{\normalfont\normalsize\bfseries\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{0.5em}%
	{0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
	{-\z@\relax}%
	{\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

\usepackage{longtable, booktabs, multirow, multicol, colortbl, hhline, caption, array, float, xpatch}
\setcounter{topnumber}{2}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{4}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.85}
\renewcommand{\textfraction}{0.15}
\renewcommand{\floatpagefraction}{0.7}

\usepackage{tcolorbox}
\tcbuselibrary{listings,theorems, breakable, skins}
\usepackage{fontawesome5}

\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{ACACAC}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582EC}
\definecolor{quarto-callout-important-color-frame}{HTML}{D9534F}
\definecolor{quarto-callout-warning-color-frame}{HTML}{F0AD4E}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02B875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{FD7E14}

%\newlength\Oldarrayrulewidth
%\newlength\Oldtabcolsep

\usepackage{hyperref}

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{241,243,245}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}

\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{newtx}

\defaultfontfeatures{Scale=MatchLowercase}
\defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}

\title{A Tutorial on Tailored Simulation-Based Sample Size Planning for
Experimental Designs with Generalized Linear Mixed Models}

\shorttitle{Simulation-Based Sample Size Planning for GLMMs}

\usepackage{etoolbox}

\authorsnames[{1},{1,2},{1},{1,3},{1,4}]{Florian Pargent,Timo K.
Koch,Anne-Kathrin Kleine,Eva Lermer,Susanne Gaube}

\authorsaffiliations{
{Department of Psychology, LMU Munich},{Institute of Behavioral Science
\& Technology, University of St.~Gallen},{Department of Business
Psychology, Technical University of Applied Sciences Augsburg},{Global
Business School for Health, University College London}}

\leftheader{Pargent, Koch, Kleine, Lermer and Gaube}

\date{2024-06-12}

\abstract{When planning experimental research, determining an
appropriate sample size and using suitable statistical models are
crucial for robust and informative results. The recent replication
crisis underlines the need for more rigorous statistical methodology and
adequately powered designs. Generalized linear mixed models (GLMMs)
offer a flexible statistical framework to analyze experimental data with
complex (e.g., dependent and hierarchical) data structures. However,
available methods and software for a priori sample size planning for
GLMMs are often limited to specific designs. Tailored data simulation
approaches are a more flexible alternative. Based on a practical case
study, the current tutorial equips researchers with a step-by-step guide
and corresponding code for conducting tailored a priori sample size
planning with GLMMs. We not only focus on power analysis but also
explain how to use the precision of parameter estimates to determine
appropriate sample sizes. We conclude with an outlook on the increasing
importance of simulation-based sample size planning.}

\keywords{tutorial, sample size planning, generalized linear mixed
model, power analysis, precision, data simulation}

\authornote{\par{\addORCIDlink{Florian
Pargent}{0000-0002-2388-553X}}\par{\addORCIDlink{Timo K.
Koch}{0000-0001-6728-2063}}

\par{ This is version 2.5 of our preprint published at
\url{https://doi.org/10.31234/osf.io/rpjem}. All materials (reproducible
manuscript, including R code and all simulation results) are available
in the project's repository on the Open Science Framework (OSF) at
\url{https://osf.io/dhwf4/}. A Quarto Manuscripts website is hosted at
\url{https://Timo-Ko.github.io/glmm_simulation_tutorial/}.  The authors
declare that there were no conflicts of interest with respect to the
authorship or the publication of this article. This research was funded
by a grant from the Volkswagen Foundation (Grant
No.~98525).   Author roles were classified using the Contributor Role Taxonomy (CRediT; https://credit.niso.org/) as follows: Florian
Pargent:   formal
analysis, methodology, visualization, writing, editing; Timo K.
Koch:   conceptualization, formal
analysis, writing, editing; Anne-Kathrin Kleine:   formal
analysis, editing; Eva Lermer:   conceptualization, funding
aquisition, editing; Susanne
Gaube:   conceptualization, supervision, editing}
\par{Correspondence concerning this article should be addressed to Timo
K. Koch, Institute of Behavioral Science \& Technology, University of
St.~Gallen, Torstrasse
25, St.~Gallen 9000, Switzerland, Email: timo.koch@unisg.ch}
}

\makeatletter
\let\endoldlt\endlongtable
\def\endlongtable{
\hline
\endoldlt
}
\makeatother

\urlstyle{same}

\usepackage{float}
\usepackage{tabularray}
\usepackage[normalem]{ulem}
\usepackage{graphicx}
\UseTblrLibrary{booktabs}
\UseTblrLibrary{siunitx}
\NewTableCommand{\tinytableDefineColor}[3]{\definecolor{#1}{#2}{#3}}
\newcommand{\tinytableTabularrayUnderline}[1]{\underline{#1}}
\newcommand{\tinytableTabularrayStrikeout}[1]{\sout{#1}}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

% From https://tex.stackexchange.com/a/645996/211326
%%% apa7 doesn't want to add appendix section titles in the toc
%%% let's make it do it
\makeatletter
\xpatchcmd{\appendix}
  {\par}
  {\addcontentsline{toc}{section}{\@currentlabelname}\par}
  {}{}
\makeatother

\begin{document}

\maketitle

\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\setlength\LTleft{0pt}

\section{Introduction}\label{introduction}

When planning experimental research, it is essential to determine an
appropriate sample size and to use appropriate statistical models to
analyze the data to ensure that the results obtained are both robust and
informative (\citeproc{ref-lakensSampleSizeJustification2022}{Lakens,
2022a}). The recent replication crisis has illustrated many challenges
surrounding the reproducibility and reliability of study findings
(\citeproc{ref-yarkoniGeneralizabilityCrisis2022}{Yarkoni, 2022}). As a
result, there is a growing need for more rigorous statistical
methodology and the adoption of adequately powered experimental designs.
Multiple easy-to-use software solutions exist for simple statistical
models and experimental designs. However, many researchers lack the
skills and tools to conduct ``a priori'' (i.e., before data collection)
sample size planning for more complex research designs such as flexible
generalized linear mixed models (GLMM) framework. In the present work,
we provide a tutorial on how to determine adequate sample sizes by
performing tailored simulation-based sample size planning for GLMMs.
After introducing some theoretical background on sample size planning,
we review existing software solutions in R and discuss under which
circumstances tailored data simulations are necessary. Then we describe
general steps and decisions involved in tailored data simulation. To
illustrate the details of these steps, we finish with a hypothetical
case study from the field of human-AI (artificial intelligence)
interaction research.

To benefit most of this tutorial paper, we recommend readers to
familiarise themselves with basic statistical concepts like hypothesis
tests (HTs) and their statistical power as well as confidence intervals
(CIs) and their precision
(\citeproc{ref-kumleEstimatingPowerGeneralized2021}{Kumle et al., 2021};
\citeproc{ref-lakensImprovingYourStatistical2022}{Lakens, 2022b};
\citeproc{ref-riesthuisSimulationBasedPowerAnalyses2024}{Riesthuis,
2024}). Some knowledge of causal inference is useful but not necessary
(\citeproc{ref-deffnerCausalFrameworkCrossCultural2022}{Deffner et al.,
2022}; \citeproc{ref-lundbergWhatYourEstimand2021}{Lundberg et al.,
2021}). In addition, readers should have a rough understanding of R
(\citeproc{ref-wickhamDataScienceImport2023}{Wickham et al., 2023}) and
how to simulate data. For data simulation, we use functions from the
tidyverse (\citeproc{ref-wickhamWelcomeTidyverse2019}{Wickham et al.,
2019}) and the faux package (\citeproc{ref-R-faux}{DeBruine, 2023}).
Finally, readers should be familiar with regression modeling in general
and GLMMs in particular. In this tutorial, we simulate data by manually
specifying the model equation of a GLMM that represents our assumed data
generating process
(\citeproc{ref-debruineUnderstandingMixedEffectsModels2021}{DeBruine \&
Barr, 2021}). It is not necessary to understand the technical details of
how GLMMs are estimated. However it is crucial to understand the
structure of a basic GLMM (e.g., logistic regression with random
intercepts) and how the model assumes that the values in the dependent
variable are determined by the predictor variables and the random
effects.

\section{Theoretical background}\label{theoretical-background}

\subsection{Planning for statistical power or
precision}\label{planning-for-statistical-power-or-precision}

Conducting research with insufficiently large sample sizes can have many
negative consequences. First, experiments may yield inconclusive or
misleading results, hindering the accumulation of knowledge. Second,
studies that are doomed to never finding a postulated effect waste
resources by consuming time, effort, and funding without delivering
meaningful results. For these reasons, many journals and funding bodies
now require that a justification of sample size is included in study
protocols and grant proposals, recognizing its significance in ensuring
robust and meaningful findings. Although scientists often do not justify
sample size or use general heuristics from the literature, resource
constraints often play an important role
(\citeproc{ref-lakensSampleSizeJustification2022}{Lakens, 2022a}). But
ideally, a suitable sample size should be determined a priori (i.e.,
before the study is conducted) based on some meaningful computation to
ensure that the study will be able to fulfill its purpose.

The majority of empirical studies in psychology and other social
sciences apply hypothesis testing. As a consequence, the dominant
approach for determining an adequate sample size is based on power
analysis (i.e., planning for power)
(\citeproc{ref-lakensSampleSizeJustification2022}{Lakens, 2022a};
\citeproc{ref-maxwellSampleSizePlanning2008}{Maxwell et al., 2008}).
Statistical power is defined as the probability that a HT has a
significant p-value when analyzing repeated samples from a population
with a true effect of some pre-specified size
(\citeproc{ref-cohenPowerPrimer1992}{Cohen, 1992}). Less formally, power
is described as the probability that a HT correctly rejects the null
hypothesis when the alternative hypothesis is true. If the sample size
(i.e., the number of participants and/or stimuli) used for data
collection is insufficient to detect the effects or relationships being
investigated with high probability, the study is considered
``underpowered''. When planning for power, a target is set for the
statistical power of a HT of interest. Assuming some effect size of
interest and a desired significance level, a minimum sample size can be
determined that, on average, would guarantee reaching this target.
Although it is recommended to justify the desired significance level and
power (\citeproc{ref-lakensJustifyYourAlpha2018}{Lakens, Adolfi, et al.,
2018}), most empirical studies adopt the heuristic of \(\alpha = 0.05\)
and \(1 - \beta = 0.80\).

In contrast to power analysis, sample size planning can also be based on
the precision of parameter estimates (i.e., planning for precision or
planning for accuracy)
(\citeproc{ref-lakensSampleSizeJustification2022}{Lakens, 2022a};
\citeproc{ref-maxwellSampleSizePlanning2008}{Maxwell et al., 2008}). Not
all research questions are best answered using hypothesis testing. It
has been argued that basic research rarely requires making discrete
decisions on whether some effect has been ``discovered'' and should thus
shift from a hypothesis testing towards an estimation framework
(\citeproc{ref-cummingNewStatisticsWhy2014}{Cumming, 2014};
\citeproc{ref-kruschkeBayesianNewStatistics2018}{Kruschke \& Liddell,
2018};
\citeproc{ref-mcelreathStatisticalRethinkingBayesian2020}{McElreath,
2020}). Although this view is not without critique
(\citeproc{ref-uyguntuncEpistemicPragmaticFunction2023}{Uygun Tunç et
al., 2023}), at least for exploratory or pilot studies where little
previous research has been conducted, more scientists seem to agree that
simply estimating the effects of interest and making the estimation
uncertainty transparent by reporting CIs is more useful. Assuming that
no HTs are conducted for a planned study, power analysis is not relevant
for sample size planning. Nonetheless, the sample size still has a
crucial effect on how informative the planned study will be, because an
effect of interest is estimated more precisely with bigger samples. In
the precision framework, the target quantity commonly used for sample
size planning is the expected width of a CI
(\citeproc{ref-kelleySampleSizePlanning2006}{Kelley \& Rausch, 2006};
\citeproc{ref-lakensSampleSizeJustification2022}{Lakens, 2022a};
\citeproc{ref-maxwellSampleSizePlanning2008}{Maxwell et al., 2008}). The
values inside a CI are often interpreted as plausible values for the
quantity of interest it is supposed to estimate. More formally, a CI
with confidence level 0.95 provides the smallest interval with the
property that upon repeated sampling, 95\% of individual CIs would
include the true quantity of interest. Thus, a narrow CI is more
informative about the size of the true effect than a wide interval.
Apart from the desired confidence level, the width of a CI depends
strongly on the sample size. Because bigger samples carry more
information, they lead to smaller CIs. When planning for precision, a
target can be set for the expected width of a CI of interest. Assuming
some effect size of interest and a desired confidence level, a minimum
sample size can be determined that, on average, would guarantee reaching
this target. Because planning for precision is still rare, there are no
common heuristics on how to choose the desired width of the CI
(\citeproc{ref-lakensSampleSizeJustification2022}{Lakens, 2022a}).

\subsection{Generalized linear mixed models
(GLMMs)}\label{generalized-linear-mixed-models-glmms}

As study designs become more complex, psychological researchers require
more sophisticated statistical models to capture the nuanced
relationships and grouping structures introduced by them
(\citeproc{ref-yarkoniGeneralizabilityCrisis2022}{Yarkoni, 2022}). GLMMs
(also called multilevel models) are gaining increasing popularity
because they offer great flexibility
(\citeproc{ref-fahrmeirRegressionModelsMethods2021}{Fahrmeir et al.,
2021}). GLMMs are an extension of LMMs (Linear Mixed Models), which are,
in turn, extensions of linear regression models that account for
correlated data, including hierarchical structures
(\citeproc{ref-fahrmeirRegressionModelsMethods2021}{Fahrmeir et al.,
2021}). In this context, correlated data means that the value in the
outcome variable for one observation may be related to the value of
another observation in a systematic way that is not already accounted
for by the usual (fixed) predictor variables (e.g., age of
participants). This correlation can arise for various reasons: For
instance, responses to some stimuli from some participants might be more
similar because the same person was measured twice (repeated
measurements), participants come from the same neighborhood
(clustering), or participants responded to the same stimulus (stimulus
effects). Thus, modeling such correlations is important whenever the
data has a clear structure, while the grouping variables can be
hierarchically nested (e.g., grouping variables students and schools:
each student belongs to exactly one school) or cross-classified (e.g.,
grouping variables students and math exercises: each student is
presented with several math exercises). LMMs are used when the outcome
variable is continuous and follows a normal distribution. They allow for
the modeling of fixed effects, which capture the relationships between
the usual predictors and the outcome, as well as random effects, which
account for the different types of correlation structure and grouping
effects. Random effects are typically assumed to follow a normal
distribution with a mean of zero and a variance that quantifies the
heterogeneity across groups. GLMMs extend the LMM framework to
accommodate non-normally distributed continuous and categorical outcome
variables. GLMMs involve a link function that connects the linear
combination of predictor variables to the expected value of the outcome
variable. The link function allows for modeling the relationship between
predictors and the outcome in a non-linear way that is appropriate for
the specific distribution family of the outcome variable. For example,
think of an experiment with different design factors (e.g., picture
positions, headline aesthetics) impacting the likelihood of users
clicking on an online advertisement. The participants' behavior is
measured repeatedly over several sessions. The click patterns of
participants in one session are likely to be correlated with their
previous sessions and the outcome variable is binary (click/no click)
for each session, which follows a binomial distribution.

\section{Simulation-based sample size planning with
GLMMs}\label{simulation-based-sample-size-planning-with-glmms}

To our knowledge, existing approaches for sample size planning for GLMMs
have exclusively focused on planning for power. Power analysis methods
for multilevel models can be categorized into formula-based methods and
simulation-based methods
(\citeproc{ref-murayamaSummarystatisticsbasedPowerAnalysis2022}{Murayama
et al., 2022}). Formula-based methods rely on formulas to calculate
power directly while simulation-based methods rely on repeatedly
simulating data with a known true effect size and estimating power
empirically (i.e., what percentage of simulated datasets produces a
significant p-value). Available formula-based software packages for
multilevel models often do not include GLMMs or are limited to very
simple designs
(\citeproc{ref-murayamaSummarystatisticsbasedPowerAnalysis2022}{Murayama
et al., 2022};
\citeproc{ref-westfallStatisticalPowerOptimal2014}{Westfall et al.,
2014}), making it necessary to build data simulations tailored
specifically to the study design. A number of tutorials have been
published describing how to perform such simulation-based power analysis
for multilevel models
(\citeproc{ref-arendStatisticalPowerTwolevel2019}{Arend \& Schäfer,
2019}; \citeproc{ref-brysbaertPowerAnalysisEffect2018}{Brysbaert \&
Stevens, 2018};
\citeproc{ref-debruineUnderstandingMixedEffectsModels2021}{DeBruine \&
Barr, 2021}; \citeproc{ref-greenSIMRPackagePower2016a}{Green \& MacLeod,
2016}; \citeproc{ref-kainPracticalGuidePower2015}{Kain et al., 2015};
\citeproc{ref-kumleEstimatingPowerGeneralized2021}{Kumle et al., 2021};
\citeproc{ref-lafitSelectionNumberParticipants2021}{Lafit et al., 2021};
\citeproc{ref-zimmerSampleSizePlanning2022}{Zimmer et al., 2022}).
However, many of these tutorials focus on linear mixed models (LMMs) and
the common study designs (but see
\citeproc{ref-kumleEstimatingPowerGeneralized2021}{Kumle et al., 2021},
for a tutorial that also covers more advanced settings). This narrow
focus provides limited guidance for researchers faced with more complex
study designs, especially when little prior knowledge about plausible
effect sizes is available (see the discussion in
\citeproc{ref-kumleEstimatingPowerGeneralized2021}{Kumle et al., 2021}).
Simulation-based power analysis with GLMMs requires making a range of
assumptions: The (conditional) distribution assumption specifies the
distributional family for the outcome variable. Assumptions about the
random effects include the assumption of normality (i.e., that the
random effects follow a normal distribution) and the covariance
structure among the random effects (i.e., if and how they are
correlated). Making these decisions requires understanding the
underlying assumptions of the model and ensuring they align with the
characteristics of the data being analyzed. Existing tutorials often
rely on heuristics for specifying variance components (e.g., the
standard deviation of random intercepts) or assume that results from
meta-analyses or data from pilot studies are available to determine
plausible values for all model parameters. However, in practice,
knowledge about those parameters from prior studies is often limited,
which makes specifying assumptions a practical challenge (see the
discussion in (\citeproc{ref-maxwellSampleSizePlanning2008}{Maxwell et
al., 2008}) and
(\citeproc{ref-kumleEstimatingPowerGeneralized2021}{Kumle et al.,
2021})).

INSERT TABLE 1 HERE!

In Table 1, we give a short review of existing R packages that can be
used for power analysis for GLMMs.

\section{When to use tailored data
simulation?}\label{when-to-use-tailored-data-simulation}

Performing tailored simulation-based sample size planning is more
complicated and time-consuming than using the existing software tools
outlined in TABLE 1. The most important circumstances under which
tailored simulation-based sample size planning is necessary are 1)
complex study designs, 2) complex statistical hypotheses, 3) planning
for precision, 4) no available prior studies or pilot data.

Requirements of real-world studies are often more complex than the
simplified designs assumed by many user-friendly software packages for
sample size planning. One frequent issue in applied data analysis is
missing data, and there can be various reasons for this
(\citeproc{ref-littleStatisticalAnalysisMissing2014}{Little \& Rubin,
2014}). For example, data can be missing completely at random (e.g.,
because an electronic measurement device randomly failed for some
technical reasons). Alternatively, subjects might systematically drop
out or produce missing data, but this dropout can be explained by some
attributes also measured in the dataset (e.g., older subjects have a
higher probability to refuse answering a question on income). In a more
complicated scenario, missing data in some variable is caused by the
measured attribute itself (e.g., wealthy people are more likely to
refuse reporting their income). Moreover, many experimental designs
contain conditions in which values of the predictor variables are
missing by design. This can make data analysis more complicated because
predictors have to be coded in specific ways that prevent the estimated
GLMM from becoming unidentified. Whether missing data has an effect on
the sample size planning depends on our theoretical assumptions on how
the missingness is caused. However, it is often challenging to decide
whether missing data can be safely ignored in the data analysis and
sample size planning process based on a merely theoretical approach
(\citeproc{ref-gomilaMissingDataExperiments2022}{Gomila \& Clark,
2022}). Tailored simulation-based approaches offer the possibility to
include the assumed process of how data become missing in the data
simulation, thereby determining the required sample based on simulated
datasets that contain missing values (for an example, see
\citeproc{ref-lanePowerStrugglesEstimating2018}{Lane \& Hennes, 2018}).
As a byproduct, the simulated datasets can also be used to test whether
the intended data analysis provides the expected (unbiased) results,
despite the missing data. Although GLMMs can handle a large variety of
outcome variables, researchers are becoming increasingly aware that many
datasets might profit from even more sophisticated models. Common
examples are zero-inflated outcomes, censoring, and nonlinear predictor
effects that can be modeled with the R packages glmmTMB
(\citeproc{ref-brooksGlmmTMBBalancesSpeed2017}{Brooks et al., 2017}) or
brms (\citeproc{ref-burknerAdvancedBayesianMultilevel2018}{Bürkner,
2018}). Tailored simulation-based approaches do not share the same
limitation than the existing software solutions for power analysis that
focus exclusively on GLMMs. As long as there is a software package
available to estimate the model of interest, it is always possible to
perform tailored simulation-based sample size planning.

The most common hypotheses tested in psychological research are of the
type \(H_0: \beta = 0\), where \(\beta\) is a slope or intercept of a
regression model. However, many research questions in psychology
actually require testing more complex statistical hypotheses. In the new
era of preregistration and registered reports
(\citeproc{ref-chambersPresentFutureRegistered2022}{Chambers \&
Tzavella, 2022}), most research questions should be tested with directed
hypotheses because good theories at least postulate whether some
psychological effect of interest is positive or negative. Even better
theories should be able to specify the smallest effect sizes of interest
(SESOI) that must be exceeded if the effect has any practical relevance
(\citeproc{ref-lakensEquivalenceTestingPsychological2018}{Lakens,
Scheel, et al., 2018}). In combination, this might require a test such
as \(H_0: \beta \leq 0.1\). More elaborate research questions often
require testing hypotheses that consist of a combination of model
parameters, for example testing simple slopes
(\citeproc{ref-preacherComputationalToolsProbing2006}{Preacher et al.,
2006}) with a hypothesis such as \(H_0: \beta_0 + \beta_1 \leq 0\). If
the research question consists only of a single hypothesis of this sort,
it might be possible to reduce the hypothesis to a single regression
coefficient by clever coding and/or centering of predictor variables.
However, interesting research questions often consist of combined
hypotheses that consist of more than one separate statistical hypothesis
(for a tutorial on contrast analysis in GLMMs, see
\citeproc{ref-schadHowCapitalizePriori2020}{Schad et al., 2020}). For
example, a combined null hypothesis \(H_0\) might consist of two single
null hypotheses \(H_{01}: \beta_1 \leq 0\) and
\(H_{02}: \beta_0 + \beta_1 \leq 0\). For some research questions, the
combined null hypothesis \(H_0\) would be rejected if both \(H_{01}\)
\emph{AND} \(H_{02}\) are rejected. For other research questions, the
combined null hypothesis \(H_0\) would be rejected if \(H_{01}\)
\emph{OR} \(H_{02}\) \emph{OR} both are rejected. If the global
hypothesis \(H_0\) is combined with \emph{OR}, the p-values of the
single hypotheses must be corrected for multiple testing to avoid
\(\alpha\)-inflation for the global hypothesis
(\citeproc{ref-dmitrienkoTraditionalMultiplicityAdjustment2013}{Dmitrienko
\& D'Agostino, 2013}). However, if the global hypothesis \(H_0\) is
combined with \emph{AND}, a correction for multiple testing is not
necessary but rather a mistake that unnecessarily reduces the power of
the global HT. None of the software packages for sample size planning in
table X can handle combined hypotheses as discussed here and only some
can handle directed hypotheses. In contrast, our case study will
demonstrate how we can test directed combined hypotheses with tailored
simulation-based sample size planning.

Although planning for precision
(\citeproc{ref-cummingNewStatisticsWhy2014}{Cumming, 2014};
\citeproc{ref-lakensSampleSizeJustification2022}{Lakens, 2022a};
\citeproc{ref-maxwellSampleSizePlanning2008}{Maxwell et al., 2008}) has
been increasingly discussed as a useful strategy for empirical research,
all available software packages for sample size planning with GLMMs are
based on power analysis. Therefore, researchers that want to want to
apply an estimation strategy in their studies, without testing any
statistical hypotheses, currently cannot use the software packages
outlined in TABLE 1. However, tailored simulation-based sample size
planning can easily handle the planning for precision approach
(\citeproc{ref-maxwellSampleSizePlanning2008}{Maxwell et al., 2008}).
The only change in procedure is that instead of computing HTs for each
simulated dataset and estimating statistical power across repetitions,
CIs are computed for each simulated dataset, and the expected width is
estimated.

All frameworks for sample size planning require the user to make
assumptions about the expected effect size. Assuming the true effect is
of this size (or greater), one can compute the (minimum) power of a HT
or the (maximum) expected width of a CI. Existing software packages for
sample size planning for GLMMs usually require to provide the assumed
effect in the unit of some standardized measure of effect size. When the
researcher has access to similar studies or pilot data, providing such
standardized effect sizes is feasible. However, note that choosing
effect sizes based on small pilot studies is generally not recommended,
as those estimates can be heavily biased
(\citeproc{ref-albersWhenPowerAnalyses2018a}{Albers \& Lakens, 2018};
\citeproc{ref-lakensSampleSizeJustification2022}{Lakens, 2022a}).
Providing an informed standardized effect size can be an almost
impossible challenge when no prior studies of pilot data are available.
This problem is further exacerbated by the fact that GLMMs are so
flexible that general heuristics of what should be considered a small
effect do not exist or are difficult to defend. In our experience, using
domain knowledge to construct a tailored data simulation is the only
solution to determine plausible effect sizes in the absence of prior
evidence. It would be possible to use these tailored simulations to
extract plausible values for standardized effect sizes that could then
be inserted in existing software packages for sample size planning.
However, we would argue that when tailored data simulations necessary to
determine effect sizes anyway, performing the whole sample size planning
in a customized way is preferred over using the existing software
packages.

\section{General steps in tailored simulation-based sample size
planning}\label{general-steps-in-tailored-simulation-based-sample-size-planning}

Although the details differ depending on the specific study
characteristics, each tailored simulation-based sample size planning
requires a series of steps and decisions. We will introduce each step in
a theoretical section, followed by the practical application based on a
hypothetical case study. All code in this manuscript and simulation
results are available in the project's repository on the Open Science
Framework (\url{https://osf.io/dhwf4/}).

\subsection{Step 1: Define the
estimand}\label{step-1-define-the-estimand}

\subsubsection{THEORY}\label{theory}

The first step in every research process should be a clear definition of
the theoretical \emph{estimand}
(\citeproc{ref-lundbergWhatYourEstimand2021}{Lundberg et al., 2021}),
i.e.~the theoretical quantity which is necessary to answer a specific
research question. The estimand consists of a quantity, that can be
described for each unit under investigation and a clear definition of
the target population, for which the quantity is of interest. For
example, an estimand might be the probability that a clinical
psychologists makes the correct diagnosis for a psychiatric patient with
major depression, averaged across all clinical psychologists and
depressed patients in psychiatric institutions in Germany.

The estimand should always be defined outside of any statistical model,
because there are usually a range of statistical methods that can be
used to estimate the same estimand, depending on the study design (e.g.,
a randomized experiment) that will produce the observed data in the
planned study. For many common research questions in psychology, the
estimand can be expressed as a statistical quantity that can be
estimated with a regression model, for example a single \(\beta\)
coefficient. However, this is not possible for all estimands, which is
why the literature discusses many estimation strategies beyond
regression
(\citeproc{ref-deffnerCausalFrameworkCrossCultural2022}{Deffner et al.,
2022}; \citeproc{ref-lundbergWhatYourEstimand2021}{Lundberg et al.,
2021}).

\subsubsection{PRACTICE}\label{practice}

In the present hypothetical case study, we consider the effectiveness of
feedback provided by an artificial intelligence (AI) embedded in a
diagnostic decision support system. The context is a clinical setting,
where expert radiologists and students under training must detect
bleeding based on head scans from computer tomography (CT). In the
investigated AI-enabled diagnostic decision support system, an AI model
can provide initial diagnostic advice, which can be used as guidance by
the humans who are required to make the final diagnostic decision. The
research goal is to validate the effectiveness of the AI-enabled advice.
We consider the AI-enabled advice as effective, if the following pattern
holds, which we will first describe verbally:

\emph{We expect that for BOTH expert radiologists and medical students,
correct AI-advice leads to a higher probability of accurately diagnosing
a CT scan compared to no AI-advice presented, AND, we expect that for
BOTH experts and non-experts, incorrect advice leads to a lower
probability of accurately diagnosing a CT scan compared to no advice
presented.}

It becomes clear that our estimand consists of four comparisons between
experimental conditions
(\citeproc{ref-lundbergWhatYourEstimand2021}{Lundberg et al., 2021}).
However, the verbal description is still somewhat vague, which is why we
try to give a more precise expression for each comparison:

\[
\begin{aligned}
& P(\text{correct diagnosis} | \text{correct advice, average expert, average scan}) \\
& \quad - P(\text{correct diagnosis} | \text{no advice, average expert, average scan})
\end{aligned}
\]

\[
\begin{aligned}
& P(\text{correct diagnosis} | \text{no advice, average expert, average scan}) \\
& \quad - P(\text{correct diagnosis} | \text{wrong advice, average expert, average scan})
\end{aligned}
\] \[
\begin{aligned}
& P(\text{correct diagnosis} | \text{correct advice, average student, average scan}) \\
& \quad - P(\text{correct diagnosis} | \text{no advice, average student, average scan})
\end{aligned}
\]

\[
\begin{aligned}
& P(\text{correct diagnosis} | \text{no advice, average student, average scan}) \\
& \quad - P(\text{correct diagnosis} | \text{wrong advice, average student, average scan})
\end{aligned}
\] For example, the first expression is concerned with the difference
between the probability that a correct diagnosis is made if correct
AI-advice is presented and the probability that a correct diagnosis is
made if no AI-advice is presented. This contrast is quantified for a
hypothetical \emph{typical} expert and a \emph{typical} scan, where
\emph{typical} is usually defined as an average score on all attributes
of the expert or scan.\footnote{Note that a different estimand would be
  the so-called average treatment effect (ATE). For the ATE, the
  probability contrast is defined for each combination of expert and
  scan, and then these contrasts are averaged across all experts and
  scans from the target population
  (\citeproc{ref-lundbergWhatYourEstimand2021}{Lundberg et al., 2021}).}.
To complete our definition of the estimand, we have to define our target
population that consists of persons, stimuli, and AI-advice: With
respect to persons, we are only interested in expert radiologists and
medical students at German universities. With respect to stimuli, we are
only interested in the head CT scans made from subjects that do or do
not suffer from intracerebral hemorrhage. Lastly, we are only interested
in AI-advice given by a specific AI-enabled diagnostic decision support
system.

Although the estimand is initially defined outside of any statistical
model, it is only useful if we find a way to estimate it based on
observed data. For our exemplary research question, it is possible to
construct an experimental study, where all participants are confronted
with the same set of head CT scans, but the kind of AI-advice given for
each scan is randomly assigned within participants. This random
intervention allows us to produce an empirical estimate of our estimand,
although, in reality, each person receives only one kind of AI-advice
(correct advice, wrong advice, no advice) for each scan. We will see
later how each of the probability expressions in our estimand can be
modeled with the same GLMM. Estimating this GLMM based on the data
observed in our planned study will produce an estimate for each
probability, and these estimates can then be combined to compute an
estimate for each of the four probability contrasts. For pedagogical
reasons, we will skip the concrete definition of our estimate until we
have discussed how to simulate data based on a concrete GLMM in the next
section.

\subsection{Step 2: Simulate the data generating
process}\label{step-2-simulate-the-data-generating-process}

\paragraph{THEORY.}\label{theory-1}

When the estimand has been defined, the next step in the research
process is to write code that simulates the (assumed) data generating
process of the planned study. This requires specifying a generative
process for all predictor variables used in the final data analysis.
While such assumptions can be quite challenging for observational
studies or continuous predictor variables, this is less of a problem for
experimental studies with only categorical predictor variables. When all
predictor variables have been simulated, one can use the structure of a
suitable GLMM to simulate the dependent variable. To simulate the GLMM,
one requires plausible values for all model parameters. We will discuss
strategies on how these values can be obtained later. Because we have
full control over the data generating process in a tailored simulation,
it is possible to model specific aspects of the planned study, like data
missing by design or assuming that some subjects might drop out. The
quality of the results of the sample size planning crucially depends on
the plausibility of the simulated data generating process. However, we
would argue that even a strongly simplified data generating process
(e.g.~only a small number of interaction effects; only random intercepts
and no random slopes; assuming that data is missing completely at
random) can yield informative results.

\paragraph{PRACTICE.}\label{practice-1}

In our case study, we simulate data for an experiment where the
diagnostic performance of users of an AI-enabled diagnostic decision
support system will be evaluated. Radiologists (task experts) and
students/interns (non experts) review a series of head computer
tomography (CT) scans to assess the presence of a bleeding. An AI model
provides initial diagnostic advice to assist their decision-making. In
the control condition, no AI advice is presented. When AI advice is
given, this advice can be either correct or incorrect. The type of
advice (no advice, wrong advice, correct advice) is randomized within
subjects across brain scans. After reviewing a CT scan, participants
deliver a medical diagnosis (bleeding or no bleeding), which may be
either accurate or inaccurate. This experimental design introduces some
missing values by design since the advice is neither correct nor
incorrect when no advice is present, which must be taken into account
when simulating and analyzing the data. In this example, recruiting task
experts (i.e., radiologists) is more challenging due to their limited
availability, while non-experts (i.e., students/interns) are more
readily accessible. The goal of simulation-based sample size planning is
to determine how many task experts and non-experts must be recruited to
achieve sufficient statistical power or precision in the planned
experiment.

\subparagraph{Our specific GLMM.}\label{our-specific-glmm}

In a GLMM, the expected value of the dependent variable \(Y\)
conditioned on the vector of predictor variables \(\mathbf{X}\) and
random effects \(\mathbf{U}\), transformed by a link function \(g()\) is
modeled as a linear combination \(\eta\) of the predictor variables
\(\mathbf{X}\), the random effects \(\mathbf{U}\) and the model
parameters \(\mathbf{\beta}\)
(\citeproc{ref-fahrmeirRegressionModelsMethods2021}{Fahrmeir et al.,
2021}): \[
g(E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = \eta
\] Equivalently, the conditional expected value is modeled as the linear
combination \(\eta\), transformed by the inverse link function
\(g^{-1}()\): \[
E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u}) = g^{-1}(\eta)
\] If the dependent variable (i.e., diagnostic decision) \(Y\) is a
binary variable with values \(0\) (i.e., inaccurate), or \(1\) (i.e.,
accurate), the conditional expected value is equivalent to the
probability: \[
P_{si} := P(Y = 1|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})
\] In our case study, \(P_{si}\) is the conditional probability that
subject \(s\) gives the correct response to item (i.e., CT scan) \(i\).

In such a setting, we model this probability as \[
P_{si} = \text{inverse\_logit}(\eta_{si})
\] with the inverse-logit link
\(g^{-1}(\eta_{si}) = inverse\_logit(\eta_{si}) = \frac{exp(\eta_{si})}{1 + exp(\eta_{si})}\)
or equivalently \[
\text{logit}(P_{si}) = \eta_{si}
\] with the logit link
\(g(P_{si}) = \text{logit}(P_{si}) = \text{ln} (\frac{P_{si}}{1 - P_{si}})\).

In our case study, the probability of making an accurate diagnostic
decision is assumed to depend on the predictors:

\begin{itemize}
\tightlist
\item
  \(advice\_present_{si}\): whether subject \(s\) was presented with AI
  advice (1) or not (0) when asked to assess item \(i\)
\item
  \(advice\_correct_{si}\): whether this advice was correct (1) or not
  (0)
\item
  \(expert_s\): whether subject \(s\) was a task expert (1) or not (0)
\end{itemize}

and the random effects:

\begin{itemize}
\tightlist
\item
  \(u_{0s}\): the deviation of subject \(s\) from the average ability to
  solve an item (i.e., CT scan) with average difficulty; assumed to be
  distributed as \(u_{0s} \sim N(0, \sigma_S^2)\)
\item
  \(u_{0i}\): the deviation of item (i.e., CT scan) \(i\) from the
  average difficulty to be solved by a person with average ability;
  assumed to be distributed as \(u_{0i} \sim N(0, \sigma_I^2)\)
\end{itemize}

In total, we assume the model \[
\begin{aligned}
\text{logit}[P_{si}] =\ (&\beta_0 + u_{0s} + u_{0i}) + \\
&\beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
&\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}
\end{aligned}
\] or equivalently \[
\begin{aligned}
P_{si} = \text{inverse\_logit}[&(\beta_0 + u_{0s} + u_{0i}) + \\
&\beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
&\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}]
\end{aligned}
\] with model parameters \(\beta_0\), \(\beta_e\), \(\beta_a\),
\(\beta_c\), \(\beta_{ea}\), \(\beta_{ec}\), \(\sigma_S\), and
\(\sigma_I\).

In the GLMM literature, this would be called a binomial GLMM with two
random intercepts (for subjects and items), two level-1 predictors
(\(advice\_present\), \(advice\_correct\)), one level-2 predictor
(\(expert\)) and two cross-level interactions
(\(expert \cdot advice\_present\), \(expert \cdot advice\_correct\)). To
limit complexity, we do not consider random slopes, additional
predictors or higher-level interactions.

\subparagraph{Simulation function in R.}\label{simulation-function-in-r}

The following R function simulates a full dataset structured according
to the design of our case study.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{simulate }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(}\AttributeTok{n\_subjects =} \DecValTok{100}\NormalTok{, }\AttributeTok{n\_items =} \DecValTok{50}\NormalTok{,}
  \AttributeTok{b\_0 =} \FloatTok{0.847}\NormalTok{, }\AttributeTok{b\_e =} \FloatTok{1.350}\NormalTok{, }\AttributeTok{b\_a =} \SpecialCharTok{{-}}\FloatTok{1.253}\NormalTok{, }\AttributeTok{b\_c =} \FloatTok{2.603}\NormalTok{,}
  \AttributeTok{b\_ea =} \FloatTok{0.790}\NormalTok{, }\AttributeTok{b\_ec =} \SpecialCharTok{{-}}\FloatTok{1.393}\NormalTok{,}
  \AttributeTok{sd\_u0s =} \FloatTok{0.5}\NormalTok{, }\AttributeTok{sd\_u0i =} \FloatTok{0.5}\NormalTok{, ...)\{}
  \FunctionTok{require}\NormalTok{(dplyr)}
  \FunctionTok{require}\NormalTok{(faux)}
  \CommentTok{\# simulate design}
\NormalTok{  dat }\OtherTok{\textless{}{-}} \FunctionTok{add\_random}\NormalTok{(}\AttributeTok{subject =}\NormalTok{ n\_subjects, }\AttributeTok{item =}\NormalTok{ n\_items) }\SpecialCharTok{|\textgreater{}}
    \FunctionTok{add\_between}\NormalTok{(}\StringTok{"subject"}\NormalTok{, }\AttributeTok{expert =} \FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{), }\AttributeTok{.prob =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.25}\NormalTok{, }\FloatTok{0.75}\NormalTok{)) }\SpecialCharTok{|\textgreater{}}
    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{advice\_present =} \FunctionTok{rbinom}\NormalTok{(}\FunctionTok{n}\NormalTok{(), }\DecValTok{1}\NormalTok{, }\AttributeTok{prob =} \DecValTok{2}\SpecialCharTok{/}\DecValTok{3}\NormalTok{)) }\SpecialCharTok{|\textgreater{}}
    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{advice\_correct =} \FunctionTok{if\_else}\NormalTok{(advice\_present }\SpecialCharTok{==} \DecValTok{1}\NormalTok{L, }
                                    \FunctionTok{rbinom}\NormalTok{(}\FunctionTok{n}\NormalTok{(), }\DecValTok{1}\NormalTok{L, }\AttributeTok{prob =} \FloatTok{0.8}\NormalTok{), }\DecValTok{0}\NormalTok{L)) }\SpecialCharTok{|\textgreater{}}
    \CommentTok{\# add random effects}
    \FunctionTok{add\_ranef}\NormalTok{(}\StringTok{"subject"}\NormalTok{, }\AttributeTok{u0s =}\NormalTok{ sd\_u0s) }\SpecialCharTok{|\textgreater{}}
    \FunctionTok{add\_ranef}\NormalTok{(}\StringTok{"item"}\NormalTok{, }\AttributeTok{u0i =}\NormalTok{ sd\_u0i) }\SpecialCharTok{|\textgreater{}}
    \CommentTok{\# compute dependent variable}
    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{linpred =}\NormalTok{ b\_0 }\SpecialCharTok{+}\NormalTok{ u0i }\SpecialCharTok{+}\NormalTok{ u0s }\SpecialCharTok{+}
\NormalTok{        b\_e }\SpecialCharTok{*}\NormalTok{ expert }\SpecialCharTok{+}\NormalTok{ b\_a }\SpecialCharTok{*}\NormalTok{ advice\_present }\SpecialCharTok{+}\NormalTok{ b\_c }\SpecialCharTok{*}\NormalTok{ advice\_correct }\SpecialCharTok{+}
\NormalTok{        b\_ea }\SpecialCharTok{*}\NormalTok{ expert }\SpecialCharTok{*}\NormalTok{ advice\_present }\SpecialCharTok{+}\NormalTok{ b\_ec }\SpecialCharTok{*}\NormalTok{ expert }\SpecialCharTok{*}\NormalTok{ advice\_correct) }\SpecialCharTok{|\textgreater{}}
    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{y\_prob =} \FunctionTok{plogis}\NormalTok{(linpred)) }\SpecialCharTok{|\textgreater{}}
    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{y\_bin =} \FunctionTok{rbinom}\NormalTok{(}\AttributeTok{n =} \FunctionTok{n}\NormalTok{(), }\AttributeTok{size =} \DecValTok{1}\NormalTok{, }\AttributeTok{prob =}\NormalTok{ y\_prob))}
\NormalTok{  dat}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

In the first four lines of the function definition, we set some default
parameter values (which we will explain in the next section) and load
the packages we use to manipulate and simulate data.\footnote{The faux
  package (\citeproc{ref-R-faux}{DeBruine, 2023}) contains useful
  functions when simulating factorial designs, including random effects.}
In our case study, each subject (\texttt{n\_subjects} in total) is
assumed to respond to each item (i.e., CT scan; \texttt{n\_items} in
total). Thus, the \texttt{add\_random} command creates a fully-crossed
\texttt{data.frame} with \texttt{n\_subjects} \(\times\)
\texttt{n\_items} rows. We add a between-subject effect with the
\texttt{add\_between} command, simulating that about \(25\%\) of
subjects are experts. The next two lines simulate that in
\(\frac{2}{3}\) of trials, subjects will be presented with AI advice,
and if advice is presented, the advice will be correct in about \(80\%\)
of cases (the variable \texttt{advice\_correct} is always 0 when no
advice is presented). Next, we simulate one random effect for each
subject (\texttt{u0s}) and for each item (\texttt{u0i}). As assumed by
standard GLMMs, the \texttt{add\_ranef} function draws the random
effects from a normal distribution with a mean 0 and a standard
deviation specified by the user. With all design variables done, we are
ready to simulate our model equation outlined in the last section. The
linear predictor variable \texttt{linpred} (\(\eta\) in the GLMM model
equations) combines the predictor variables, random effects, and model
parameters as assumed by our model. We then transform the linear
predictor with the inverse-link function to compute \texttt{y\_prob},
the probability that the subject correctly solved the item (in R, the
inverse-logit link is computed with \texttt{plogis} and the logit link
with \texttt{qlogis}). In the final step, we simulate the binary
dependent variable \texttt{y\_bin} (i.e., whether the subject makes an
accurate diagnostic decision for the CT scan) by -- for each trial --
drawing from a Bernoulli distribution with success probability
\texttt{y\_prob}.

\subsection{Step 3: Specify the population
parameters}\label{step-3-specify-the-population-parameters}

\subsubsection{THEORY}\label{theory-2}

In the absence of previous studies with the same design or pilot data,
researchers require strategies on how to specify the population
parameters used in their data simulation. Population parameters are all
model parameters estimated in a GLMM, in particular the regression
coefficients of the fixed effects and the standard deviation of the
random effects (and the correlation between random effects in more
complicated models). In contrast to non-hierarchical linear regression,
common heuristics based on standardized effect sizes are less useful or
not even available for GLMMs. Our strategies to specify population
parameters will require access to domain knowledge from domain experts.
Because most domain knowledge can only be expressed in unstandardized
measurement units of a specific application, we argue that
unstandardized effect sizes are usually preferable over standardized
effect sizes for tailored simulation-based sample size planning. The
basic idea of all strategies is how the data generating process implied
by certain values of population parameters can be quantified or
visualized in an intuitive way that enables a calibration of population
parameters based on the available knowledge of domain experts.

Although we use frequentist model estimation in our tutorial, many
strategies described in this chapter are inspired by research on how to
monitor the plausibility of model assumptions in applied Bayesian
statistics (\citeproc{ref-gelmanBayesianWorkflow2020}{Gelman et al.,
2020}).

\subsubsection{PRACTICE}\label{practice-2}

When introducing the simulation function for our case study, we have
used theoretically plausible values as defaults for all model parameters
(\(\beta_0\), \(\beta_e\), \(\beta_a\), \(\beta_c\), \(\beta_{ea}\),
\(\beta_{ec}\), \(\sigma_S\), and \(\sigma_I\)), but have not talked
about where these numbers came from.

The starting point for all parameter values in our present case study
were based on results from distantly related study designs in the
literature. Additionally, we had repeated discussions with our
affiliated domain experts in radiology to check whether our assumptions
seem plausible.

We now outline our main strategy to determine plausible parameter values
for the fixed effects (\(\beta\) parameters): Unfortunately, the model
parameters in a binomial GLMM are hard to interpret in isolation because
1) the \(\beta\) parameters are connected to the modeled probability via
the non-linear inverse-logit link, and 2) we also have to consider the
random effects. The most simple interpretation, that allows us to ignore
the random effects for now, works by imagining a subject with average
ability (\(u_{0s} = 0\)) responding to an item (i.e., CT scan) with
average difficulty (\(u_{0i} = 0\)). Then the model implied probability
that such a person solves such an item accurately is given by:

\[
\begin{aligned}
P(Y=1|\mathbf{X=x}, \mathbf{U} = \mathbf{0}) = \\
= \text{inverse\_logit}[&\beta_0 + \beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
&\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}]
\end{aligned}
\] In fact, we would only need the full equation if the subject is an
expert and correct advice is presented. In all other experimental
conditions, some terms drop from the equation because they are
multiplied by \(0\). For example, the probability that a non-expert with
average ability solves an item with average difficulty when no advice is
presented, only requires the intercept: \[
\begin{aligned}
P(Y=1| advice\_present = 0, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0) = \\
= \text{inverse\_logit}[\beta_0]
\end{aligned}
\] We can revert this perspective by choosing plausible probability
values based on domain knowledge and deriving the parameter values
implied by these probabilities for each experimental condition.

\begin{table}

{\caption{{Assumed probabilities that an average subject solves an
average item in each experimental condition.}{\label{tbl-probtable}}}
\vspace{-20pt}}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1646}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3291}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.5063}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Experimental condition
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\(P(Y=1 \mid \mathbf{X=x}, \mathbf{U} = \mathbf{0})\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Implied equation
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
no advice, no expert & 0.70 & \(logit(0.70) = \beta_0\) \\
no advice, expert & 0.90 & \(logit(0.90) = \beta_0 + \beta_e\) \\
false advice, no expert & 0.40 & \(logit(0.40) = \beta_0 + \beta_a\) \\
false advice, expert & 0.85 &
\(logit(0.85) = \beta_0 + \beta_e + \beta_{a} + \beta_{ea}\) \\
correct advice, no expert & 0.90 &
\(logit(0.90) = \beta_0 + \beta_a + \beta_c\) \\
correct advice, expert & 0.95 &
\(logit(0.95) = \beta_0 + \beta_e + \beta_a + \beta_c + \beta_{ea} + \beta_{ec}\) \\
\end{longtable}

{\vspace{-20pt}
\noindent \emph{Note.} Implied equations are derived based on the model equations and setting all random intercept terms to 0.}

\end{table}

Table~\ref{tbl-probtable} shows our set of assumptions concerning the
probability that an average subject solves an average item for each
experimental condition, as well as the corresponding equations implied
by the model. The table can be used to compute the implied values for
the \(\beta\) parameters, starting with the first equation and
reinserting the computed \(\beta\) values in all following equations
(\texttt{b\_0} stands for the intercept \(\beta_0\), \texttt{b\_e} for
the slope \(\beta_e\), and so on):

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{b\_0 }\OtherTok{\textless{}{-}} \FunctionTok{qlogis}\NormalTok{(}\FloatTok{0.7}\NormalTok{)}
\NormalTok{b\_e }\OtherTok{\textless{}{-}} \FunctionTok{qlogis}\NormalTok{(}\FloatTok{0.9}\NormalTok{) }\SpecialCharTok{{-}}\NormalTok{ b\_0}
\NormalTok{b\_a }\OtherTok{\textless{}{-}} \FunctionTok{qlogis}\NormalTok{(}\FloatTok{0.4}\NormalTok{) }\SpecialCharTok{{-}}\NormalTok{ b\_0}
\NormalTok{b\_ea }\OtherTok{\textless{}{-}} \FunctionTok{qlogis}\NormalTok{(}\FloatTok{0.85}\NormalTok{) }\SpecialCharTok{{-}}\NormalTok{ b\_0 }\SpecialCharTok{{-}}\NormalTok{ b\_e }\SpecialCharTok{{-}}\NormalTok{ b\_a}
\NormalTok{b\_c }\OtherTok{\textless{}{-}} \FunctionTok{qlogis}\NormalTok{(}\FloatTok{0.9}\NormalTok{) }\SpecialCharTok{{-}}\NormalTok{ b\_0 }\SpecialCharTok{{-}}\NormalTok{ b\_a}
\NormalTok{b\_ec }\OtherTok{\textless{}{-}} \FunctionTok{qlogis}\NormalTok{(}\FloatTok{0.95}\NormalTok{) }\SpecialCharTok{{-}}\NormalTok{ b\_0 }\SpecialCharTok{{-}}\NormalTok{ b\_e }\SpecialCharTok{{-}}\NormalTok{ b\_a }\SpecialCharTok{{-}}\NormalTok{ b\_c }\SpecialCharTok{{-}}\NormalTok{ b\_ea}
\FunctionTok{c}\NormalTok{(}\AttributeTok{b\_0 =}\NormalTok{ b\_0, }\AttributeTok{b\_e =}\NormalTok{ b\_e, }\AttributeTok{b\_a =}\NormalTok{ b\_a, }\AttributeTok{b\_c =}\NormalTok{ b\_c, }\AttributeTok{b\_ea =}\NormalTok{ b\_ea, }\AttributeTok{b\_ec =}\NormalTok{ b\_ec)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
       b_0        b_e        b_a        b_c       b_ea       b_ec 
 0.8472979  1.3499267 -1.2527630  2.6026897  0.7901394 -1.3928518 
\end{verbatim}

It is always possible to double-check these computations by transforming
the parameter values back to probabilities, e.g.~ \[
\begin{aligned}
P(Y=1|expert = 1, advice\_present = 1, advice\_correct = 1, u_{0s} = 0, u_{0i} = 0) = \\
= inverse\_logit[\beta_0 + \beta_e + \beta_a + \beta_c + \beta_{ea} + \beta_{ec}]
\end{aligned}
\]

which we compute in R as:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{plogis}\NormalTok{(b\_0 }\SpecialCharTok{+}\NormalTok{ b\_e }\SpecialCharTok{+}\NormalTok{ b\_a }\SpecialCharTok{+}\NormalTok{ b\_c }\SpecialCharTok{+}\NormalTok{ b\_ea }\SpecialCharTok{+}\NormalTok{ b\_ec)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
[1] 0.95
\end{verbatim}

This leaves us with the question on how to determine plausible values
for the two remaining model parameters (\(\sigma_S\), and \(\sigma_I\))
that are the standard deviations for the random intercepts. For this, we
introduce two more strategies.

\subsubsection{Insightful descriptive
statistics}\label{insightful-descriptive-statistics}

\paragraph{THEORY.}\label{theory-3}

The mathematical structure of GLMMs determines which patterns in data
would be produced by the model, if a specific set of values for the
population parameters would be specified. The knowledge of how to
simulate from a GLMM enables us to compute insightful descriptive
statistics that can be compared to available domain knowledge much more
easily than the opaque values of model parameters. For example, domain
experts might not be able to directly choose plausible values for the
coefficients in a logistic regression model (which are measured on the
log-odds scale). However, they should be able to reason about the
expected ratio of the binary dependent variable in different
experimental conditions, i.e.~which relative frequency they expect to
observe (on average). The job of the analyst who is familiar with the
mathematical structure of the GLMM is to produce the model implied value
of the insightful descriptive statistic expected by the domain expert.
Although insightful descriptive statistics usually depend on the model
parameters in a non-linear way, it is not necessary to solve the exact
relationship mathematically. Instead, one can simply adjust the
population parameters by trial and error until the model implied
quantities produce the desired result.

\paragraph{PRACTICE.}\label{practice-3}

In the last section, we showed how we can derive the model implied
probability that a subject with average ability solves an item with
average difficulty for each experimental condition. Although these
derivations are straightforward, it is important not to misinterpret
their implications: In binomial GLMMs, the average probability to solve
an item (averaged across persons of varying ability and items of varying
difficulty) is \textbf{not} equal to the probability that a person with
average ability solves an item with average difficulty
(\citeproc{ref-fahrmeirRegressionModelsMethods2021}{Fahrmeir et al.,
2021}). The first perspective implies a so-called marginal
interpretation, while the second one implies a conditional
interpretation.

For example, we determined the \(\beta\) parameters in a way that
corresponds to a desired conditional probability of \(0.95\), that an
expert with average ability solves an item with average difficulty when
presented with correct advice (the conditional perspective). However,
even if the model were true, we would not observe that 95\% of experts
responding to items presented with correct advice from a big sample of
subjects drawn from their natural distribution of ability, and items
drawn from their natural distribution of difficulty (the marginal
perspective). How much the two probabilities differ depends on the
standard deviations of the random intercepts (the two probabilities are
only equal if both standard deviations would be zero). We want to use
the model implied observed proportion of correct diagnoses in each
experimental condition as an insightful descriptive statistics to
determine plausible values for the random effect standard deviations.

We will simulate a large dataset (for which the observed values of the
descriptive statistic will be close to their model implied true values)
and simply compute the relative frequency of correct diagnoses for each
experimental condition.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(tidyverse)}
\FunctionTok{set.seed}\NormalTok{(}\DecValTok{1}\NormalTok{)}
\NormalTok{dat }\OtherTok{\textless{}{-}} \FunctionTok{simulate}\NormalTok{(}\AttributeTok{n\_subjects =} \DecValTok{3000}\NormalTok{, }\AttributeTok{n\_items =} \DecValTok{3000}\NormalTok{,}
  \AttributeTok{sd\_u0s =} \FloatTok{0.5}\NormalTok{, }\AttributeTok{sd\_u0i =} \FloatTok{0.5}\NormalTok{)}
\NormalTok{dat }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{condition =} \FunctionTok{fct\_cross}\NormalTok{(}
    \FunctionTok{factor}\NormalTok{(expert), }\FunctionTok{factor}\NormalTok{(advice\_present), }\FunctionTok{factor}\NormalTok{(advice\_correct))) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{condition =} \FunctionTok{fct\_recode}\NormalTok{(condition,}
 \StringTok{"no expert, no advice"} \OtherTok{=} \StringTok{"0:0:0"}\NormalTok{, }\StringTok{"expert, no advice"} \OtherTok{=} \StringTok{"1:0:0"}\NormalTok{, }
 \StringTok{"no expert, wrong advice"} \OtherTok{=} \StringTok{"0:1:0"}\NormalTok{, }\StringTok{"expert, wrong advice"} \OtherTok{=} \StringTok{"1:1:0"}\NormalTok{,}
 \StringTok{"no expert, correct advice"} \OtherTok{=} \StringTok{"0:1:1"}\NormalTok{, }\StringTok{"expert, correct advice"} \OtherTok{=} \StringTok{"1:1:1"}\NormalTok{)) }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{group\_by}\NormalTok{(condition) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{summarize}\NormalTok{(}\AttributeTok{relative\_frequency =} \FunctionTok{sum}\NormalTok{(y\_bin) }\SpecialCharTok{/} \FunctionTok{n}\NormalTok{())}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
# A tibble: 6 x 2
  condition                 relative_frequency
  <fct>                                  <dbl>
1 no expert, no advice                   0.683
2 expert, no advice                      0.881
3 no expert, wrong advice                0.409
4 expert, wrong advice                   0.828
5 no expert, correct advice              0.883
6 expert, correct advice                 0.938
\end{verbatim}

We tried using these descriptive statistics to judge together with
domain experts whether our chosen values for the random effect standard
deviations would produce data that align with out domain expertise.
However, although the result was deemed plausible, these statistics were
not informative enough to determine a final set of plausible parameter
values. For this reason, we will additionally look at insightful model
based quantities.

\subsubsection{Insightful model based
quantities}\label{insightful-model-based-quantities}

\paragraph{THEORY.}\label{theory-4}

Because GLMMs are complicated models, descriptive statistics alone are
usually not enough to specify plausible values for all model parameters.
This is especially true for the standard deviation of random effects
that have non-linear (and often unexpected) effects on the model-implied
results. An important advantage of data simulation (where one has full
control over parameter values and sample sizes) is that one can produce
insightful model based quantities that can never be directly observed in
an actual empirical dataset. For example, in a logistic model with
random intercepts for participants, one can produce a visualization of
the implied distribution of the probability that a participant on
average solves a cognitive task. Although domain knowledge will probably
not suffice to specify this distribution completely, it should be
possible to rule out implausible boundary conditions. For example, the
domain expert might deem it implausible that the 5\% most able
participants have a probability of more than 0.99 to solve the difficult
cognitive task.

\paragraph{PRACTICE.}\label{practice-4}

The discussed inequality of conditional and marginal effects in GLMMs
(\citeproc{ref-fahrmeirRegressionModelsMethods2021}{Fahrmeir et al.,
2021}) makes their interpretation more difficult. One must be careful
when specifying parameter values based on previous studies or pilot data
that use the marginal interpretation (e.g., a pilot study providing an
estimate of how often radiologists make an accurate diagnosis based on
brain scans). However, this does not mean that we cannot use the
marginal interpretation (average probability across persons and items)
to inform plausible parameter values: When parameter values have been
selected, we can compute the implied marginal distributions and compare
this information to our domain knowledge. Then, we can iteratively
adjust the parameter values until we are satisfied with the implied
distributions. In the last section, we simulated a large dataset and
computed descriptive statistics, the relative frequencies of correct
diagnoses, for each experimental condition. We will now use the model
implied probability of each simulated data point (stored in the variable
\texttt{y\_prob}) to visualize the whole model implied marginal
distribution of correct diagnoses for each experimental condition.

\phantomsection\label{cell-fig-margdist1}
\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(ggdist)}
\NormalTok{dat }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{condition =} \FunctionTok{fct\_cross}\NormalTok{(}
    \FunctionTok{factor}\NormalTok{(expert), }\FunctionTok{factor}\NormalTok{(advice\_present), }\FunctionTok{factor}\NormalTok{(advice\_correct))) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{condition =} \FunctionTok{fct\_recode}\NormalTok{(condition,}
\StringTok{"no expert, no advice"} \OtherTok{=} \StringTok{"0:0:0"}\NormalTok{, }\StringTok{"expert, no advice"} \OtherTok{=} \StringTok{"1:0:0"}\NormalTok{, }
\StringTok{"no expert, wrong advice"} \OtherTok{=} \StringTok{"0:1:0"}\NormalTok{, }\StringTok{"expert, wrong advice"} \OtherTok{=} \StringTok{"1:1:0"}\NormalTok{,}
\StringTok{"no expert, correct advice"} \OtherTok{=} \StringTok{"0:1:1"}\NormalTok{, }\StringTok{"expert, correct advice"} \OtherTok{=} \StringTok{"1:1:1"}\NormalTok{)) }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ y\_prob, }\AttributeTok{y =}\NormalTok{ condition)) }\SpecialCharTok{+}
  \FunctionTok{stat\_histinterval}\NormalTok{(}\AttributeTok{point\_interval =} \StringTok{"mean\_qi"}\NormalTok{, }\AttributeTok{slab\_color =} \StringTok{"gray45"}\NormalTok{,}
    \AttributeTok{breaks =} \StringTok{"Sturges"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{scale\_x\_continuous}\NormalTok{(}\AttributeTok{breaks =} \FunctionTok{seq}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{, }\FloatTok{0.1}\NormalTok{), }\AttributeTok{limits =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{figure}[H]

\caption{\label{fig-margdist1}Marginal distributions including means,
66\% and 95\% confidence intervals for all experimental conditions.}

\centering{

\includegraphics{manuscript_files/figure-pdf/fig-margdist1-1.pdf}

}

\end{figure}%

Figure~\ref{fig-margdist1} shows the model implied marginal
distributions, including the mean, 66\% and 95\% intervals. We can see
that, indeed, the average probabilities (black dots) slightly differ
from the probabilities of average subjects and items considered in the
previous section. This difference increases with the variability of the
random effects. We can use plots like the one above as a useful tool to
decide whether the specified standard deviations of the subject and item
random intercepts (\(\sigma_S\) and \(\sigma_I\)) are reasonable by
comparing the ranges and overlap between conditions to domain knowledge.

\phantomsection\label{cell-fig-margdist2}
\begin{figure}[H]

\caption{\label{fig-margdist2}Marginal distributions including means,
66\% and 95\% confidence intervals for all experimental conditions while
setting the standard deviation of item random intercepts to 0.01.}

\centering{

\includegraphics{manuscript_files/figure-pdf/fig-margdist2-1.pdf}

}

\end{figure}%

In the next plot, we have set the item standard deviation to almost zero
(\(\sigma_I = 0.01\)). This gives us a better way to see the variability
between persons. As an example, Figure~\ref{fig-margdist2} reveals a
number of implicit assumptions about the comparison between experts and
non-experts: With wrong advice, virtually all experts have a higher
probability of making a correct diagnosis compared to non-experts when
considering only items with average difficulty. In contrast, there is
considerable overlap in probability between experts and non-experts with
no advice and even higher overlap with correct advice. Patterns like
these should be considered carefully and discussed with the domain
experts. Parameter values (\(\beta\) parameters, and \(\sigma_S\))
should be adjusted if the implications do not seem reasonable. We could
also have a closer look at variability between items by setting the
subject standard deviation to almost zero (\(\sigma_S = 0.01\)).

The final plot demonstrates that these plots are also useful for
spotting standard deviations that are specified too high. For
Figure~\ref{fig-margdist3} we have set \(\sigma_S = 3\) and
\(\sigma_I = 3\). This implies that in each experimental condition, the
probabilities that a subject solves an item are usually close to either
0 or 1, which is not a plausible assumption. These high standard
deviations do not account for the inherent variability and complexity of
human performance. For example, we would expect that a participant with
low ability compared to other task experts to solve a difficult item
with a probability substantially larger than zero even when presented
with wrong advice.

\phantomsection\label{cell-fig-margdist3}
\begin{figure}[H]

\caption{\label{fig-margdist3}Marginal distributions including means,
66\% and 95\% confidence intervals for all experimental conditions while
setting the standard deviation of subject and item random intercepts to
3.}

\centering{

\includegraphics{manuscript_files/figure-pdf/fig-margdist3-1.pdf}

}

\end{figure}%

\subsubsection{Iterative process with domain
experts}\label{iterative-process-with-domain-experts}

\paragraph{THEORY.}\label{theory-5}

In our experience, the gathering of domain knowledge by domain experts
and the consecutive specification of population parameter values used in
data simulation is not a one-time event but rather an iterative process.
In a first step, domain experts can be interviewed to ``elicit'' their
domain knowledge about how the future data of the planned study is
expected to look like. As most domain experts are no experts in
statistical modeling and GLMMs, they often struggle without further
guidance to communicate their knowledge in a way that is useful when
specifying the parameters for data simulation. For this reason, we
suggest that after an initial unstructured interview of domain experts,
the analyst who is familiar with the structure of the GLMM under study
selects an initial set of insightful descriptive statistics and model
based quantities. Then they reenter into an iterative discussion where
some set of population values are selected and the plausibility of
resulting implied quantities are discussed with the domain experts.
Then, the population parameters are updated based on this discussion
until the domain experts are satisfied with the final result. During
this process, the monitored model based quantities and descriptive
statistics can be updated or extended to capture as much available
domain knowledge as possible.

\subparagraph{PRACTICE.}\label{practice-5}

All parameter values in our present case study have been determined
based on repeated discussions with our affiliated domain experts in
Radiology to validate our assumptions. Initially, we reviewed the
literature to establish a reasonable baseline performance rate for
examining head CT scans for intracranial hemorrhage. Existing studies
indicate that radiologists typically demonstrate high accuracies, often
exceeding or hovering around 90\%, while interns have been shown to
perform below 80\%, and medical students fall even shorter. For
simplicity, we assumed plausible probability values of .90 for experts
and .70 for non-experts, respectively. Our experts confirmed that these
values are realistic baselines for reviewing diverse head CT images
without AI assistance. Subsequently, we consulted several published
papers investigating the effect of correct and incorrect advice on
decision-making performance in other settings. From their findings, we
inferred that both experts and non-experts should benefit from correct
and suffer losses from incorrect advice. However, the magnitude of these
effects should be substantially greater for non-experts, given their
demonstrated reliance on advice compared to experts. We further
validated the plausibility of our estimated gains and losses with the
collaborating radiologists. For our simulation, we used the
probabilities of average participants to solve an average case, as shown
in Table 1.

\subsection{Step 4: Estimate the statistical
model}\label{step-4-estimate-the-statistical-model}

\subsubsection{THEORY}\label{theory-6}

At this point, the researcher is capable of producing a simulated
dataset similar to the actual dataset that will later be collected in
the planned study. The next step is to specify how the statistical model
shall be estimated in the actual study collected later. This usually
includes the selection of 1) a statistical framework (e.g., frequentist
statistics), 2) a software package that is capable of estimating the
model class of interest (e.g., the lme4 R package), 3) an estimation
algorithm (e.g., the default optimizer ``bobyqa''), and 4) the specific
model structure including all fixed effects, random effects, and the
model family of the dependent variable.\\
Note that this does not always mean that one will specify the same GLMM
that was used when specifying the data generating process. On the one
hand, using a simpler model for data simulation than for model
estimation can be a useful strategy in scenarios where making plausible
assumptions for complicated random effect structures and interactions is
not feasible. On the other hand, using a more complex model for data
simulation than for model estimation can be a useful strategy in
scenarios where one has specific domain knowledge about aspects of the
data generating process that are still difficult to estimate with the
current state-of-the-art in multilevel modeling.

\subsubsection{PRACTICE}\label{practice-6}

In our case study, we use the lme4 R package
(\citeproc{ref-batesFittingLinearMixedEffects2015}{Bates et al., 2015}),
which is a state-of-the-art tool for fitting frequentist
GLMMs.\footnote{For Bayesian GLMMs, the brms R package is currently the
  most prominent option
  (\citeproc{ref-burknerBrmsPackageBayesian2017}{Bürkner, 2017}).} For
the current example, we simulate data according to our model, in which
100 subjects respond to 50 items (we use \texttt{set.seed} to make the
simulation reproducible). However, for the sake of the exercise, we can
imagine that this would be real data resulting from our future
experiment and think about how we would analyze this data.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(tidyverse)}
\FunctionTok{set.seed}\NormalTok{(}\DecValTok{1}\NormalTok{)}
\NormalTok{dat }\OtherTok{\textless{}{-}} \FunctionTok{simulate}\NormalTok{(}\AttributeTok{n\_subjects =} \DecValTok{100}\NormalTok{, }\AttributeTok{n\_items =} \DecValTok{50}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

The lme4 package uses a special syntax for model specification. Our
specific GLMM is represented by the formula:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(lme4)}
\NormalTok{f }\OtherTok{\textless{}{-}}\NormalTok{ y\_bin }\SpecialCharTok{\textasciitilde{}} \DecValTok{1} \SpecialCharTok{+}\NormalTok{ expert }\SpecialCharTok{+}\NormalTok{ advice\_present }\SpecialCharTok{+}\NormalTok{ advice\_correct }\SpecialCharTok{+} 
\NormalTok{  expert}\SpecialCharTok{:}\NormalTok{advice\_present }\SpecialCharTok{+}\NormalTok{ expert}\SpecialCharTok{:}\NormalTok{advice\_correct }\SpecialCharTok{+}
\NormalTok{  (}\DecValTok{1}\SpecialCharTok{|}\NormalTok{subject) }\SpecialCharTok{+}\NormalTok{ (}\DecValTok{1}\SpecialCharTok{|}\NormalTok{item)}
\end{Highlighting}
\end{Shaded}

The first two lines look similar to any linear model in R (general
intercept indicated by \texttt{1}; main effects indicated by variable
names in the dataset; interactions indicated by
\texttt{variable1:variable2}). The third line specifies a random
intercept for each subject \texttt{(1\textbar{}subject)} and for each
item \texttt{(1\textbar{}item)}. The complete set of rules for the
syntax is outlined in Bates et al.
(\citeproc{ref-batesFittingLinearMixedEffects2015}{2015}) and in the
documentation of the lme4 package.

In lme4, a GLMM is fitted with the \texttt{glmer} function. By setting
\texttt{family\ =\ \ "binomial"}, we request a binomial GLMM appropriate
for our binary dependent variable \texttt{y\_bin} (the binomial GLMM
uses the canonical logit link by default), which is defined as an
accurate (1) vs.~inaccurate (0) diagnosis.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{fit }\OtherTok{\textless{}{-}} \FunctionTok{glmer}\NormalTok{(f, }\AttributeTok{data =}\NormalTok{ dat, }\AttributeTok{family =} \StringTok{"binomial"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

We can inspect the estimates for all model parameters with the
\texttt{summary} command:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{summary}\NormalTok{(fit)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
Generalized linear mixed model fit by maximum likelihood (Laplace
  Approximation) [glmerMod]
 Family: binomial  ( logit )
Formula: 
y_bin ~ 1 + expert + advice_present + advice_correct + expert:advice_present +  
    expert:advice_correct + (1 | subject) + (1 | item)
   Data: dat

     AIC      BIC   logLik deviance df.resid 
  4149.4   4201.6  -2066.7   4133.4     4992

Scaled residuals: 
    Min      1Q  Median      3Q     Max 
-5.7669  0.2125  0.3046  0.4317  2.1056

Random effects:
 Groups  Name        Variance Std.Dev.
 subject (Intercept) 0.3148   0.5611  
 item    (Intercept) 0.1624   0.4029  
Number of obs: 5000, groups:  subject, 100; item, 50

Fixed effects:
                      Estimate Std. Error z value Pr(>|z|)    
(Intercept)             1.0339     0.1103   9.374  < 2e-16 ***
expert                  1.1849     0.2096   5.654 1.56e-08 ***
advice_present         -1.3436     0.1206 -11.143  < 2e-16 ***
advice_correct          2.6154     0.1273  20.540  < 2e-16 ***
expert:advice_present   1.0589     0.2940   3.601 0.000317 ***
expert:advice_correct  -1.8104     0.2915  -6.210 5.29e-10 ***
---
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

Correlation of Fixed Effects:
            (Intr) expert advc_p advc_c exprt:dvc_p
expert      -0.377                                 
advic_prsnt -0.349  0.176                          
advic_crrct  0.023  0.001 -0.668                   
exprt:dvc_p  0.143 -0.448 -0.412  0.276            
exprt:dvc_c -0.008  0.004  0.292 -0.435 -0.686     
\end{verbatim}

In the lme4 output, the \texttt{Estimate} column in the
\texttt{Fixed\ effects} table contains the estimates for the \(\beta\)
parameters, while the \texttt{Std.Dev.} column in the
\texttt{Random\ effects} table contains the estimates for \(\sigma_S\)
and \(\sigma_I\).

\subsection{Step 5: Compute the
estimate}\label{step-5-compute-the-estimate}

\subsubsection{THEORY}\label{theory-7}

In previous steps, we have defined the theoretical estimand, written a
data simulation function and specified how to estimate a GLMM using
simulated data. The next step is to specify how to compute a concrete
point estimate of the theoretical estimand within the framework of the
fitted GLMM. For some research question, the estimate corresponds with a
single regression coefficient. In more complicated scenarios, the
estimate is computed from a combination of coefficients. Beyond
computing the point estimate, we have already discussed that both
hypothesis testing and interval estimation can be used to answer the
research question. The decision on testing or estimating is then
followed by selecting the specific statistical method that shall be
applied to compute the HT(s) or CI(s) (e.g., compute HTs and CIs with
the marginaleffects R package using the delta method).

\subsubsection{PRACTICE}\label{practice-7}

In the estimand section, we have translated a verbal description of our
research question into four probability statements that are specified
outside of any specific statistical model. For a concrete estimate
within the context of our specified GLMM, we must compute the following
probability contrasts:

\[
\begin{aligned}
& P(Y=1|advice\_present = 1, advice\_correct = 1, expert = 1, u_{0s} = 0, u_{0i} = 0) \\
& \quad - P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 1, u_{0s} = 0, u_{0i} = 0)
\end{aligned}
\] \[
\begin{aligned}
& P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 1, u_{0s} = 0, u_{0i} = 0) \\
& \quad - P(Y=1|advice\_present = 1, advice\_correct = 0, expert = 1, u_{0s} = 0, u_{0i} = 0)
\end{aligned}
\]

\[
\begin{aligned}
& P(Y=1|advice\_present = 1, advice\_correct = 1, expert = 0, u_{0s} = 0, u_{0i} = 0) \\
& \quad - P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0)
\end{aligned}
\]

\[
\begin{aligned}
& P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0) \\
& \quad - P(Y=1|advice\_present = 1, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0)
\end{aligned}
\] We have already discussed how to compute the involved probabilities
in the section on specifying population parameters. Plugging in the
model equation of the GLMM produces an equation on how to compute each
contrast if all model parameters were known. When we want to estimate
the above contrasts based on \emph{observed} data, the only difference
is that model parameters are not known and we instead use the
corresponding parameter \emph{estimates}.

We could use our knowledge of the structure of the GLMM to determine the
exact formula needed to compute the contrasts of interest and then plug
in the parameter estimates manually from the \texttt{summary(fit)}
output. However, this would be tedious and we can use R to compute this
contrast without doing the math. Using the first contrast \emph{(correct
advice, expert) - (no advice, expert)} as our example, we could apply
the \texttt{predict} function of the lme4 package to compute the
predicted probability for a correct diagnosis based on our fitted model,
plug in the two sets of predictor values, and compute the difference
between the two estimated probabilities.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{grid1 }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{advice\_present =} \FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{), }\AttributeTok{advice\_correct =} \FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{), }
  \AttributeTok{expert =} \FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{1}\NormalTok{))}
\NormalTok{grid1}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
  advice_present advice_correct expert
1              1              1      1
2              0              0      1
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{pred }\OtherTok{\textless{}{-}} \FunctionTok{predict}\NormalTok{(fit, }\AttributeTok{newdata =}\NormalTok{ grid1, }\AttributeTok{type =} \StringTok{"response"}\NormalTok{, }\AttributeTok{re.form =} \ConstantTok{NA}\NormalTok{)}
\NormalTok{pred}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
       1        2 
0.939292 0.901923 
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{pred[}\DecValTok{1}\NormalTok{] }\SpecialCharTok{{-}}\NormalTok{ pred[}\DecValTok{2}\NormalTok{]}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
         1 
0.03736901 
\end{verbatim}

The argument \texttt{type\ =\ "response"} specifies that predictions are
made on the probability scale (instead of the log-odds scale of the
\(\beta\) parameters), while \texttt{re.form\ =\ NA} sets all random
effects to 0. We could use this method to compute point estimates for
all four contrasts that are part of our estimand. However, depending on
whether we are interested in hypothesis testing or parameter estimation,
we also need a method to compute HTs or CI. The marginaleffects package
(\citeproc{ref-R-marginaleffects}{Arel-Bundock, 2024}) is a very
flexible, increasingly popular package to compute HTs and CIs for
contrasts with a variety of statistical models, including GLMMs
estimated with lme4. First, we specify a grid of all combinations of
predictor variable and then compute estimated probabilities for all
experimental conditions in our experiment with the \texttt{predictions}
function:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(tidyverse)}
\FunctionTok{library}\NormalTok{(marginaleffects)}
\FunctionTok{library}\NormalTok{(tinytable)}
\NormalTok{grid2 }\OtherTok{\textless{}{-}} \FunctionTok{expand\_grid}\NormalTok{(}\AttributeTok{advice\_present =} \DecValTok{0}\SpecialCharTok{:}\DecValTok{1}\NormalTok{, }
  \AttributeTok{advice\_correct =} \DecValTok{0}\SpecialCharTok{:}\DecValTok{1}\NormalTok{, }\AttributeTok{expert =} \DecValTok{0}\SpecialCharTok{:}\DecValTok{1}\NormalTok{)}
\NormalTok{grid2}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
# A tibble: 8 x 3
  advice_present advice_correct expert
           <int>          <int>  <int>
1              0              0      0
2              0              0      1
3              0              1      0
4              0              1      1
5              1              0      0
6              1              0      1
7              1              1      0
8              1              1      1
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{preds }\OtherTok{\textless{}{-}} \FunctionTok{predictions}\NormalTok{(fit, }\AttributeTok{newdata =}\NormalTok{ grid2, }
  \AttributeTok{type =} \StringTok{"response"}\NormalTok{, }\AttributeTok{re.form =} \ConstantTok{NA}\NormalTok{)}
\FunctionTok{print}\NormalTok{(preds, }\AttributeTok{style =} \StringTok{"tinytable"}\NormalTok{) }\SpecialCharTok{|\textgreater{}} \FunctionTok{theme\_tt}\NormalTok{(}\AttributeTok{theme =} \StringTok{"resize"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}
\centering
\resizebox{\ifdim\width>\linewidth 1\linewidth\else\width\fi}{!}{
\begin{talltblr}[         %% tabularray outer open
entry=none,label=none,
note{}={Type:  response 
},
note{ }={Columns: rowid, estimate, std.error, statistic, p.value, s.value, conf.low, conf.high, advice\_present, advice\_correct, expert, y\_bin 
},
]                     %% tabularray outer close
{                     %% tabularray inner open
colspec={Q[]Q[]Q[]Q[]Q[]Q[]Q[]Q[]Q[]Q[]},
}                     %% tabularray inner close
\toprule
Estimate & Std. Error & z & Pr(>|z|) & S & 2.5 \% & 97.5 \% & advice\_present & advice\_correct & expert \\ \midrule %% TinyTableHeader
0.738 & 0.02134 &  34.6 & <0.001 & 867.2 & 0.696 & 0.779 & 0 & 0 & 0 \\
0.902 & 0.01739 &  51.9 & <0.001 & Inf & 0.868 & 0.936 & 0 & 0 & 1 \\
0.975 & 0.00421 & 231.6 & <0.001 & Inf & 0.966 & 0.983 & 0 & 1 & 0 \\
0.954 & 0.01454 &  65.6 & <0.001 & Inf & 0.925 & 0.982 & 0 & 1 & 1 \\
0.423 & 0.03221 &  13.1 & <0.001 & 128.6 & 0.360 & 0.486 & 1 & 0 & 0 \\
0.874 & 0.02793 &  31.3 & <0.001 & 711.4 & 0.819 & 0.928 & 1 & 0 & 1 \\
0.909 & 0.00967 &  94.1 & <0.001 & Inf & 0.890 & 0.928 & 1 & 1 & 0 \\
0.939 & 0.01091 &  86.1 & <0.001 & Inf & 0.918 & 0.961 & 1 & 1 & 1 \\
\bottomrule
\end{talltblr}
}
\end{table}

The point estimates for all experimental conditions are reported in the
\texttt{Estimate} column. Note that the output also contains the two
missing by design conditions that will never be observed in the actual
study (\(advice\_present = 0, advice\_correct = 1, expert = 1\) and
\(advice\_present = 0, advice\_correct = 1, expert = 0\)). This is no
problem as long as we never interpret those estimates. Next, we use the
estimated probabilities to compute the four specific contrasts that are
part of our estimand. For this we must specify which rows in
\texttt{probs} have to be subtracted from each other. We will use the
\texttt{hypotheses} function to compute our four contrasts of interest
together with HTs and CIs. We use the default inference options of the
marginaleffects package that compute HTs and CIs based on the
approximate delta method.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{contrasts }\OtherTok{\textless{}{-}}\NormalTok{ preds }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{hypotheses}\NormalTok{(}\AttributeTok{hypothesis =} \FunctionTok{c}\NormalTok{(}
    \StringTok{"b8 = b2"}\NormalTok{,  }\CommentTok{\# (correct advice, expert) {-} (no advice, expert)}
    \StringTok{"b2 = b6"}\NormalTok{,  }\CommentTok{\# (no advice, expert) {-} (wrong advice, expert) }
    \StringTok{"b7 = b1"}\NormalTok{,  }\CommentTok{\# (correct advice, no expert) {-} (no advice, no expert)}
    \StringTok{"b1 = b5"}\NormalTok{), }\CommentTok{\# (no advice, no expert) {-} (wrong advice, no expert)}
    \AttributeTok{equivalence =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{))}
\FunctionTok{print}\NormalTok{(contrasts, }\AttributeTok{style =} \StringTok{"tinytable"}\NormalTok{) }\SpecialCharTok{|\textgreater{}} \FunctionTok{theme\_tt}\NormalTok{(}\AttributeTok{theme =} \StringTok{"resize"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}
\centering
\resizebox{\ifdim\width>\linewidth 1\linewidth\else\width\fi}{!}{
\begin{talltblr}[         %% tabularray outer open
entry=none,label=none,
note{}={Type:  response 
},
note{ }={Columns: term, estimate, std.error, statistic, p.value, s.value, conf.low, conf.high, statistic.noninf, statistic.nonsup, p.value.noninf, p.value.nonsup, p.value.equiv 
},
]                     %% tabularray outer close
{                     %% tabularray inner open
colspec={Q[]Q[]Q[]Q[]Q[]Q[]Q[]Q[]Q[]Q[]Q[]},
}                     %% tabularray inner close
\toprule
Term & Estimate & Std. Error & z & Pr(>|z|) & S & 2.5 \% & 97.5 \% & p (NonSup) & p (NonInf) & p (Equiv) \\ \midrule %% TinyTableHeader
b8=b2 & 0.0374 & 0.0162 &  2.31 & 0.021 & 5.6 &  0.00563 & 0.0691 & 0.989 & 0.0105 & 0.989 \\
b2=b6 & 0.0282 & 0.0279 &  1.01 & 0.312 & 1.7 & -0.02653 & 0.0830 & 0.844 & 0.1562 & 0.844 \\
b7=b1 & 0.1717 & 0.0173 &  9.93 & <0.001 & 74.8 &  0.13780 & 0.2056 & 1.000 & <0.001 & 1.000 \\
b1=b5 & 0.3145 & 0.0280 & 11.24 & <0.001 & 95.0 &  0.25965 & 0.3693 & 1.000 & <0.001 & 1.000 \\
\bottomrule
\end{talltblr}
}
\end{table}

The expression \texttt{"b8\ =\ b2"} is special syntax to subtract the
estimate in row number 8 from the estimate in row number 2 in the
\texttt{probs}-output. The argument \texttt{equivalence\ =\ c(0,\ 0)}
can be used to compute one-sided p-values, testing whether the contrast
in the population is smaller than 0 (\texttt{p\ (NonSub)} column) or
greater than 0 (\texttt{p\ (NonInf)} column). The point estimates for
four contrasts are reported in the \texttt{Estimate} column. Note that
to facilitate interpretation, we arranged the contrasts in a way that we
theoretically expect positive values for all four of them.

\paragraph{Hypothesis testing.}\label{hypothesis-testing}

If we chose hypothesis testing for our case study, we would test a
combined null hypothesis \(H_0\) that consists of four separate null
hypotheses: \[
\begin{aligned}
H_{01}:\ & P(Y=1|advice\_present = 1, advice\_correct = 1, expert = 1, u_{0s} = 0, u_{0i} = 0) \leq \\
& P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 1, u_{0s} = 0, u_{0i} = 0) \\
H_{02}:\ &P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 1, u_{0s} = 0, u_{0i} = 0) \leq \\
& P(Y=1|advice\_present = 1, advice\_correct = 0, expert = 1, u_{0s} = 0, u_{0i} = 0) \\
H_{03}:\ &P(Y=1|advice\_present = 1, advice\_correct = 1, expert = 0, u_{0s} = 0, u_{0i} = 0) \leq \\
& P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0) \\
H_{04}:\ & P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0) \leq \\
& P(Y=1|advice\_present = 1, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0)
\end{aligned}
\] The combined null hypothesis \(H_0\) should only be rejected if
\textbf{all} individual null hypotheses are rejected (i.e.,
intersection-union setting;
\citeproc{ref-dmitrienkoTraditionalMultiplicityAdjustment2013}{Dmitrienko
\& D'Agostino, 2013}). In such cases, the error probabilities do not
accumulate, and we would waste power when correcting for multiple tests.

With a standard significance level of \(\alpha = 0.05\), we would not
reject all four null hypotheses (the p-value in the \texttt{p\ (NonInf)}
column for the second hypothesis is not significant) and therefore also
not reject the combined null hypothesis for this particular (simulated)
dataset. Note that this decision would be wrong because we have
simulated the data such that the combined alternative hypothesis \(H_1\)
is actually true in the population.

\paragraph{Interval estimation.}\label{interval-estimation}

If we chose parameter estimation for our case study, we would focus on
the two-sided CIs of the four contrasts of interest. With a standard
confidence level of \(1 - \alpha = 0.95\), plausible values are clearly
in the positive range for the first, third and fourth contrast, while
both negative and positive values seem plausible for the second
contrast. Note that due to the constrained range of the probability
scale, the width of the CI differs between the four contrasts (which is
the expected behavior for binomial GLMMs). The smallest width is
observed for the first contrast (expert with correct advice vs.~expert
without advice) where both underlying probabilities are close to 1. The
largest width is observed for the fourth contrast (non-expert with wrong
advice vs.~non-expert without advice), where both underlying
probabilities are closer to 0.5.

\subsection{Step 6: Perform repeated
simulations}\label{step-6-perform-repeated-simulations}

\subsubsection{THEORY}\label{theory-8}

Conducting all previous steps enables the analyst to 1) simulate a
dataset, 2) estimate a GLMM, and 3) compute HTs or CIs for estimands of
interest, mirroring the analysis that will later be performed for the
actual dataset of the planned study. The last missing piece is to write
code to perform the above steps repeatedly and allow for a setting using
different sample sizes. On a conceptual level, we first require a
function that takes as input the sample size and the full set of
population parameter values. When planning for power, the function
should return the p-value(s) of the HT(s) of interest when conducted on
the simulated dataset. When planning for precision, the function should
return the width of the CI(s) of interest. Secondly, we must run this
function repeatedly with the same sample size and population parameters.
Because even fitting GLMMs with frequentist methods can quickly become
time-consuming, it is recommended to use parallel computing, that is
running simulations on multiple cores of the computer at the same time
to reduce total run time. Thirdly, the results of the repeated
simulation must be collected and aggregated. When planning for power, we
compute the relative frequency of (a) significant p-value(s) across
repeated simulations. When planning for precision, we compute the
average width of the CI(s). Lastly, we have to repeat the complete
simulation for different sample sizes, to determine how big the sample
must be in order to achieve the targeted power or precision.

\subsubsection{PRACTICE}\label{practice-8}

We are finally ready to run our simulation-based sample size planning
analyses to plan for power and for precision. Wrapping the
\texttt{simulate} function already constructed earlier, the helper
function \texttt{sim\_and\_analyse} performs all previous steps
(simulate a dataset, fit a GLMM, compute p-values and CIs) in a single
command.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{sim\_and\_analyse }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(}
  \AttributeTok{formula\_chr =} \StringTok{"y\_bin \textasciitilde{} 1 + expert + advice\_present + advice\_correct + }
\StringTok{    expert:advice\_present + expert:advice\_correct + (1|subject) + (1|item)"}\NormalTok{,}
  \AttributeTok{contrasts =} \FunctionTok{c}\NormalTok{(}\StringTok{"b8 = b2"}\NormalTok{, }\StringTok{"b2 = b6"}\NormalTok{, }\StringTok{"b7 = b1"}\NormalTok{, }\StringTok{"b1 = b5"}\NormalTok{), ...)\{}
  \FunctionTok{require}\NormalTok{(lme4)}
  \FunctionTok{require}\NormalTok{(marginaleffects)}
  \FunctionTok{require}\NormalTok{(tidyr)}
  \CommentTok{\# simulate data}
\NormalTok{  dat }\OtherTok{\textless{}{-}} \FunctionTok{simulate}\NormalTok{(...)}
  \CommentTok{\# fit model}
\NormalTok{  model }\OtherTok{\textless{}{-}} \FunctionTok{glmer}\NormalTok{(}\FunctionTok{as.formula}\NormalTok{(formula\_chr), }\AttributeTok{data =}\NormalTok{ dat, }\AttributeTok{family =} \StringTok{"binomial"}\NormalTok{)}
  \CommentTok{\# compute contrasts}
\NormalTok{  contr\_df }\OtherTok{\textless{}{-}} \FunctionTok{expand\_grid}\NormalTok{(}\AttributeTok{advice\_present =} \DecValTok{0}\SpecialCharTok{:}\DecValTok{1}\NormalTok{, }\AttributeTok{advice\_correct =} \DecValTok{0}\SpecialCharTok{:}\DecValTok{1}\NormalTok{,}
    \AttributeTok{expert =} \DecValTok{0}\SpecialCharTok{:}\DecValTok{1}\NormalTok{)}
  \FunctionTok{predictions}\NormalTok{(model, }\AttributeTok{newdata =}\NormalTok{ contr\_df, }\AttributeTok{type =} \StringTok{"response"}\NormalTok{, }\AttributeTok{re.form =} \ConstantTok{NA}\NormalTok{) }\SpecialCharTok{|\textgreater{}}
    \FunctionTok{hypotheses}\NormalTok{(}\AttributeTok{hypothesis =}\NormalTok{ contrasts, }\AttributeTok{equivalence =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{)) }\SpecialCharTok{|\textgreater{}}
    \FunctionTok{data.frame}\NormalTok{()}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

Simulation-based sample size planning can quickly become computationally
intensive when we repeatedly simulate data and fit models for different
parameter combinations or sample sizes. Thus, we use the future
(\citeproc{ref-R-RJ-2021-048}{Bengtsson, 2021}) and furrr
(\citeproc{ref-R-furrr}{Vaughan \& Dancho, 2022}) packages to perform
computations in parallel. First, we enable parallelization with the
\texttt{plan} function and specify how many parallel cores (``workers'')
of our computer to use (users can find out the maximum number of cores
on their computer with the command \texttt{parallel::detectCores()}),
and set a seed to make the simulation reproducible.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(future)}
\FunctionTok{plan}\NormalTok{(}\StringTok{"multisession"}\NormalTok{, }\AttributeTok{workers =} \DecValTok{6}\NormalTok{)}
\FunctionTok{set.seed}\NormalTok{(}\DecValTok{2}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

The next code chunk specifies a simulation grid with different settings
for both the number of subjects (\texttt{n\_subjects}) and the number of
items (\texttt{n\_items}), each combination being repeated \texttt{rep}
times. We chose 300 repetitions for the data simulation at hand as it
strikes a balance between achieving a robust statistical estimate and
remaining computationally feasible. With the current settings, this
simulation takes about one hour on a MacBook Pro from 2020 with M1 chip
and 16 GB working memory. If you want to quickly experiment with the
code yourself, a setting with \texttt{workers\ =\ 4} and
\texttt{rep\ =\ 5} should finish in less than 5 minutes, even on smaller
machines.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(furrr)}
\NormalTok{sim\_result }\OtherTok{\textless{}{-}} \FunctionTok{crossing}\NormalTok{(}
  \AttributeTok{rep =} \DecValTok{1}\SpecialCharTok{:}\DecValTok{300}\NormalTok{,}
  \AttributeTok{n\_subjects =} \FunctionTok{c}\NormalTok{(}\DecValTok{100}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{250}\NormalTok{),}
  \AttributeTok{n\_items =} \FunctionTok{c}\NormalTok{(}\DecValTok{10}\NormalTok{, }\DecValTok{30}\NormalTok{, }\DecValTok{50}\NormalTok{, }\DecValTok{70}\NormalTok{)}
\NormalTok{) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{res =} \FunctionTok{future\_pmap}\NormalTok{(., sim\_and\_analyse, }
    \AttributeTok{.options =} \FunctionTok{furrr\_options}\NormalTok{(}\AttributeTok{seed =} \ConstantTok{TRUE}\NormalTok{))) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{unnest}\NormalTok{(}\AttributeTok{col =}\NormalTok{ res)}
\end{Highlighting}
\end{Shaded}

The result of this computation is a data frame that contains the
p-values and CIs of all specified contrasts for each simulated dataset.
In some iterations (predominantly in conditions with small sample
sizes), model estimation did not converge with the lme4 package. When
the model fails to converge, it means that the statistical model being
fitted to the data failed to reach a stable or valid solution during the
estimation process. We do not remove these results because
non-convergence can also happen when analyzing the real data we plan to
collect, thus, we want to factor in this possibility to keep our
simulation more realistic.

\paragraph{Power results.}\label{power-results}

For our exemplary combined hypothesis, power is defined as the
(long-run) percentage of simulations in which all four p-values of our
individual hypotheses are significant at the \(\alpha = 0.05\) level.
Based on our simulation outcomes, we compute a power estimate for each
combination of \texttt{n\_subjects} \(\times\) \texttt{n\_items}
(including 95\% CIs) and visualize the results with the following
code.\footnote{This code was inspired by the ``Mixed Design Simulation''
  vignette of the faux package at
  \url{https://debruine.github.io/faux/articles/sim_mixed.html}.}

\phantomsection\label{cell-fig-finalpwr}
\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(binom)}
\NormalTok{alpha }\OtherTok{\textless{}{-}} \FloatTok{0.05}
\NormalTok{power }\OtherTok{\textless{}{-}}\NormalTok{ sim\_result }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{pivot\_wider}\NormalTok{(}\AttributeTok{names\_from =}\NormalTok{ term, }\AttributeTok{names\_sep =} \StringTok{"\_"}\NormalTok{, }
    \AttributeTok{values\_from =}\NormalTok{ estimate}\SpecialCharTok{:}\NormalTok{p.value.equiv) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{group\_by}\NormalTok{(n\_subjects, n\_items) }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{summarise}\NormalTok{(}
    \AttributeTok{power =} \FunctionTok{mean}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{p.value.noninf\_b1=b5}\StringTok{\textasciigrave{}} \SpecialCharTok{\textless{}}\NormalTok{ alpha }\SpecialCharTok{\&} 
        \StringTok{\textasciigrave{}}\AttributeTok{p.value.noninf\_b8=b2}\StringTok{\textasciigrave{}} \SpecialCharTok{\textless{}}\NormalTok{ alpha }\SpecialCharTok{\&} \StringTok{\textasciigrave{}}\AttributeTok{p.value.noninf\_b2=b6}\StringTok{\textasciigrave{}} \SpecialCharTok{\textless{}}\NormalTok{ alpha }\SpecialCharTok{\&} 
        \StringTok{\textasciigrave{}}\AttributeTok{p.value.noninf\_b7=b1}\StringTok{\textasciigrave{}} \SpecialCharTok{\textless{}}\NormalTok{ alpha), }
    \AttributeTok{n\_sig =} \FunctionTok{sum}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{p.value.noninf\_b1=b5}\StringTok{\textasciigrave{}} \SpecialCharTok{\textless{}}\NormalTok{ alpha }\SpecialCharTok{\&} 
        \StringTok{\textasciigrave{}}\AttributeTok{p.value.noninf\_b8=b2}\StringTok{\textasciigrave{}} \SpecialCharTok{\textless{}}\NormalTok{ alpha }\SpecialCharTok{\&} \StringTok{\textasciigrave{}}\AttributeTok{p.value.noninf\_b2=b6}\StringTok{\textasciigrave{}} \SpecialCharTok{\textless{}}\NormalTok{ alpha }\SpecialCharTok{\&} 
        \StringTok{\textasciigrave{}}\AttributeTok{p.value.noninf\_b7=b1}\StringTok{\textasciigrave{}} \SpecialCharTok{\textless{}}\NormalTok{ alpha),}
    \AttributeTok{n =} \FunctionTok{n}\NormalTok{(),}
    \AttributeTok{ci.lwr =} \FunctionTok{binom.confint}\NormalTok{(n\_sig, n, }\AttributeTok{method =} \StringTok{"wilson"}\NormalTok{)}\SpecialCharTok{$}\NormalTok{lower,}
    \AttributeTok{ci.upr =} \FunctionTok{binom.confint}\NormalTok{(n\_sig, n, }\AttributeTok{method =} \StringTok{"wilson"}\NormalTok{)}\SpecialCharTok{$}\NormalTok{upper, }
    \AttributeTok{.groups =} \StringTok{"drop"}\NormalTok{)}
\NormalTok{power }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{mutate}\NormalTok{(}\FunctionTok{across}\NormalTok{(}\FunctionTok{c}\NormalTok{(n\_subjects, n\_items), factor)) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(n\_subjects, n\_items, }\AttributeTok{fill =}\NormalTok{ power)) }\SpecialCharTok{+}
  \FunctionTok{geom\_tile}\NormalTok{() }\SpecialCharTok{+}
  \FunctionTok{geom\_text}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{label =} \FunctionTok{sprintf}\NormalTok{(}\StringTok{"\%.2f }\SpecialCharTok{\textbackslash{}n}\StringTok{ [\%.2f; \%.2f]"}\NormalTok{, }
\NormalTok{                                power, ci.lwr, ci.upr)), }
    \AttributeTok{color =} \StringTok{"white"}\NormalTok{, }\AttributeTok{size =} \DecValTok{4}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{scale\_fill\_viridis\_c}\NormalTok{(}\AttributeTok{limits =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{)) }\SpecialCharTok{+}
  \FunctionTok{xlab}\NormalTok{(}\StringTok{"number of subjects"}\NormalTok{) }\SpecialCharTok{+} \FunctionTok{ylab}\NormalTok{(}\StringTok{"number of items"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[H]

\caption{\label{fig-finalpwr}Simulation-based power estimates including
95\% confidence interval of the case study for different numbers of
subjects and items, based on a significance level of 0.05.}

\centering{

\includegraphics{manuscript_files/figure-pdf/fig-finalpwr-1.pdf}

}

\end{figure}%

As should be the case, power estimates in Figure~\ref{fig-finalpwr}
increase with both the number of subjects and the number of items. The
CIs reported here indicate how precisely power was estimated by our
simulation. Higher precision (which would be reflected in narrower CIs)
could be obtained by increasing the number of repetitions (\texttt{rep})
in the simulation. In practice, data simulations are often run multiple
times with adjusted combinations of sample sizes. When running for the
first time, it might be revealed that power is way too low (or much
higher than required) for some combinations of \texttt{n\_subjects} and
\texttt{n\_items}. When narrowing down the best combination that
achieves sufficient power while at the same time striking a good balance
of how many subjects and items are practically feasible, later rounds of
data simulation will typically include a smaller grid of sample sizes
combined with a higher number of repetitions. This will assure high
precision for the final power estimates, which are then used for the
sample size justification of the future study.

Much has been written on the optimal amount of power to target in
empirical research. The most prominent heuristic is to target a power of
0.8 (when combined with a type I error rate of \(\alpha = 0.05\)), but
depending on the research goals of the study, there are often good
reasons to move away from this standard depending on the research goals
and resource constraints
(\citeproc{ref-lakensJustifyYourAlpha2018}{Lakens, Adolfi, et al.,
2018}; \citeproc{ref-lakensSampleSizeJustification2022}{Lakens, 2022a}).
When target power has been specified, the number of subjects and the
number of items in our study design can be traded against each other
based on practical considerations. For the sake of the example, let the
targeted power be indeed about 0.8, using an \(\alpha\) of 0.05 to
detect an effect of the expected size implied by our data simulation.
This could be achieved by collecting data from 200 subjects (about 25\%
of which will be experts), each completing the same 50 items (with
advice present in about 67\% of cases, which is correct in about 80\% of
cases with present advice). If collecting data from 200 subjects is not
feasible, an alternative would be to recruit 150 subjects but increase
the length of the experiment to over 70 items. However, 70 items might
take too long to complete for the radiologists participating in the
study, who have a busy schedule. The simulation suggests that it might
also be possible to plan a shorter experiment with only 30 items if it
is feasible to recruit an even higher number of subjects (\textgreater{}
250, to be determined by additional rounds of power analysis). Design
parameters that also affect power, and which could be investigated in
the simulation to find a more optimal trade-off, are the ratio of
experts, the frequency of whether advice is presented and whether it is
correct.

\paragraph{Precision results.}\label{precision-results}

When planning for precision, one could monitor the width of all four CIs
at the same time. However, because the CIs of the four contrasts
strongly differ in width, it is not trivial to decide which width one
should target when deciding on the appropriate sample size. In contrast
to planning for power, there are no common standards on how to specify
the targeted precision. For our example, we use a simple heuristic but
we strongly encourage readers to think about better alternatives that
are appropriate in their own applications. Our simulations show that the
smallest CI can be expected for the first contrast (expert with correct
advice vs.~expert without advice). The true contrast in probability for
an average expert and an average item in this condition is
\texttt{plogis(b\_0\ +\ b\_e\ +\ b\_a\ +\ b\_c\ +\ b\_ea\ +\ b\_ec)\ -\ plogis(b\_0\ +\ b\_e)\ =}
\(0.05\). We want the width of this CI to be smaller than 0.1. This
would mean that if the point estimate happens to be close to the true
value, the plausible values inside of a 95\% CI would all be positive.

Thus in our example, precision is defined as the (long-run) average
width of a 95\% CI for the probability contrast between experts with
correct advice and experts without advice. Of course, lower width
implies better precision. Based on our simulation outcomes, we compute
the precision estimate for each combination of \texttt{n\_subjects}
\(\times\) \texttt{n\_items} (including 95\% CIs) and visualize the
results with the following code.

\phantomsection\label{cell-fig-finalprecision}
\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{precision }\OtherTok{\textless{}{-}}\NormalTok{ sim\_result }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{pivot\_wider}\NormalTok{(}\AttributeTok{names\_from =}\NormalTok{ term, }\AttributeTok{names\_sep =} \StringTok{"\_"}\NormalTok{, }
    \AttributeTok{values\_from =}\NormalTok{ estimate}\SpecialCharTok{:}\NormalTok{p.value.equiv) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{group\_by}\NormalTok{(n\_subjects, n\_items) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{width =} \StringTok{\textasciigrave{}}\AttributeTok{conf.high\_b8=b2}\StringTok{\textasciigrave{}} \SpecialCharTok{{-}} \StringTok{\textasciigrave{}}\AttributeTok{conf.low\_b8=b2}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{precision =} \FunctionTok{mean}\NormalTok{(width),}
    \AttributeTok{ci.lwr =} \FunctionTok{t.test}\NormalTok{(width)}\SpecialCharTok{$}\NormalTok{conf.int[}\DecValTok{1}\NormalTok{],}
    \AttributeTok{ci.upr =} \FunctionTok{t.test}\NormalTok{(width)}\SpecialCharTok{$}\NormalTok{conf.int[}\DecValTok{2}\NormalTok{], }
    \AttributeTok{.groups =} \StringTok{"drop"}\NormalTok{)}
\NormalTok{precision }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{mutate}\NormalTok{(}\FunctionTok{across}\NormalTok{(}\FunctionTok{c}\NormalTok{(n\_subjects, n\_items), factor)) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(n\_subjects, n\_items, }\AttributeTok{fill =}\NormalTok{ precision)) }\SpecialCharTok{+}
  \FunctionTok{geom\_tile}\NormalTok{() }\SpecialCharTok{+}
  \FunctionTok{geom\_text}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{label =} \FunctionTok{sprintf}\NormalTok{(}\StringTok{"\%.2f }\SpecialCharTok{\textbackslash{}n}\StringTok{ [\%.2f; \%.2f]"}\NormalTok{, }
\NormalTok{                                precision, ci.lwr, ci.upr)), }
    \AttributeTok{color =} \StringTok{"white"}\NormalTok{, }\AttributeTok{size =} \DecValTok{4}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{scale\_fill\_viridis\_c}\NormalTok{(}\AttributeTok{limits =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\FloatTok{0.3}\NormalTok{), }\AttributeTok{direction =} \SpecialCharTok{{-}}\DecValTok{1}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{guides}\NormalTok{(}\AttributeTok{fill =} \FunctionTok{guide\_legend}\NormalTok{(}\AttributeTok{reverse=}\ConstantTok{FALSE}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{figure}[H]

\caption{\label{fig-finalprecision}Simulation-based precision estimates
(expected width of confidence intervals) including 95\% confidence
interval of the case study for different numbers of subjects and items,
based on a confidence level of 0.95.}

\centering{

\includegraphics{manuscript_files/figure-pdf/fig-finalprecision-1.pdf}

}

\end{figure}%

As should be the case, precision estimates in
Figure~\ref{fig-finalprecision} increase (i.e., average width of CI
decreases) with both the number of subjects and the number of items. The
CIs reported here indicate how precisely the expected width of the CI
for our focal contrast was estimated by our simulation. Applying our
simple heuristic of targeting an expected width smaller than 0.1, we see
the same trade-off between the number of subjects and the number of
items as with planning for power. We could either choose 100 subjects
and 30 items or 200 subjects and 10 items. Note that our simple
heuristic for determining sample size in the planning for precision
scenario was quite liberal. This is reflected by the result that we
would need a smaller sample size than in the planning for power
scenario. With a more conservative precision target, the result is
generally the opposite: As a rule, precise parameter estimates usually
require bigger samples than null hypothesis testing.

\subsection{Sensitivity analysis}\label{sensitivity-analysis}

In our case study, we have performed simulation-based sample size
planning from a single set of parameter values that reflect our
assumptions of an expected effect size. Instead of extracting this
expected effect size from meta-analyses or pilot data, which has been
the main focus of previous tutorials (e.g.,
\citeproc{ref-kumleEstimatingPowerGeneralized2021}{Kumle et al., 2021}),
we have demonstrated some strategies to determine plausible parameter
values in GLMMs based on domain knowledge. Domain knowledge can be
considered a vague theoretical model about the data-generating process
that is less formal and can only be accessed by a back-and-forth
exchange in which domain experts assess the plausibility of simulated
data. When sample sizes are chosen based on the results of our
simulation-based power analysis, a future study will be informative to
reject the null hypothesis if an effect of our \emph{expected size} is
present (or estimate the effect with satisfying precision). However, if
the true effect is indeed smaller, power (or precision) will be lower,
and the study might not be sufficiently informative. A common, more
conservative strategy for sample size justification is to perform sample
size planning for the smallest effect size of interest (SESOI). An
effect smaller than the SESOI would be considered too small to be
interesting or practically meaningful, even if the effect is not
actually zero (\citeproc{ref-kingPointMinimalImportant2011}{King, 2011};
\citeproc{ref-lakensEquivalenceTestingPsychological2018}{Lakens, Scheel,
et al., 2018}). For strategies on the even more difficult task of
specifying a plausible SESOI, as well as a thorough discussion of
various topics concerning power analysis, see
(\citeproc{ref-lakensImprovingYourStatistical2022}{Lakens, 2022b}). When
domain knowledge or formal theories about the research topic of interest
are too vague to specify a meaningful SESOI, it is still recommended to
demonstrate power or precision for different effect sizes in what is
called \emph{sensitivity power analysis}. By simulating power (or
precision) for different effect sizes (in addition to the different
number of subjects and items), one can make sure that power (or
precision) would still be sufficient to detect smaller effect sizes than
our expected effect or at least get an impression of how strongly power
(or precision) depends on the size of the true effect. For our case
study that investigates combined hypotheses in a GLMM modeling
framework, sensitivity analysis would require manually specifying
additional sets of plausible parameter values that reflect scenarios
with smaller or larger differences between groups with respect to our
specific research question. Power (or precision) could then be simulated
for several of these scenarios (across different numbers of subjects and
items, as considered earlier).

\section{Conclusion and outlook}\label{conclusion-and-outlook}

The goal of this tutorial was to teach researchers how to perform
tailored simulation-based sample size planning for GLMMs. Beyond the
specifics of our concrete case study, we want to outline six
developments regarding the future role of simulation-based sample size
planning in experimental research:

In light of the ongoing replication crisis and an existing literature
full of underpowered studies, there is a growing need for
simulation-based sample size planning in experimental research: In order
to conduct informative research, GLMMs offer a flexible statistical
framework to analyze complex experimental study designs. However,
existing formula-based heuristics and user-friendly software tools for a
priori power analysis are often not sufficient. Therefore,
simulation-based power analysis is becoming increasingly needed since it
provides experimental researchers with a tailored approach to estimating
required sample sizes before data collection.

Managing data simulations more easily with discrete predictor variables:
Simulation-based sample size planning becomes more manageable when all
predictor variables are discrete (like in the presented case study) and
fixed by the study design. This allows researchers to focus on
simulating outcome variables while avoiding the need for complex
simulations of predictor values, which would introduce additional
assumptions. By simplifying the simulation process, researchers can
obtain reliable estimates for power or precision without compromising
realistic assumptions about the data-generating process implied by the
study design.

Teaching data simulation skills: The ability to conduct simulation-based
sample size planning is a valuable skill that should be taught to
experimental researchers. By incorporating such training into research
methods courses and workshops, researchers can gain a deeper
understanding of statistical power or precision, and improve the quality
of their experimental designs. Equipping researchers with the knowledge
and tools to perform simulation-based sample size planning enables them
to make informed decisions and enhance the rigor of their studies. The
need to reason about how to simulate plausible data that is in line with
the research hypothesis, while not violating domain expertise on how
plausible data should look, might also contribute to planning more
insightful studies that can answer more precise research questions
(\citeproc{ref-yarkoniGeneralizabilityCrisis2022}{Yarkoni, 2022}).

Addressing the mismatch in effort perception: There is often a
significant disconnect between the amount of effort required to perform
tailored simulation-based sample size planning and the perceived effort
estimated by researchers and collaborators in experimental research.
Many researchers request simulation-based power analyses from
statisticians or methodological experts without fully comprehending the
complexity and time-consuming nature of these tailored simulations. It
is crucial to raise awareness about the effort involved to ensure
realistic expectations and effective collaboration between researchers
and methodological experts.

Recognizing the value of simulation-based design analysis: Tailored data
simulations and power analyses are not mere technicalities; they are
valuable research contributions that deserve recognition in experimental
research. They offer insights into the robustness and sensitivity of
experimental designs, helping researchers make informed decisions about
sample sizes, effect sizes, and statistical power or precision. Their
importance can be reflected by allocating them a separate publication or
incorporating them as a significant component of stage 1 preregistered
reports (\citeproc{ref-chambersPresentFutureRegistered2022}{Chambers \&
Tzavella, 2022}).

Integration with Open Science and preregistration practices:
Simulation-based sample size planning aligns well with the principles of
Open Science and preregistration in experimental research. When
researchers have access to simulated data based on their pre-specified
model, analyzing the collected dataset becomes straightforward and
unambiguous. By preregistering their simulation-based sample size plan,
researchers enhance the transparency and accountability of their
experimental procedures, contributing to the credibility and
reproducibility of research.

\section{References}\label{references}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-albersWhenPowerAnalyses2018a}
Albers, C., \& Lakens, D. (2018). When power analyses based on pilot
data are biased: {Inaccurate} effect size estimators and follow-up bias.
\emph{Journal of Experimental Social Psychology}, \emph{74}, 187--195.
\url{https://doi.org/10.1016/j.jesp.2017.09.004}

\bibitem[\citeproctext]{ref-R-marginaleffects}
Arel-Bundock, V. (2024). \emph{Marginaleffects: Predictions,
comparisons, slopes, marginal means, and hypothesis tests}.
\url{https://marginaleffects.com/}

\bibitem[\citeproctext]{ref-arendStatisticalPowerTwolevel2019}
Arend, M. G., \& Schäfer, T. (2019). Statistical power in two-level
models: {A} tutorial based on {Monte Carlo} simulation.
\emph{Psychological Methods}, \emph{24}(1), 1--19.
\url{https://doi.org/10.1037/met0000195}

\bibitem[\citeproctext]{ref-batesFittingLinearMixedEffects2015}
Bates, D., Mächler, M., Bolker, B., \& Walker, S. (2015). Fitting
{Linear Mixed-Effects Models Using} {\textbf{Lme4}}. \emph{Journal of
Statistical Software}, \emph{67}(1).
\url{https://doi.org/10.18637/jss.v067.i01}

\bibitem[\citeproctext]{ref-R-RJ-2021-048}
Bengtsson, H. (2021). A unifying framework for parallel and distributed
processing in r using futures. \emph{The R Journal}, \emph{13}(2),
208--227. \url{https://doi.org/10.32614/RJ-2021-048}

\bibitem[\citeproctext]{ref-brooksGlmmTMBBalancesSpeed2017}
Brooks, M., E., Kristensen, K., Benthem, van, J., Magnusson, A., Berg,
C., W., Nielsen, A., Skaug, H., J., Mächler, M., \& Bolker, B., M.
(2017). {glmmTMB Balances Speed} and {Flexibility Among Packages} for
{Zero-inflated Generalized Linear Mixed Modeling}. \emph{The R Journal},
\emph{9}(2), 378. \url{https://doi.org/10.32614/RJ-2017-066}

\bibitem[\citeproctext]{ref-brysbaertPowerAnalysisEffect2018}
Brysbaert, M., \& Stevens, M. (2018). Power {Analysis} and {Effect Size}
in {Mixed Effects Models}: {A Tutorial}. \emph{Journal of Cognition},
\emph{1}(1), 9. \url{https://doi.org/10.5334/joc.10}

\bibitem[\citeproctext]{ref-burknerBrmsPackageBayesian2017}
Bürkner, P.-C. (2017). Brms: {An R Package} for {Bayesian Multilevel
Models Using Stan}. \emph{Journal of Statistical Software}, \emph{80},
1--28. \url{https://doi.org/10.18637/jss.v080.i01}

\bibitem[\citeproctext]{ref-burknerAdvancedBayesianMultilevel2018}
Bürkner, P.-C. (2018). Advanced {Bayesian Multilevel Modeling} with the
{R Package} brms. \emph{The R Journal}, \emph{10}(1), 395.
\url{https://doi.org/10.32614/RJ-2018-017}

\bibitem[\citeproctext]{ref-chambersPresentFutureRegistered2022}
Chambers, C. D., \& Tzavella, L. (2022). The past, present and future of
{Registered Reports}. \emph{Nature Human Behaviour}, \emph{6}(1),
29--42. \url{https://doi.org/10.1038/s41562-021-01193-7}

\bibitem[\citeproctext]{ref-cohenPowerPrimer1992}
Cohen, J. (1992). A power primer. \emph{Psychological Bulletin},
\emph{112}(1), 155--159.
\url{https://doi.org/10.1037/0033-2909.112.1.155}

\bibitem[\citeproctext]{ref-cummingNewStatisticsWhy2014}
Cumming, G. (2014). The {New Statistics}: {Why} and {How}.
\emph{Psychological Science}, \emph{25}(1), 7--29.
\url{https://doi.org/10.1177/0956797613504966}

\bibitem[\citeproctext]{ref-R-faux}
DeBruine, L. (2023). \emph{Faux: Simulation for factorial designs}.
Zenodo. \url{https://doi.org/10.5281/zenodo.2669586}

\bibitem[\citeproctext]{ref-debruineUnderstandingMixedEffectsModels2021}
DeBruine, L., \& Barr, D. J. (2021). Understanding {Mixed-Effects Models
Through Data Simulation}. \emph{Advances in Methods and Practices in
Psychological Science}, \emph{4}(1), 2515245920965119.
\url{https://doi.org/10.1177/2515245920965119}

\bibitem[\citeproctext]{ref-deffnerCausalFrameworkCrossCultural2022}
Deffner, D., Rohrer, J. M., \& McElreath, R. (2022). A {Causal
Framework} for {Cross-Cultural Generalizability}. \emph{Advances in
Methods and Practices in Psychological Science}, \emph{5}(3),
251524592211063. \url{https://doi.org/10.1177/25152459221106366}

\bibitem[\citeproctext]{ref-dmitrienkoTraditionalMultiplicityAdjustment2013}
Dmitrienko, A., \& D'Agostino, R. (2013). Traditional multiplicity
adjustment methods in clinical trials. \emph{Statistics in Medicine},
\emph{32}(29), 5172--5218. \url{https://doi.org/10.1002/sim.5990}

\bibitem[\citeproctext]{ref-fahrmeirRegressionModelsMethods2021}
Fahrmeir, L., Kneib, T., Lang, S., \& Marx, B. D. (2021).
\emph{Regression: {Models}, {Methods} and {Applications}}. Springer
Berlin Heidelberg. \url{https://doi.org/10.1007/978-3-662-63882-8}

\bibitem[\citeproctext]{ref-gelmanBayesianWorkflow2020}
Gelman, A., Vehtari, A., Simpson, D., Margossian, C. C., Carpenter, B.,
Yao, Y., Kennedy, L., Gabry, J., Bürkner, P.-C., \& Modrák, M. (2020).
\emph{Bayesian {Workflow}} (arXiv:2011.01808). arXiv.
\url{https://arxiv.org/abs/2011.01808}

\bibitem[\citeproctext]{ref-gomilaMissingDataExperiments2022}
Gomila, R., \& Clark, C. S. (2022). Missing data in experiments:
{Challenges} and solutions. \emph{Psychological Methods}, \emph{27}(2),
143--155. \url{https://doi.org/10.1037/met0000361}

\bibitem[\citeproctext]{ref-greenSIMRPackagePower2016a}
Green, P., \& MacLeod, C. J. (2016). {\textsc{SIMR}} : An {R} package
for power analysis of generalized linear mixed models by simulation.
\emph{Methods in Ecology and Evolution}, \emph{7}(4), 493--498.
\url{https://doi.org/10.1111/2041-210X.12504}

\bibitem[\citeproctext]{ref-kainPracticalGuidePower2015}
Kain, M. P., Bolker, B. M., \& McCoy, M. W. (2015). A practical guide
and power analysis for {GLMMs}: Detecting among treatment variation in
random effects. \emph{PeerJ}, \emph{3}, e1226.
\url{https://doi.org/10.7717/peerj.1226}

\bibitem[\citeproctext]{ref-kelleySampleSizePlanning2006}
Kelley, K., \& Rausch, J. R. (2006). Sample size planning for the
standardized mean difference: {Accuracy} in parameter estimation via
narrow confidence intervals. \emph{Psychological Methods}, \emph{11}(4),
363--385. \url{https://doi.org/10.1037/1082-989X.11.4.363}

\bibitem[\citeproctext]{ref-kingPointMinimalImportant2011}
King, M. T. (2011). A point of minimal important difference ({MID}): A
critique of terminology and methods. \emph{Expert Review of
Pharmacoeconomics \& Outcomes Research}, \emph{11}(2), 171--184.
\url{https://doi.org/10.1586/erp.11.9}

\bibitem[\citeproctext]{ref-kruschkeBayesianNewStatistics2018}
Kruschke, J. K., \& Liddell, T. M. (2018). The {Bayesian New
Statistics}: {Hypothesis} testing, estimation, meta-analysis, and power
analysis from a {Bayesian} perspective. \emph{Psychonomic Bulletin \&
Review}, \emph{25}(1), 178--206.
\url{https://doi.org/10.3758/s13423-016-1221-4}

\bibitem[\citeproctext]{ref-kumleEstimatingPowerGeneralized2021}
Kumle, L., Võ, M. L.-H., \& Draschkow, D. (2021). Estimating power in
(generalized) linear mixed models: {An} open introduction and tutorial
in {R}. \emph{Behavior Research Methods}, \emph{53}(6), 2528--2543.
\url{https://doi.org/10.3758/s13428-021-01546-0}

\bibitem[\citeproctext]{ref-lafitSelectionNumberParticipants2021}
Lafit, G., Adolf, J. K., Dejonckheere, E., Myin-Germeys, I.,
Viechtbauer, W., \& Ceulemans, E. (2021). Selection of the {Number} of
{Participants} in {Intensive Longitudinal Studies}: {A User-Friendly
Shiny App} and {Tutorial} for {Performing Power Analysis} in {Multilevel
Regression Models That Account} for {Temporal Dependencies}.
\emph{Advances in Methods and Practices in Psychological Science},
\emph{4}(1), 251524592097873.
\url{https://doi.org/10.1177/2515245920978738}

\bibitem[\citeproctext]{ref-lakensSampleSizeJustification2022}
Lakens, D. (2022a). Sample {Size Justification}. \emph{Collabra:
Psychology}, \emph{8}(1), 33267.
\url{https://doi.org/10.1525/collabra.33267}

\bibitem[\citeproctext]{ref-lakensImprovingYourStatistical2022}
Lakens, D. (2022b). \emph{Improving {Your Statistical Inferences}}.
Zenodo. \url{https://doi.org/10.5281/ZENODO.6409077}

\bibitem[\citeproctext]{ref-lakensJustifyYourAlpha2018}
Lakens, D., Adolfi, F. G., Albers, C. J., Anvari, F., Apps, M. A. J.,
Argamon, S. E., Baguley, T., Becker, R. B., Benning, S. D., Bradford, D.
E., Buchanan, E. M., Caldwell, A. R., Van Calster, B., Carlsson, R.,
Chen, S.-C., Chung, B., Colling, L. J., Collins, G. S., Crook, Z.,
\ldots{} Zwaan, R. A. (2018). Justify your alpha. \emph{Nature Human
Behaviour}, \emph{2}(3), 168--171.
\url{https://doi.org/10.1038/s41562-018-0311-x}

\bibitem[\citeproctext]{ref-lakensEquivalenceTestingPsychological2018}
Lakens, D., Scheel, A. M., \& Isager, P. M. (2018). Equivalence
{Testing} for {Psychological Research}: {A Tutorial}. \emph{Advances in
Methods and Practices in Psychological Science}, \emph{1}(2), 259--269.
\url{https://doi.org/10.1177/2515245918770963}

\bibitem[\citeproctext]{ref-lanePowerStrugglesEstimating2018}
Lane, S. P., \& Hennes, E. P. (2018). Power struggles: {Estimating}
sample size for multilevel relationships research. \emph{Journal of
Social and Personal Relationships}, \emph{35}(1), 7--31.
\url{https://doi.org/10.1177/0265407517710342}

\bibitem[\citeproctext]{ref-littleStatisticalAnalysisMissing2014}
Little, R. J. A., \& Rubin, D. B. (2014). \emph{Statistical {Analysis}
with {Missing Data}} (2nd ed). Wiley.

\bibitem[\citeproctext]{ref-lundbergWhatYourEstimand2021}
Lundberg, I., Johnson, R., \& Stewart, B. M. (2021). What {Is Your
Estimand}? {Defining} the {Target Quantity Connects Statistical
Evidence} to {Theory}. \emph{American Sociological Review},
\emph{86}(3), 532--565. \url{https://doi.org/10.1177/00031224211004187}

\bibitem[\citeproctext]{ref-maxwellSampleSizePlanning2008}
Maxwell, S. E., Kelley, K., \& Rausch, J. R. (2008). Sample {Size
Planning} for {Statistical Power} and {Accuracy} in {Parameter
Estimation}. \emph{Annual Review of Psychology}, \emph{59}(1), 537--563.
\url{https://doi.org/10.1146/annurev.psych.59.103006.093735}

\bibitem[\citeproctext]{ref-mcelreathStatisticalRethinkingBayesian2020}
McElreath, R. (2020). \emph{Statistical {Rethinking}: {A Bayesian
Course} with {Examples} in {R} and {Stan}} (2nd ed.). {Chapman and
Hall/CRC}. \url{https://doi.org/10.1201/9780429029608}

\bibitem[\citeproctext]{ref-murayamaSummarystatisticsbasedPowerAnalysis2022}
Murayama, K., Usami, S., \& Sakaki, M. (2022). Summary-statistics-based
power analysis: {A} new and practical method to determine sample size
for mixed-effects modeling. \emph{Psychological Methods}.
\url{https://doi.org/10.1037/met0000330}

\bibitem[\citeproctext]{ref-preacherComputationalToolsProbing2006}
Preacher, K. J., Curran, P. J., \& Bauer, D. J. (2006). Computational
{Tools} for {Probing Interactions} in {Multiple Linear Regression},
{Multilevel Modeling}, and {Latent Curve Analysis}. \emph{Journal of
Educational and Behavioral Statistics}, \emph{31}(4), 437--448.
\url{https://doi.org/10.3102/10769986031004437}

\bibitem[\citeproctext]{ref-riesthuisSimulationBasedPowerAnalyses2024}
Riesthuis, P. (2024). Simulation-{Based Power Analyses} for the
{Smallest Effect Size} of {Interest}: {A Confidence-Interval Approach}
for {Minimum-Effect} and {Equivalence Testing}. \emph{Advances in
Methods and Practices in Psychological Science}, \emph{7}(2),
25152459241240722. \url{https://doi.org/10.1177/25152459241240722}

\bibitem[\citeproctext]{ref-schadHowCapitalizePriori2020}
Schad, D. J., Vasishth, S., Hohenstein, S., \& Kliegl, R. (2020). How to
capitalize on a priori contrasts in linear (mixed) models: {A} tutorial.
\emph{Journal of Memory and Language}, \emph{110}, 104038.
\url{https://doi.org/10.1016/j.jml.2019.104038}

\bibitem[\citeproctext]{ref-uyguntuncEpistemicPragmaticFunction2023}
Uygun Tunç, D., Tunç, M. N., \& Lakens, D. (2023). The epistemic and
pragmatic function of dichotomous claims based on statistical hypothesis
tests. \emph{Theory \& Psychology}, \emph{33}(3), 403--423.
\url{https://doi.org/10.1177/09593543231160112}

\bibitem[\citeproctext]{ref-R-furrr}
Vaughan, D., \& Dancho, M. (2022). \emph{Furrr: Apply mapping functions
in parallel using futures}. \url{https://github.com/DavisVaughan/furrr}

\bibitem[\citeproctext]{ref-westfallStatisticalPowerOptimal2014}
Westfall, J., Kenny, D. A., \& Judd, C. M. (2014). Statistical power and
optimal design in experiments in which samples of participants respond
to samples of stimuli. \emph{Journal of Experimental Psychology:
General}, \emph{143}, 2020--2045.
\url{https://doi.org/10.1037/xge0000014}

\bibitem[\citeproctext]{ref-wickhamWelcomeTidyverse2019}
Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L., François,
R., Grolemund, G., Hayes, A., Henry, L., Hester, J., Kuhn, M., Pedersen,
T., Miller, E., Bache, S., Müller, K., Ooms, J., Robinson, D., Seidel,
D., Spinu, V., \ldots{} Yutani, H. (2019). Welcome to the {Tidyverse}.
\emph{Journal of Open Source Software}, \emph{4}(43), 1686.
\url{https://doi.org/10.21105/joss.01686}

\bibitem[\citeproctext]{ref-wickhamDataScienceImport2023}
Wickham, H., Çetinkaya-Rundel, M., \& Grolemund, G. (2023). \emph{R for
data science: Import, tidy, transform, visualize, and model data} (2nd
edition). O'Reilly.

\bibitem[\citeproctext]{ref-yarkoniGeneralizabilityCrisis2022}
Yarkoni, T. (2022). The generalizability crisis. \emph{Behavioral and
Brain Sciences}, \emph{45}, e1.
\url{https://doi.org/10.1017/S0140525X20001685}

\bibitem[\citeproctext]{ref-zimmerSampleSizePlanning2022}
Zimmer, F., Henninger, M., \& Debelak, R. (2022). \emph{Sample {Size
Planning} for {Complex Study Designs}: {A Tutorial} for the mlpwr
{Package}}. PsyArXiv. \url{https://doi.org/10.31234/osf.io/r9w6t}

\end{CSLReferences}

\end{document}